### 2014-10-05 Paper/ms.tex
SimpleExample

\documentclass[a4paper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{dsfont}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

\title{Nested sampling with multiple scalars}
\author{Brendon J. Brewer}
\date{\vspace{3cm}This work is licensed under the Creative Commons Attribution-ShareAlike 3.0 Unported License. To view a copy of this license, visit {\tt http://creativecommons.org/licenses/by-sa/3.0/deed.en\_GB}.}

\begin{document}
\maketitle

\abstract{Nested Sampling is an effective and popular
Monte Carlo algorithm for Bayesian computation. It is based on exploring the
prior $\pi(\mathbf{x})$ and successively imposing constraints on the value of
the likelihood $\mathcal{L}(\mathbf{x})$ that compress the prior mass by a
``known'' factor. This enables the calculation of the marginal likelihood
or evidence
$\mathcal{Z} = \int \pi(\mathbf{x})\mathcal{L}(\mathbf{x}) \, d^N\mathbf{x}$ and
properties of the posterior, or any other distribution that is intermediate
between the prior and the posterior. However, in some inference and
statistical mechanics problems, there are two or more scalar functions of
$\mathbf{x}$ that are relevant. For example, we might be interested in the
properties of
$p(\mathbf{x} | \lambda_1, \lambda_2) \propto \pi(\mathbf{x})\exp\left[-\lambda_1 f_1(\mathbf{x}) - \lambda_2 f_2(\mathbf{x})\right]$
for many different values of $\lambda_1$ and $\lambda_2$, including the
normalisations or ``partition function'' $\mathcal{Z}(\lambda_1, \lambda_2)$.
This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.
}

If the prior is $\pi(\mathbf{x})$, then a family of canonical
probability distributions is given by
\begin{eqnarray}
p(\mathbf{x} | \lambda_1, ..., \lambda_n) = 
\frac{\pi(\mathbf{x})\exp(-\sum_{i=1}^n \lambda_i f_i(\mathbf{x}))}
{\mathcal{Z}(\lambda_1, ..., \lambda_n)}
\end{eqnarray}
This family of distributions arises from a MaxEnt update from the prior
$\pi(\mathbf{x})$ if the expected values of the functions
$\{f_i(\mathbf{x})\}$ are specified. In statistical mechanics it is useful to
quantify the properties of
$p(\mathbf{x} | \lambda_1, ..., \lambda_n)$ as a function of the
$\lambda$s.

\section{Nested sampling}

\section{Demonstration Example}
The prior is a uniform distribution over the $n$-dimensional unit hypercube:
\begin{eqnarray}
p(\mathbf{x}) &=& 1
\end{eqnarray}
as long as $x_i \in [0, 1]$ for all $i$. The scalars are:
\begin{eqnarray}
S_1(\mathbf{x}) &=& -\sum_{i=1}^n \left(x_i - 0.5\right)^2\\
S_2(\mathbf{x}) &=& -\sum_{i=1}^n \sin^2\left(4\pi x_i\right)
\end{eqnarray}

For this section, we used $n=1000$ dimensions.

\section{Example: Potts model with some embellishments}
TBD

\end{document}

### 2015-01-07 Paper/ms.tex
Notation

@@
 \author{Brendon J. Brewer}
-\date{\vspace{3cm}This work is licensed under the Creative Commons Attribution-ShareAlike 3.0 Unported License.
+\date{\vspace{2cm}This work is licensed under the Creative Commons Attribution-ShareAlike 3.0 Unported License.
 To view a copy of this license, visit {\tt http://creativecommons.org/licenses/by-sa/3.0/deed.en\_GB}.}
@@
 the likelihood $\mathcal{L}(\mathbf{x})$ that compress the prior mass by a
-``known'' factor. This enables the calculation of the marginal likelihood
+factor that is approximately known.
+This enables the calculation of the marginal likelihood
 or evidence
@@
 $\mathbf{x}$ that are relevant. For example, we might be interested in the
-properties of
-$p(\mathbf{x} | \lambda_1, \lambda_2) \propto \pi(\mathbf{x})\exp\left[-\lambda_1 f_1(\mathbf{x}) - \lambda_2 f_2(\mathbf{x})\right]$
-for many different values of $\lambda_1$ and $\lambda_2$, including the
-normalisations or ``partition function'' $\mathcal{Z}(\lambda_1, \lambda_2)$.
+properties of the canonical distributions
+$p(\mathbf{x} | \alpha, \beta) \propto \pi(\mathbf{x})\exp\left[\alpha f_1(\mathbf{x}) + \beta f_2(\mathbf{x})\right]$
+for many different values of $\alpha$ and $\beta$, including the
+normalisations or ``partition function'' $\mathcal{Z}(\alpha, \beta)$.
 This work describes
@@
 
+\section{Introduction}
 If the prior is $\pi(\mathbf{x})$, then a family of canonical

### 2015-01-19 Paper/ms.tex
A bit of writing

@@
 
+\newcommand{\xx}{\mathbf{x}}	% The unknown parameters
+\newcommand{\dx}{d^N\mathbf{x}} % Volume element in parameter space
 \renewcommand{\topfraction}{0.85}
@@
 \title{Nested sampling with multiple scalars}
-\author{Brendon J. Brewer}
-\date{\vspace{2cm}This work is licensed under the Creative Commons Attribution-ShareAlike 3.0 Unported License.
-To view a copy of this license, visit {\tt http://creativecommons.org/licenses/by-sa/3.0/deed.en\_GB}.}
+\author{Brendon J. Brewer and Ewan Cameron}
 
@@
 
-\abstract{Nested Sampling is an effective and popular
-Monte Carlo algorithm for Bayesian computation. It is based on exploring the
-prior $\pi(\mathbf{x})$ and successively imposing constraints on the value of
-the likelihood $\mathcal{L}(\mathbf{x})$ that compress the prior mass by a
+\abstract{The abstract goes here.}
+
+\section{Introduction}
+Nested Sampling \citep{skilling} is an effective and popular
+Monte Carlo algorithm for Bayesian computation and statistical mechanics.
+In a Bayesian inference problem with unknown parameters denoted collectively
+by a vector $\mathbf{x}$, the
+posterior distribution for the parameters is
+\begin{eqnarray}
+p(\mathbf{x} | {\rm data}) &=&
+\frac{p(\mathbf{x})p({\rm data} | \mathbf{x})}{p(\rm data)}\\
+&=& \frac{\pi(\xx)L(\xx)}{Z}
+\end{eqnarray}
+where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
+function, and $Z$ is the normalising constant, known as the
+``marginal likelihood'' or the ``evidence'':
+\begin{eqnarray}
+Z &=& \int \pi(\xx) L(\xx) \, \dx.
+\end{eqnarray}
+
+Nested Sampling begins by drawing particles from the
+prior $\pi(\xx)$ and successively imposing constraints on the value of
+the likelihood $L(\xx)$ that compress the prior mass by a
 factor that is approximately known.
 This enables the calculation of the marginal likelihood
-or evidence
-$\mathcal{Z} = \int \pi(\mathbf{x})\mathcal{L}(\mathbf{x}) \, d^N\mathbf{x}$ and
-properties of the posterior, or any other distribution that is intermediate
-between the prior and the posterior. However, in some inference and
+and properties of the posterior, or any other distribution that is intermediate
+between the prior and the posterior.
+
+Relative to alternative approaches such as annealing, Nested Sampling continues
+to work when the problem contains a {\it phase transition}.
+
+In some inference and
 statistical mechanics problems, there are two or more scalar functions of
-$\mathbf{x}$ that are relevant. For example, we might be interested in the
+$\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
+we obtain testable information that fixes the expected values of two scalar
+functions of $\xx$, $S_1(\xx)$ and $S_2(\xx)$. It is well known that the
+updated probability distribution that takes into account the constraints is
+of the ``canonical'' form:
+
+\begin{eqnarray}
+p(\xx; \alpha_1, \alpha_2) &=& \frac{\pi(\xx)\exp\left[\alpha_1S_1(\xx)+\alpha_2S_2(\xx)\right]}
+{Z(\alpha_1, \alpha_2)}
+\end{eqnarray}
+
+ For example, we might be interested in the
 properties of the canonical distributions
@@
 transitions.
-}
 
-\section{Introduction}
 If the prior is $\pi(\mathbf{x})$, then a family of canonical
@@
 
+\begin{thebibliography}{99}
+
+\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+
+\end{thebibliography}
+
 \end{document}

### 2015-01-19 Paper/ms.tex
A bit of writing

@@
 and properties of the posterior, or any other distribution that is intermediate
-between the prior and the posterior.
+between the prior and the posterior. For example, we might be interested in
+a ``power posterior'' where the likelihood is raised to a power $\beta$:
+\begin{eqnarray}
+p(\xx; \beta) &=& \frac{\pi(\xx)L(\xx)^\beta}{Z(\beta)}
+\end{eqnarray}
+The normalisation and posterior samples from this distribution can be obtained
+from the original Nested Sampling run. An example application, in the case of
+a Bayesian model with a ``gaussian noise'' assumption in the likelihood,
+computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
+posterior distribution would have been if the noise variance had been greater.
 
@@
 
- For example, we might be interested in the
-properties of the canonical distributions
-$p(\mathbf{x} | \alpha, \beta) \propto \pi(\mathbf{x})\exp\left[\alpha f_1(\mathbf{x}) + \beta f_2(\mathbf{x})\right]$
-for many different values of $\alpha$ and $\beta$, including the
-normalisations or ``partition function'' $\mathcal{Z}(\alpha, \beta)$.
-This work describes
+If we were only interested in a single canonical distribution, for example
+with $\alpha_1 = 0.3$ and $\alpha_2 = 0.7$, we could estimate its normalising
+constant and by running standard Nested Sampling with ``likelihood''
+$L(\xx) = \exp\left[0.3S_1(\xx) + 0.7S_2(\xx)\right]$. However, usually we
+are interested in a range of values for $\alpha_1$ and $\alpha_2$, and we
+want to know the normalisation $Z(\alpha_1, \alpha_2)$, called the
+partition function. This work describes
 progress towards solving this class of problems while maintaining the benefits
@@
 
-If the prior is $\pi(\mathbf{x})$, then a family of canonical
-probability distributions is given by
-\begin{eqnarray}
-p(\mathbf{x} | \lambda_1, ..., \lambda_n) = 
-\frac{\pi(\mathbf{x})\exp(-\sum_{i=1}^n \lambda_i f_i(\mathbf{x}))}
-{\mathcal{Z}(\lambda_1, ..., \lambda_n)}
-\end{eqnarray}
-This family of distributions arises from a MaxEnt update from the prior
-$\pi(\mathbf{x})$ if the expected values of the functions
-$\{f_i(\mathbf{x})\}$ are specified. In statistical mechanics it is useful to
-quantify the properties of
-$p(\mathbf{x} | \lambda_1, ..., \lambda_n)$ as a function of the
-$\lambda$s.
-
-\section{Nested sampling}
+\section{Properties of Nested sampling}
+We seek an algorithm for computing the partition function
+$Z(\alpha_1, \alpha_2)$ for a range of values of the $\alpha$s, from a single
+run. To be considered a variant of Nested Sampling, we require that the method
+satisfies the following requirements:
+\begin{itemize}
+\item The algorithm should begin with $n$ points drawn from the prior $\pi(\xx)$.
+\item The algorithm should seek to explore regions where the values of
+$S_1(\xx)$ and $S_2(\xx)$ are greater than what would be expected from the
+prior.
+\item The algorithm should try to move through a sequence of probability
+distributions whose enclosed prior masses are approximately known.
+\item The algorithm should be invariant to monotonic transformations of
+$S_1$ and $S_2$, i.e. it should only depend on rankings of $S_1$ and $S_2$
+values and not the values themselves.
+\end{itemize}
 

### 2015-05-29 Paper/ms.tex
References and notation

@@
 \newcommand{\xx}{\mathbf{x}}	% The unknown parameters
+\newcommand{\data}{\mathbf{D}}  % The data
 \newcommand{\dx}{d^N\mathbf{x}} % Volume element in parameter space
@@
 Nested Sampling \citep{skilling} is an effective and popular
-Monte Carlo algorithm for Bayesian computation and statistical mechanics.
+Monte Carlo algorithm for Bayesian computation and
+statistical mechanics
+\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
 In a Bayesian inference problem with unknown parameters denoted collectively
-by a vector $\mathbf{x}$, the
-posterior distribution for the parameters is
+by a vector $\xx$, the
+posterior distribution for the parameters given data $\data$ (and prior
+information $I$) is:
 \begin{eqnarray}
-p(\mathbf{x} | {\rm data}) &=&
-\frac{p(\mathbf{x})p({\rm data} | \mathbf{x})}{p(\rm data)}\\
+p(\xx | \data, I) &=&
+\frac{p(\xx | I)p(\data | \xx, I)}{p(\data | I)}\\
 &=& \frac{\pi(\xx)L(\xx)}{Z}
@@
 \begin{eqnarray}
-p(\mathbf{x}) &=& 1
+p(\xx) &=& 1
 \end{eqnarray}
@@
 \begin{eqnarray}
-S_1(\mathbf{x}) &=& -\sum_{i=1}^n \left(x_i - 0.5\right)^2\\
-S_2(\mathbf{x}) &=& -\sum_{i=1}^n \sin^2\left(4\pi x_i\right)
+S_1(\xx) &=& -\sum_{i=1}^n \left(x_i - 0.5\right)^2\\
+S_2(\xx) &=& -\sum_{i=1}^n \sin^2\left(4\pi x_i\right)
 \end{eqnarray}
@@
 
+\bibitem[Baldock et al.(2015)]{2015arXiv150303404B} Baldock, R.~J.~N., 
+P{\'a}rtay, L.~v.~B., Bart{\'o}k, A.~P., Payne, M.~C., Cs{\'a}nyi, G.\ 
+2015.\ Determining pressure-temperature phase diagrams of materials.\ ArXiv 
+e-prints arXiv:1503.03404.
+
+\bibitem[P{\'a}rtay et al.(2014)]{2014PhRvE..89b2302P} P{\'a}rtay, L.~B., 
+Bart{\'o}k, A.~P., Cs{\'a}nyi, G.\ 2014.\ Nested sampling for materials: 
+The case of hard spheres.\ Physical Review E 89, 022302.
+
+\bibitem[P{\'a}rtay et al.(2009)]{2009arXiv0906.3544P} P{\'a}rtay, L.~B., 
+Bart{\'o}k, A.~P., Cs{\'a}nyi, G.\ 2009.\ Efficient sampling of atomic 
+configurational spaces.\ ArXiv e-prints arXiv:0906.3544.
+
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.

### 2015-08-20 Paper/ms.tex
Corner mass

@@
 
+\section{A Diffusive Variant}
+Have one walker, or a few (not a large number).
+
+Explore the prior $\pi(\xx)$, and save sampled values of the scalars
+(thinned, presumably). As a function of position $(s_1, s_2)$ there is a well
+defined value of the cumulative prior mass
+\begin{eqnarray}
+M(s_1, s_2) &=& \int \pi(\xx)\mathds{1}\left(S_1(\xx) \leq s_1,
+S_2(\xx) \leq s_2\right)\, d\xx
+\end{eqnarray}
+which can be approximated in a Monte Carlo fashion by
+\begin{eqnarray}
+\hat{M}(s_1, s_2) &=& \frac{1}{N}\sum_{i=1}^N \mathds{1}\left(S_1^i \leq s_1,
+S_2^i \leq s_2\right)
+\end{eqnarray}
+
 \begin{thebibliography}{99}

### 2015-09-15 Code/truth.py
interpolation

@@ -34,13 +34,13 @@ def grid():
 
   figure(1)
   subplot(1, 2, 1)
-  imshow(logZ, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())))
+  imshow(logZ, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest')
   xlabel('ln($T_1$)')
   ylabel('ln($T_2$)')
   title('log(Z)')
 
   subplot(1, 2, 2)
-  imshow(H, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())))
+  imshow(H, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest')
   xlabel('ln($T_1$)')
   ylabel('ln($T_2$)')
   title('H, max={m}'.format(m=H.max()))

### 2015-09-15 Paper/ms.tex
Entropy format

@@
-\documentclass[a4paper, 11pt]{article}
-\usepackage{graphicx}
-\usepackage{natbib}
+%  LaTeX support: latex@mdpi.com
+%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+%=================================================================
+
+% LaTeX Class File and Rendering Mode (choose one)
+% You will need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.
+
+%=================================================================
+
+\documentclass[journal,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
+%--------------------
+% Class Options:
+%--------------------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% actuators, administrativesciences, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences, arts, atmosphere, atoms, axioms, batteries, behavioralsciences, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsciences, buildings, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, cosmetics, crystals, data, dentistryjournal, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, inorganics, insects, ijerph, ijfs, ijms, ijns, ijgi, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, joi, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, lubricants, machines, marinedrugs, materials, mathematics, medicalsciences, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, sinusitis, socialsciences, societies, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, universe, vaccines, veterinarysciences, viruses, water
+%---------
+% article
+%---------
+% The default type of manuscript is article, but could be replaced by using one of the class options:
+% article, review, communication, commentary, bookreview, correction, addendum, editorial, changes, supfile, casereport, comment, conceptpaper, conferencereport, meetingreport, discussion, essay, letter, newbookreceived, opinion, projectreport, reply, retraction, shortnote, technicalnote, creative, datadescriptor (for journal Data), briefreport, hypothesis, interestingimage
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g.
+the logo of the journal will get visible), the headings, and the copyright information.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+% Please insert a blank line is before and after all equation and eqnarray environments to ensure proper line numbering when option submit is chosen
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+%---------
+% pdftex
+%---------
+% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use the optioin "dvipdfm", with LaTeX and dvi2pdf only.
+
+%=================================================================
+\setcounter{page}{1}
+\lastpage{x}
+\doinum{10.3390/------}
+\pubvolume{xx}
+\pubyear{2015}
+%\externaleditor{Academic Editor: xx}
+\history{Received: xx / Accepted: xx / Published: xx}
+
 \usepackage{dsfont}
-\usepackage[left=3cm,top=3cm,right=3cm]{geometry}
 
-\newcommand{\xx}{\mathbf{x}}	% The unknown parameters
-\newcommand{\data}{\mathbf{D}}  % The data
-\newcommand{\dx}{d^N\mathbf{x}} % Volume element in parameter space
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\parindent=0cm
+\newcommand{\xx}{\boldsymbol{x}}
+\newcommand{\dx}{d\boldsymbol{x}}
+\newcommand{\data}{\boldsymbol{D}}
 
-\title{Nested sampling with multiple scalars}
-\author{Brendon J. Brewer and Ewan Cameron}
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+
+% Add packages and commands to include here
+% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are loaded by the MDPI class.
+%\usepackage{graphicx}
+%\usepackage{subfigure,psfig}
+
+%=================================================================
+%% Please use the following mathematics environments:
+% \theoremstyle{mdpi}
+% \newcounter{thm}
+% \setcounter{thm}{0}
+% \newcounter{ex}
+% \setcounter{ex}{0}
+% \newcounter{re}
+% \setcounter{re}{0}
+%
+% \newtheorem{Theorem}[thm]{Theorem}
+% \newtheorem{Lemma}[thm]{Lemma}
+% \newtheorem{Corollary}[thm]{Corollary}
+% \newtheorem{Proposition}[thm]{Proposition}
+%
+% \theoremstyle{mdpidefinition}
+% \newtheorem{Characterization}[thm]{Characterization}
+% \newtheorem{Property}[thm]{Property}
+% \newtheorem{Problem}[thm]{Problem}
+% \newtheorem{Example}[ex]{Example}
+% \newtheorem{ExamplesandDefinitions}[ex]{Examples and Definitions}
+% \newtheorem{Remark}[re]{Remark}
+% \newtheorem{Definition}[thm]{Definition}
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+
+% Full title of the paper (Capitalized)
+\Title{Nested sampling with multiple scalars}
+
+% Authors (Add full first names)
+\Author{Brendon J. Brewer$^{1,}$* and Ewan Cameron$^{2}$}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ Department of Statistics, The University of Auckland, Private Bag 92019,
+Auckland 1142, New Zealand\\
+$^{2}$ Spatial Ecology and Epidemiology Group, Tinbergen Building, Department
+of Zoology, University of Oxford, South Parks Road, Oxford, UK}
+
+%\contributed{$^\dagger$ These authors contributed equally to this work.}
+
+% Contact information of the corresponding author (Add [2] after \corres if there are more than one corresponding author.)
+\corres{{\tt bj.brewer@auckland.ac.nz}}
+
+% Abstract (Do not use inserted blank lines, i.e. \\)
+\abstract{The abstract goes here.}
+
+% Keywords: add 3 to 10 keywords
+\keyword{keyword; keyword; keyword}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{}
+%\MSC{}
+%\JEL{}
+
+% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\dagger$ in the end of the title of this article.
+%\conference{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% For journal Data:
+
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\maketitle
 
-\abstract{The abstract goes here.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{thebibliography}{99}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\authorcontributions{Author Contributions}
+
+Required if more than one author.
+Authorship must include and be strictly limited to researchers who have substantially contributed to the reported work.
+Please carefully review our criteria regarding the Qualification for Authorship: \web /instructions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\conflictofinterests{Conflicts of Interest}
+The authors declare no conflict of interest.
+
+%=================================================================
+% References: Variant A
+%=================================================================
+% Back Matter (References and Notes)
+%----------------------------------------------------------
+% Style and layout of the references
+\bibliographystyle{mdpi}
+\makeatletter
+\renewcommand\@biblabel[1]{#1. }
+\makeatother
+
+\begin{thebibliography}{999} % if there are less than 10 entries, enter a one digit number
 
@@
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+%% Reference 1
+%\bibitem{ref-journal}
+%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
+
+%% Reference 2
+%\bibitem{ref-book}
+%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
 
@@
 
+%=================================================================
+% References:  Variant B
+%=================================================================
+% Use the following option to include external BibTeX files:
+%\bibliography{lite}
+%\bibliographystyle{mdpi}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\abbreviations{Abbreviations/Nomenclature}
+%
+%Main text.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\appendix
+%\section{Appendix Title}
+%
+%Main text.
+
 \end{document}

### 2015-09-15 Paper/ms.tex
Spaces and acknowledgements

@@
 % Keywords: add 3 to 10 keywords
-\keyword{keyword; keyword; keyword}
+\keyword{nested sampling; bayesian computation; statistical mechanics}
 
@@
 \section{Introduction}
+
 Nested Sampling \citep{skilling} is an effective and popular
@@
 \section{Properties of Nested sampling}
+
 We seek an algorithm for computing the partition function
@@
 \section{Example: Potts model with some embellishments}
+
 TBD
@@
 \section{A Diffusive Variant}
+
 Have one walker, or a few (not a large number).
@@
 \end{eqnarray}
+
+\acknowledgments{Acknowledgements}
+
+It is a pleasure to thank Gábor Csányi (Cambridge), Livia B. Pártay (Cambridge),
+and Robert Baldock (Cambridge) for valuable conversations.
 

### 2015-09-15 Code/C++/SamplerImpl.h
Got rid of warning

@@ -228,7 +228,7 @@ double Sampler<Type>::compute_frac_below(int i,
 	double result = 0.;
 	for(size_t j=0; j<keep.size(); j++)
 	{
-		if(i != j)
+		if(i != static_cast<int>(j))
 			result += is_below(keep[j], keep[i],
 						keep_tiebreakers[j],
 						keep_tiebreakers[i]);

### 2015-09-29 Paper/ms.tex
A bit of writing

@@
 \end{eqnarray}
-
-Nested Sampling begins by drawing particles from the
+It begins by drawing particles from the
 prior $\pi(\xx)$ and successively imposing constraints on the value of
@@
 factor that is approximately known.
-This enables the calculation of the marginal likelihood
-and properties of the posterior, or any other distribution that is intermediate
+This enables the calculation of the marginal likelihood $Z$
+and properties of the posterior $\pi(\xx)L(\xx)/Z$.
+
+NS also allows us to calculate the properties of any other distribution that
+is in some sense intermediate
 between the prior and the posterior. For example, we might be interested in
@@
 \begin{eqnarray}
-p(\xx; \beta) &=& \frac{\pi(\xx)L(\xx)^\beta}{Z(\beta)}
+p(\xx; \beta) &=& \frac{\pi(\xx)L(\xx)^\beta}{Z(\beta)}\label{eqn:power_posterior}
 \end{eqnarray}
@@
 posterior distribution would have been if the noise variance had been greater.
-
 Relative to alternative approaches such as annealing, Nested Sampling continues
 to work when the problem contains a {\it phase transition}.
+
+In statistical mechanics, Equation~\ref{eqn:power_posterior} defines the
+family of {\it canonical distributions}, usually written as:
+\begin{eqnarray}
+p(\xx; \beta) &=& \frac{\pi(\xx)\exp[-\beta E(\xx)}{Z(\beta)}
+\end{eqnarray}
+where $E(\xx)$ is the energy function. In this context we usually want to
+compute $Z(\beta)$ as a function of $\beta$, which is called the
+{\it partition function}.
 

### 2015-09-29 Paper/ms.tex
Enumerate instead of itemize

@@
 $Z(\alpha_1, \alpha_2)$ for a range of values of the $\alpha$s, from a single
-run. To be considered a variant of Nested Sampling, we require that the method
-satisfies the following requirements:
-\begin{itemize}
-\item The algorithm should begin with $n$ points drawn from the prior $\pi(\xx)$.
+run. To be considered a variant of Nested Sampling, we require the method
+to satisfy the following requirements:
+\begin{enumerate}
+\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
 \item The algorithm should seek to explore regions where the values of
@@
 \item The algorithm should try to move through a sequence of probability
-distributions whose enclosed prior masses are approximately known.
+distributions proportional to $\pi$, but restricted to smaller and smaller
+domains for which we can approximately measure the enclosed prior mass.
 \item The algorithm should be invariant to monotonic transformations of
@@
 values and not the values themselves.
-\end{itemize}
+\end{enumerate}
 

### 2015-09-29 Paper/ms.tex
Walter reference

@@
 
-Nested Sampling \citep{skilling} is an effective and popular
+Nested Sampling \citep[NS][]{skilling} is an effective and popular
 Monte Carlo algorithm for Bayesian computation and
@@
 \begin{eqnarray}
-Z &=& \int \pi(\xx) L(\xx) \, \dx.
+Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
 \end{eqnarray}
@@
 and properties of the posterior $\pi(\xx)L(\xx)/Z$.
+NS is required because Equation~\ref{eqn:evidence}
+is the expected value of $L$ with respect to a very heavy-tailed distribution
+(the prior for the $L$ value implied by $\pi(\xx)$). Hence, there is a strong
+connection between NS and ideas from rare event simulation \citep{walter}.
 
@@
 \begin{eqnarray}
-p(\xx; \beta) &=& \frac{\pi(\xx)\exp[-\beta E(\xx)}{Z(\beta)}
+p(\xx; \beta) &=& \frac{\pi(\xx)\exp[-\beta E(\xx)]}{Z(\beta)}
 \end{eqnarray}
@@
 2015.\ Determining pressure-temperature phase diagrams of materials.\ ArXiv 
-e-prints arXiv:1503.03404.
+e-prints arXiv: 1503.03404.
 
@@
 Bart{\'o}k, A.~P., Cs{\'a}nyi, G.\ 2009.\ Efficient sampling of atomic 
-configurational spaces.\ ArXiv e-prints arXiv:0906.3544.
+configurational spaces.\ ArXiv e-prints arXiv: 0906.3544.
 
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+
+\bibitem[Walter(2015)]{walter}
+Walter, C.\ Point Process-based Monte Carlo estimation.\ arXiv: 1412.6368.
 %% Reference 1

### 2015-09-29 Paper/ms.tex
inputenc
alpha --> beta

@@
 \usepackage{dsfont}
+\usepackage[utf8]{inputenc}
 
@@
 \begin{eqnarray}
-p(\xx; \alpha_1, \alpha_2) &=& \frac{\pi(\xx)\exp\left[\alpha_1S_1(\xx)+\alpha_2S_2(\xx)\right]}
-{Z(\alpha_1, \alpha_2)}
+p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1S_1(\xx)+\beta_2S_2(\xx)\right]}
+{Z(\beta_1, \beta_2)}
 \end{eqnarray}
@@
 If we were only interested in a single canonical distribution, for example
-with $\alpha_1 = 0.3$ and $\alpha_2 = 0.7$, we could estimate its normalising
+with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
 constant and by running standard Nested Sampling with ``likelihood''
 $L(\xx) = \exp\left[0.3S_1(\xx) + 0.7S_2(\xx)\right]$. However, usually we
-are interested in a range of values for $\alpha_1$ and $\alpha_2$, and we
-want to know the normalisation $Z(\alpha_1, \alpha_2)$, called the
+are interested in a range of values for $\beta_1$ and $\beta_2$, and we
+want to know the normalisation $Z(\beta_1, \beta_2)$, called the
 partition function. This work describes
@@
 We seek an algorithm for computing the partition function
-$Z(\alpha_1, \alpha_2)$ for a range of values of the $\alpha$s, from a single
+$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
 run. To be considered a variant of Nested Sampling, we require the method
@@
 \end{enumerate}
+
+The algorithm works as follows.
 

### 2015-10-04 Paper/ms.tex
Place to ramble about algorithm possibilities

@@
 
+\section{Draft algorithm}
+
 The algorithm works as follows.
-
-\section{Demonstration Example}
-The prior is a uniform distribution over the $n$-dimensional unit hypercube:
+Start with $N$ particles drawn from $\pi$.
+Define the {\it corner count} of particle $i$ by
 \begin{eqnarray}
-p(\xx) &=& 1
-\end{eqnarray}
-as long as $x_i \in [0, 1]$ for all $i$. The scalars are:
-\begin{eqnarray}
-S_1(\xx) &=& -\sum_{i=1}^n \left(x_i - 0.5\right)^2\\
-S_2(\xx) &=& -\sum_{i=1}^n \sin^2\left(4\pi x_i\right)
+C_i &=& \sum_{j=1}^N \mathds{1}\left(S_1^j < S_1^i, S_2^j < S_2^i\right)
 \end{eqnarray}
 
-For this section, we used $n=1000$ dimensions.
-
-\section{Example: Potts model with some embellishments}
-
-TBD
-
-\section{A Diffusive Variant}
-
-Have one walker, or a few (not a large number).
-
-Explore the prior $\pi(\xx)$, and save sampled values of the scalars
-(thinned, presumably). As a function of position $(s_1, s_2)$ there is a well
-defined value of the cumulative prior mass
-\begin{eqnarray}
-M(s_1, s_2) &=& \int \pi(\xx)\mathds{1}\left(S_1(\xx) \leq s_1,
-S_2(\xx) \leq s_2\right)\, d\xx
-\end{eqnarray}
-which can be approximated in a Monte Carlo fashion by
-\begin{eqnarray}
-\hat{M}(s_1, s_2) &=& \frac{1}{N}\sum_{i=1}^N \mathds{1}\left(S_1^i \leq s_1,
-S_2^i \leq s_2\right)
-\end{eqnarray}
+Need to be able to sort particles into a total order. So add tiebreaker.
 

### 2015-10-08 Paper/ms.tex
Words
Easier-to-understand title

@@
 % Full title of the paper (Capitalized)
-\Title{Nested sampling with multiple scalars}
+\Title{Nested sampling with two or more objective functions}
 
@@
 
-Need to be able to sort particles into a total order. So add tiebreaker.
+Need to be able to sort particles into a total order. Since $C_i$ is discrete,
+each particle will also need a tiebreaker variable $a_i$ to accompany it.
+For iteration 1, $\{(C_i, a_i)\}_{i=1}^N$ will provide the total ordering.
+We exclude the corner with the lowest corner count.
+
+For iteration 2, we need a new quantity for total order (I think?), which
 

### 2015-10-09 Code/C++/SamplerImpl.h
Had wrong mass. Also choose scalar to sort by (within corner) at random

@@ -82,34 +82,35 @@ void Sampler<MyModel>::do_iteration()
 	int choice = usable_indices[rng.rand_int(usable_indices.size())];
 
 	// Find min of scalar 2 among particles in the rectangle
-	std::vector<double> s2;
+	int which_scalar = rng.rand_int(2);
+	std::vector<double> ss;
 	std::vector<int> indices;
 	for(int i=0; i<num_particles; i++)
 	{
 		if(is_in_lower_rectangle(particles[i].get_scalars(),
 										particles[choice].get_scalars()))
 		{
-			s2.push_back(particles[i].get_scalars()[1]);
+			ss.push_back(particles[i].get_scalars()[which_scalar]);
 			indices.push_back(i);
 		}
 	}
 
-	double s2_min = *min_element(s2.begin(), s2.end());
+	double ss_min = *min_element(ss.begin(), ss.end());
 	int which = 0;
-	for(size_t i=0; i<s2.size(); i++)
-		if(s2[i] == s2_min)
+	for(size_t i=0; i<ss.size(); i++)
+		if(ss[i] == ss_min)
 			which = indices[i];
 
 	std::vector<double> forbid(2);
-	forbid[0] = particles[choice].get_scalars()[0];
-	forbid[1] = s2_min;
+	forbid = particles[choice].get_scalars();
+	forbid[which_scalar] = particles[which].get_scalars()[which_scalar];
 
 	// Append its scalars to the forbidden rectangles
 	rects.push_front(forbid);
 	prune_rectangles();
 
 	// Assign prior weight
-	double logw = log((double)lccs[choice]/(num_particles + 1)) + log_prior_mass;
+	double logw = log((double)1./(num_particles + 1)) + log_prior_mass;
 	log_prior_mass = logdiffexp(log_prior_mass, logw);
 
 	// Write it out to an output file

### 2015-10-09 Code/C++/SamplerImpl.h
Back to min_element

@@ -68,10 +68,7 @@ void Sampler<MyModel>::do_iteration()
 		if(lccs[i] == 0)
 			lccs[i] = num_particles;
 	}
-	int min_lcc = num_particles;
-	for(int i=0; i<num_particles; i++)
-		if(lccs[i] < min_lcc)
-			min_lcc = lccs[i];
+	int min_lcc = *min_element(lccs.begin(), lccs.end());
 
 	std::vector<int> usable_indices;
 	for(int i=0; i<num_particles; i++)

### 2015-10-09 Paper/ms.tex
Editing intro

@@
 posterior distribution would have been if the noise variance had been greater.
-Relative to alternative approaches such as annealing, Nested Sampling continues
-to work when the problem contains a {\it phase transition}.
+This is different from including an extra parameter to allow the noise variance
+to be greater, because NS allows you to test the consequences of values of the
+variance that are very implausible given the data.
+
+Alternatives to NS include methods based on ``annealing'', where a sequence of
+distributions of the form of Equation~\ref{eqn:power_posterior} is used.
+There are many different methods based on this idea, such as the popular
+parallel tempering method \citep{pt}.
+However, unlike annealing methods, NS continues
+to work when the problem contains a phase transition.
 
@@
 
+\bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.
+
 \bibitem[P{\'a}rtay et al.(2014)]{2014PhRvE..89b2302P} P{\'a}rtay, L.~B., 

### 2015-10-09 Paper/ms.tex
More hacking on intro

@@
 \end{eqnarray}
-It begins by drawing particles from the
+NS begins by drawing particles from the
 prior $\pi(\xx)$ and successively imposing constraints on the value of
-the likelihood $L(\xx)$ that compress the prior mass by a
-factor that is approximately known.
+the likelihood $L(\xx)$ that compress the prior mass by an
+approximately-known factor.
 This enables the calculation of the marginal likelihood $Z$
@@
 The normalisation and posterior samples from this distribution can be obtained
-from the original Nested Sampling run. An example application, in the case of
-a Bayesian model with a ``gaussian noise'' assumption in the likelihood,
+from the original NS run by re-weighting the output according to $L(\xx)^\beta$
+instead of the usual $L(\xx)$ used to obtain the posterior.
+An example application, in the case of
+a Bayesian inference problem with a ``gaussian noise'' likelihood assumption,
 computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
@@
 \end{eqnarray}
-where $E(\xx)$ is the energy function. In this context we usually want to
+where $E(\xx)$ is the energy function, and $\pi(\xx)$ is often uniform over
+phase space (the set of possible positions and momenta of a collection of
+particles) or configuration space (the set of possible positions of a collection
+of particles). In this context we usually want to
 compute $Z(\beta)$ as a function of $\beta$, which is called the
-{\it partition function}.
+{\it partition function}. NS can achieve this, allowing the study of
+the properties of materials from first principles, based on hypotheses about
+the atoms or molecules that make them up \citep[e.g.][]{2009arXiv0906.3544P,
+2014PhRvE..89b2302P, 2015arXiv150303404B}.
 

### 2015-10-09 Paper/ms.tex
Call the scalars L

@@
 we obtain testable information that fixes the expected values of two scalar
-functions of $\xx$, $S_1(\xx)$ and $S_2(\xx)$. It is well known that the
+functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$. It is well known that the
 updated probability distribution that takes into account the constraints is
@@
 \begin{eqnarray}
-p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1S_1(\xx)+\beta_2S_2(\xx)\right]}
+p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1L_1(\xx)+\beta_2L_2(\xx)\right]}
 {Z(\beta_1, \beta_2)}
@@
 constant and by running standard Nested Sampling with ``likelihood''
-$L(\xx) = \exp\left[0.3S_1(\xx) + 0.7S_2(\xx)\right]$. However, usually we
+$L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
 are interested in a range of values for $\beta_1$ and $\beta_2$, and we
@@
 \item The algorithm should seek to explore regions where the values of
-$S_1(\xx)$ and $S_2(\xx)$ are greater than what would be expected from the
+$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
 prior.
@@
 \item The algorithm should be invariant to monotonic transformations of
-$S_1$ and $S_2$, i.e. it should only depend on rankings of $S_1$ and $S_2$
+$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
 values and not the values themselves.
@@
 
-\section{Draft algorithm}
+\section{Proposed algorithm}
 
 The algorithm works as follows.
-Start with $N$ particles drawn from $\pi$.
-Define the {\it corner count} of particle $i$ by
+
+\begin{enumerate}
+\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.\\
+\item Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
+of each particle.\\
+\end{enumerate}
+
+The corner count of particle $i$ is defined by
 \begin{eqnarray}
-C_i &=& \sum_{j=1}^N \mathds{1}\left(S_1^j < S_1^i, S_2^j < S_2^i\right)
+C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
+L_2(\xx_j) < L_2(\xx_i)
+\right).\label{eqn:corner_count}
 \end{eqnarray}

### 2015-10-09 Paper/ms.tex
Working on algorithm

@@
 \begin{enumerate}
-\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.\\
+\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
 \item Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
-of each particle.\\
+of each particle.
+\item Choose one of the particles which has corner count equal to the
+minimum nonzero value.
 \end{enumerate}
@@
 \end{eqnarray}
-
-Need to be able to sort particles into a total order. Since $C_i$ is discrete,
-each particle will also need a tiebreaker variable $a_i$ to accompany it.
-For iteration 1, $\{(C_i, a_i)\}_{i=1}^N$ will provide the total ordering.
-We exclude the corner with the lowest corner count.
-
-For iteration 2, we need a new quantity for total order (I think?), which
 

### 2015-10-09 Code/C++/showresults.py
Print effective number of NS particles thing

@@ -116,17 +116,14 @@ err1 = logZ - true_logZ
 good = logical_not(isnan(err1))
 imshow(err1, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest')
 print("H_max = " + str(H[good].max()))
-
 print(err1[good].min(), err1[good].max())
-print(sqrt(mean((logZ[good] - true_logZ[good])**2)))
+print(4.*H[good].max()/(abs(err1[good]).max())**2)
 
 subplot(1,2,2)
 err2 = (H - true_H)#/true_H
 good = logical_not(isnan(err1))
 imshow(err2, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest')
 print(err2[good].min(), err2[good].max())
-print(sqrt(mean((H[good] - true_H[good])**2/true_H[good]**2)))
-
 show()
 
 

### 2015-10-09 Paper/figures/joint.py
Awesome new figure

@@ -1,0 +1,40 @@
+import numpy as np
+import numpy.random as rng
+from matplotlib import rc
+import matplotlib.pyplot as plt
+
+"""
+Make a plot of \pi(L1, L2)
+Based on a similar plot from ABCNS
+"""
+
+rng.seed(0)
+rc("font", size=18, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+# Resolution
+N = 256
+[x, y] = np.meshgrid(np.linspace(0., 5., N), np.linspace(5., 0., N))
+
+f = np.exp(-0.5*(x-3.5)**2/1.**2)*np.exp(-0.5*((y - 5*(x/5)**2)**2)/0.3**2)
+f /= f.sum()
+
+# Generate samples
+M = 20
+xx = 3.5 + 1.*rng.randn(M)
+yy = 5*(xx/5)**2 + 0.3*rng.randn(M)
+keep = (xx > 0.) & (xx < 5.) & (yy > 0.) & (yy < 5.)
+xx = xx[keep]
+yy = yy[keep]
+
+plt.imshow(f, extent=[x.min(), x.max(), y.min(), y.max()], cmap='Blues')
+plt.plot(xx, yy, 'ko')
+plt.xlabel(r'$L_1$')
+plt.ylabel(r'$L_2$')
+plt.title(r'Prior $\pi(L_1, L_2)$')
+plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
+plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
+plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
+plt.savefig('joint.pdf', bbox_inches='tight')
+plt.show()
+

### 2015-10-09 Code/C++/SamplerImpl.h
exit more elegantly if all particles have lcc=0

@@ -69,6 +69,11 @@ void Sampler<MyModel>::do_iteration()
 			lccs[i] = num_particles;
 	}
 	int min_lcc = *min_element(lccs.begin(), lccs.end());
+	if(min_lcc == num_particles)
+	{
+		std::cerr<<"# All particles have corner counts of zero. Aborting."<<std::endl;
+		exit(1);
+	}
 
 	std::vector<int> usable_indices;
 	for(int i=0; i<num_particles; i++)

### 2015-10-11 Paper/ms.tex
Awesome new figure
Reminder to work out expected accuracy
bits and pieces

@@
 \item Choose one of the particles which has corner count equal to the
-minimum nonzero value.
+minimum nonzero value. If all particles have a corner count of zero,
+the algorithm terminates. Call the selected particle particle $K$.
+\item Choose one of the objective functions, and find $K$'s rank in terms
+of that objective function.
+\item ...
 \end{enumerate}
+
+Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
 
@@
 \end{eqnarray}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/joint.pdf}
+\caption{\it The particle at the corner of the grey rectangle
+has a corner count of 1, since 1 other particle is within the
+grey rectangle.\label{fig:joint}}
+\end{figure}
+
+\section{What kind of accuracy can we expect?}
+Imagine computing $Z(\beta_1, \beta_2)$ by looping over a grid of $\beta_2$
+values and doing an NS run with respect to $L_1$ for each. For each
+value of $\beta_1$ and $\beta_2$ that we're interested in, we can expect
+a (marginal) ``one-sigma'' accuracy in
+$\ln\left[Z(\beta_1, \beta_2)\right]$ of $\pm \sqrt{H(\beta_1, \beta_2)/N}$,
+for a computational cost proportional to...
 

### 2015-10-25 Paper/ms.tex
Some writing

@@
 \section{Introduction}
-
 Nested Sampling \citep[NS][]{skilling} is an effective and popular
@@
 the likelihood $L(\xx)$ that compress the prior mass by an
-approximately-known factor.
+approximately-known factor. Like other related algorithms such as those
+based on ``annealing'', NS
+moves through a sequence of probability distributions beginning from the
+prior. This sequence of distributions is defined by
+\begin{eqnarray}
+p(\xx; \ell) &=& \frac{1}{X(\ell)}\pi(\xx)\mathds{1}\left(L(\xx) > \ell\right).
+\end{eqnarray}
+
 This enables the calculation of the marginal likelihood $Z$
@@
 
-The corner count of particle $i$ is defined by
+The corner count of particle $i$ is defined as the number of particles with
+a lower value of {\em both} objective functions compared to particle $i$.
+Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
+particles in the rectangle whose upper right corner is at
+$\left(L_1(\xx_i), L_2(\xx_i)\right)$. The equation for the corner count is:
 \begin{eqnarray}

### 2015-10-25 Paper/ms.tex
refer to figure

@@
 particles in the rectangle whose upper right corner is at
-$\left(L_1(\xx_i), L_2(\xx_i)\right)$. The equation for the corner count is:
+$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint} for an
+example). The equation for the corner count is:
 \begin{eqnarray}

### 2015-10-26 Paper/ms.tex
Algorithm

@@
 \usepackage[utf8]{inputenc}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
 
@@
 % Full title of the paper (Capitalized)
-\Title{Nested sampling with two or more objective functions}
+\Title{Nested sampling with two objective functions}
 
@@
 of Nested Sampling, such as the ability to cope with first-order phase
-transitions.
+transitions. The algorithm is designed for problems with two objective
+functions, but should generalise to problems with more than two (though perhaps
+not more than a few).
 
@@
 prior.
-\item The algorithm should try to move through a sequence of probability
+\item The algorithm should consider a sequence of probability
 distributions proportional to $\pi$, but restricted to smaller and smaller
@@
 
-The algorithm works as follows.
+Here we describe the proposed algorithm, which took three years of
+experimentation to develop. Please contact the first author for a list of
+failed ideas, some of which might lead to potential alternatives to this
+algorithm.
 
-\begin{enumerate}
-\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
-\item Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
+\begin{algorithm}
+\begin{algorithmic}
+\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
+\While{$abc$}
+	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
 of each particle.
-\item Choose one of the particles which has corner count equal to the
+	\State Choose one of the particles which has corner count equal to the
 minimum nonzero value. If all particles have a corner count of zero,
 the algorithm terminates. Call the selected particle particle $K$.
-\item Choose one of the objective functions, and find $K$'s rank in terms
+	\State Choose one of the objective functions, and find $K$'s rank in terms
 of that objective function.
-\item ...
-\end{enumerate}
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
 

### 2015-10-26 Paper/figures/joint.py
More lines on plot

@@ -34,6 +34,9 @@ plt.ylabel(r'$L_2$')
 plt.title(r'Prior $\pi(L_1, L_2)$')
 plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
 plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
+plt.axhline(1.3095, linestyle='--', color='k')
+plt.axvline(2.6455, linestyle='--', color='k')
+plt.axhline(0.841, xmin=0., xmax=2.6455/5., linestyle='-.', color='k')
 plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
 plt.savefig('joint.pdf', bbox_inches='tight')
 plt.show()

### 2015-10-26 Paper/ms.tex
More lines on plot
random work

@@
 \newcommand{\data}{\boldsymbol{D}}
+\newcommand{\II}{\boldsymbol{I}}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{The abstract goes here.}
+\abstract{We present a generalization of the Nested Sampling algorithm to a
+broad class of problems which arise in statistical mechanics.}
 
@@
 \section{Introduction}
+
 Nested Sampling \citep[NS][]{skilling} is an effective and popular
-Monte Carlo algorithm for Bayesian computation and
-statistical mechanics
+Monte Carlo algorithm which can solve a wide range of problems in Bayesian
+computation and statistical mechanics
 \citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
@@
 posterior distribution for the parameters given data $\data$ (and prior
-information $I$) is:
+information $\II$) is:
 \begin{eqnarray}
-p(\xx | \data, I) &=&
-\frac{p(\xx | I)p(\data | \xx, I)}{p(\data | I)}\\
+p(\xx | \data, \II) &=&
+\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
 &=& \frac{\pi(\xx)L(\xx)}{Z}
@@
 where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
-function, and $Z$ is the normalising constant, known as the
+function, and $Z$ is the normalising constant, usually known as the
 ``marginal likelihood'' or the ``evidence'':
@@
 transitions. The algorithm is designed for problems with two objective
-functions, but should generalise to problems with more than two (though perhaps
+functions, but should generalize to problems with more than two (though perhaps
 not more than a few).
+
+The prior $\pi(\xx)$ implies a certain prior for $L_1$ and $L_2$, which we
+denote $\pi(L_1, L_2)$. An example of a prior is shown as the
+probability density in Figure~\ref{fig:joint}. The partition function
+$Z(\beta_1, \beta_2)$ is a set of expected values with respect to this density.
+Simple Monte Carlo sampling from $\pi$ will not work except for values of
+$(\beta_1, \beta_2)$ where the canonical distribution remains similar to $\pi$.
+We need a sampler that explores regions where $L_1$ and $L_2$ are high in order
+to accurately estimate $Z(\beta_1, \beta_2)$.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/joint.pdf}
+\caption{\it The particle at the corner of the grey rectangle
+has a corner count of 1, since 1 other particle is within the
+grey rectangle.\label{fig:joint}}
+\end{figure}
 
@@
 \State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
-\While{$abc$}
+\While{More iterations needed}
 	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
@@
 \end{eqnarray}
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.5]{figures/joint.pdf}
-\caption{\it The particle at the corner of the grey rectangle
-has a corner count of 1, since 1 other particle is within the
-grey rectangle.\label{fig:joint}}
-\end{figure}
 

### 2015-10-26 Paper/ms.tex
Random work

@@
 denote $\pi(L_1, L_2)$. An example of a prior is shown as the
-probability density in Figure~\ref{fig:joint}. The partition function
+probability density in Figure~\ref{fig:joint1}. The partition function
 $Z(\beta_1, \beta_2)$ is a set of expected values with respect to this density.
@@
 \centering
-\includegraphics[scale=0.5]{figures/joint.pdf}
-\caption{\it The particle at the corner of the grey rectangle
-has a corner count of 1, since 1 other particle is within the
-grey rectangle.\label{fig:joint}}
+\includegraphics[scale=0.5]{figures/joint1.pdf}
+\caption{\it An example of a prior for $L_1$ and $L_2$, which is implied by
+the prior $\pi(\xx)$. We have overplotted some points drawn from this prior.
+\label{fig:joint1}}
 \end{figure}
+
+Our proposed algorithm (introduced in Section~\ref{sec:algorithm}) makes use
+of a quantity we call the corner count. For any values $(\ell_1, \ell_2)$,
+the prior mass for which both objective functions are lower is
+\begin{eqnarray}
+X(\ell_1, \ell_2) &=& \int \pi(L_1, L_2) \mathds{1}\left(L_1 < \ell_1, L_2 < \ell_2 \right)
+ \, dL_1 \, dL_2
+\end{eqnarray}
+This is the two dimensional cumulative distribution function (CDF) of
+$L_1$ and $L_2$.
+
+The corner count of particle $i$ is defined as the number of particles with
+a lower value of {\em both} objective functions compared to particle $i$.
+Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
+particles in the rectangle whose upper right corner is at
+$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
+example). The equation for the corner count is:
+\begin{eqnarray}
+C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
+L_2(\xx_j) < L_2(\xx_i)
+\right).\label{eqn:corner_count}
+\end{eqnarray}
 
@@
 
-\section{Proposed algorithm}
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/joint2.pdf}
+\caption{\it The same as Figure~\ref{fig:joint1}.
+\label{fig:joint2}}
+\end{figure}
+
+\section{Proposed algorithm}\label{sec:algorithm}
 
@@
 
-The corner count of particle $i$ is defined as the number of particles with
-a lower value of {\em both} objective functions compared to particle $i$.
-Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
-particles in the rectangle whose upper right corner is at
-$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint} for an
-example). The equation for the corner count is:
-\begin{eqnarray}
-C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
-L_2(\xx_j) < L_2(\xx_i)
-\right).\label{eqn:corner_count}
-\end{eqnarray}
+\section{What kind of accuracy can we expect?}
 
-\section{What kind of accuracy can we expect?}
 Imagine computing $Z(\beta_1, \beta_2)$ by looping over a grid of $\beta_2$
@@
 for a computational cost proportional to...
+
+\section{What kind of accuracy do we need?}
+
+Use Skilling's argument about the prior distribution of a log Bayes factor.
+Even though the performance of this algorithm won't be great for high $H$, it
+will be good enough. I think the probability it won't give enough accuracy will
+be ``exponentially small''.
 

### 2015-10-26 Paper/figures/joint.py

@@ -32,12 +32,13 @@ plt.plot(xx, yy, 'ko')
 plt.xlabel(r'$L_1$')
 plt.ylabel(r'$L_2$')
 plt.title(r'Prior $\pi(L_1, L_2)$')
+plt.savefig('joint1.pdf', bbox_inches='tight')
 plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
 plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
 plt.axhline(1.3095, linestyle='--', color='k')
 plt.axvline(2.6455, linestyle='--', color='k')
 plt.axhline(0.841, xmin=0., xmax=2.6455/5., linestyle='-.', color='k')
 plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
-plt.savefig('joint.pdf', bbox_inches='tight')
+plt.savefig('joint2.pdf', bbox_inches='tight')
 plt.show()
 

### 2015-10-30 Paper/ms.tex
Words

@@
 $Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
-run. To be considered a variant of Nested Sampling, we require the method
-to satisfy the following requirements:
+run. To be considered a variant of Nested Sampling, the method should satisfy
+the following properties:
 \begin{enumerate}
@@
 \State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
-\While{More iterations needed}
+\While{more iterations desired}
 	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
@@
 \section{What kind of accuracy do we need?}
+
+The following argument is due to John Skilling. Consider Bayesian model selection
+between two models $M_1$ and $M_2$. The Bayes Factor is the ratio of the
+two marginal likelihoods:
+
+\begin{eqnarray}
+B = \frac{P(D | M_1)}{P(D | M_2)} = \frac{Z_1}{Z_2}
+\end{eqnarray}
+In large problems, the log of the Bayes Factor is usually more convenient:
+
+\begin{eqnarray}
+\ln B = \ln Z_1 - \ln Z_2.
+\end{eqnarray}
+
+Naively, we want $\ln B$ to an accuracy of about 0.1, so that we can reliably
+quantify the strength of the evidence when the two models are similarly
+supported by the data.
 

### 2015-10-30 Paper/ms.tex
working on Skilling's argument

@@
 two marginal likelihoods:
-
 \begin{eqnarray}
@@
 In large problems, the log of the Bayes Factor is usually more convenient:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 Naively, we want $\ln B$ to an accuracy of about 0.1, so that we can reliably
 quantify the strength of the evidence when the two models are similarly
-supported by the data.
+supported by the data. But is this a realistic situation?
+Prior to the data, we don't know the
+value of $\ln B$. Rather, $\ln B$ has a prior distribution:
+\begin{eqnarray}
+P(\ln B \in R) &=& \int P(D)P(\ln B \in R | D) \, dD\\
+&=& \int P(D) \mathds{1}\left(\ln B(D) \in R\right) \, dD\\
+&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \in R\right) \, dD
+\end{eqnarray}
 
-Use Skilling's argument about the prior distribution of a log Bayes factor.
-Even though the performance of this algorithm won't be great for high $H$, it
-will be good enough. I think the probability it won't give enough accuracy will
-be ``exponentially small''.
+%Use Skilling's argument about the prior distribution of a log Bayes factor.
+%Even though the performance of this algorithm won't be great for high $H$, it
+%will be good enough. I think the probability it won't give enough accuracy will
+%be ``exponentially small''.
 

### 2015-10-30 Code/C++/truth.py
More publication-ready figure

@@ -1,5 +1,8 @@
 from pylab import *
 
+rc("font", size=18, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
 x = linspace(0., 1., 10001)
 
 def truth(T1, T2, do_plot=False):
@@ -34,17 +37,18 @@ def grid():
 
   figure(1)
   subplot(1, 2, 1)
-  imshow(logZ, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest', cmap='Purples')
-  xlabel('ln($T_1$)')
-  ylabel('ln($T_2$)')
-  title('log(Z)')
+  imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+  xlabel(r'$\log_{10}(T_1)$')
+  ylabel(r'$\log_{10}(T_2)$')
+  title(r'$\ln(Z)$')
 
   subplot(1, 2, 2)
-  imshow(H, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest', cmap='Purples')
-  xlabel('ln($T_1$)')
-  ylabel('ln($T_2$)')
-  title('H, max={m}'.format(m=H.max()))
-
+  imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+  xlabel(r'$\log_{10}(T_1)$')
+#  ylabel(r'$\log_{10}(T_2)$')
+  title(r'$H$')
+  print('H_max = ', H.max())
+  savefig('truth.pdf', bbox_inches='tight')
   show()
   return [logZ, H]
 

### 2015-10-30 Paper/ms.tex
CDF
Include truth figure
Specify the test problem
More words

@@
 
+\section{Test problem}
+
+We tested the algorithm on a simple 100-dimensional example, where the
+prior is a uniform distribution between 0 and 1 for each coordinate:
+\begin{eqnarray}
+\pi(\xx) &=& \prod_{i=1}^{100}
+\left\{
+\begin{array}{lr}
+1, & x_i \in [0, 1]\\
+0, & \textnormal{otherwise.}
+\end{array}
+\right.
+\end{eqnarray}
+The two objective functions are:
+\begin{eqnarray}
+L_1(\xx) &=& -\sum_{i=1}^{100} \left(x_i - 0.5\right)^2\\
+L_2(\xx) &=& -\sum_{i=1}^{100} \log\left[1 +
+\left(\frac{x_i - 0.5}{0.01}\right)^2
+\right].
+\end{eqnarray}
+The canonical distributions at low temperature for $L_1$ and high temperature
+for $L_2$ are gaussians centered at 0.5 with standard deviations
+XX. At high temperatures for $L_2$ and low temperatures for $L_1$ the
+canonical distributions are Cauchy centered at 0.5 with width
+YY.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.75]{figures/truth.pdf}
+\caption{\label{fig:truth}}
+\end{figure}
+
 \section{What kind of accuracy can we expect?}
@@
 Prior to the data, we don't know the
-value of $\ln B$. Rather, $\ln B$ has a prior distribution:
+value of $\ln B$. Rather, $\ln B$ has a prior distribution. We can find this
+prior in terms of its CDF:
 \begin{eqnarray}
-P(\ln B \in R) &=& \int P(D)P(\ln B \in R | D) \, dD\\
-&=& \int P(D) \mathds{1}\left(\ln B(D) \in R\right) \, dD\\
-&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \in R\right) \, dD
+P(\ln B \leq R) &=& \int P(D)P(\ln B \leq R | D) \, dD\\
+&=& \int P(D) \mathds{1}\left(\ln B(D) \leq R\right) \, dD\\
+&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \leq R\right) \, dD
 \end{eqnarray}

### 2015-10-31 Code/C++/showresults.py
Results plot

@@ -2,6 +2,9 @@ from pylab import *
 import scipy
 from postprocess import logsumexp
 
+rc("font", size=18, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
 # Temperatures
 T1, T2 = 0.1, 1.
 
@@ -91,10 +94,11 @@ for i in xrange(0, N):
       exp2[i, j] = NaN
   print(i+1)
 
-figure(figsize=(11, 9))
+figure(figsize=(11, 11))
 subplot(2, 2, 1)
 imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
 title(r'$\log(Z)$')
+ylabel(r'$\log_{10}(T_2)$')
 
 subplot(2, 2, 2)
 imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
@@ -103,11 +107,14 @@ title(r'$H$')
 subplot(2, 2, 3)
 imshow(exp1, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
 title(r'$\left<L_1\right>$')
+xlabel(r'$\log_{10}(T_1)$')
+ylabel(r'$\log_{10}(T_2)$')
 
 subplot(2, 2, 4)
 imshow(exp2, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
 title(r'$\left<L_2\right>$')
-
+xlabel(r'$\log_{10}(T_1)$')
+savefig('results.pdf', bbox_inches='tight')
 show()
 
 true_logZ = loadtxt('true_logZ.txt')

### 2015-10-31 Paper/ms.tex
Results plot
widths

@@
 \end{eqnarray}
-The canonical distributions at low temperature for $L_1$ and high temperature
-for $L_2$ are gaussians centered at 0.5 with standard deviations
-XX. At high temperatures for $L_2$ and low temperatures for $L_1$ the
+The canonical distribution at low temperature for $L_1$ and high temperature
+for $L_2$ is a gaussian centered at 0.5 with standard deviation
+$\sqrt{T_1/2}$. At high temperatures for $L_2$ and low temperatures for $L_1$ the
 canonical distributions are Cauchy centered at 0.5 with width
-YY.
+$0.01\sqrt{T_2}$.
 
@@
 \caption{\label{fig:truth}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/results.pdf}
+\caption{Numerical results for the test problem. These compare well with the
+true partition function and KL divergence given in Figure~\ref{fig:truth}.
+\label{fig:results}}
 \end{figure}

### 2015-10-31 Code/C++/showresults.py
Save other plot

@@ -62,11 +62,16 @@ for i in xrange(0, N):
       break
   posterior_sample[i,:] = scalars[which,:]
 
-plot(posterior_sample[:,0], posterior_sample[:,1], 'b.', markersize=5, alpha=0.5)
+plot(posterior_sample[:,0], posterior_sample[:,1], 'b.', markersize=5,\
+					alpha=0.5, label='Discarded points')
 hold(True)
-plot(scalars[:,0], scalars[:,1], 'r.', markersize=1, alpha=0.2)
-xlabel("Scalar 1")
-ylabel("Scalar 2")
+plot(scalars[:,0], scalars[:,1], 'r.', markersize=1, alpha=0.2,\
+					label='Canonical distribution')
+xlabel("$L_1$")
+ylabel("$L_2$")
+axis([-10, 0, -650, 0])
+#legend(loc='upper left')
+savefig('output.pdf', bbox_inches='tight')
 savetxt("posterior_sample.txt", posterior_sample)
 show()
 

### 2015-10-31 Code/C++/showresults.py
New figure

@@ -71,7 +71,7 @@ xlabel("$L_1$")
 ylabel("$L_2$")
 axis([-10, 0, -650, 0])
 #legend(loc='upper left')
-savefig('output.pdf', bbox_inches='tight')
+savefig('output.png', bbox_inches='tight', dpi=300)
 savetxt("posterior_sample.txt", posterior_sample)
 show()
 

### 2015-10-31 Code/C++/showresults.py
Auto axis limits

@@ -69,7 +69,6 @@ plot(scalars[:,0], scalars[:,1], 'r.', markersize=1, alpha=0.2,\
 					label='Canonical distribution')
 xlabel("$L_1$")
 ylabel("$L_2$")
-axis([-10, 0, -650, 0])
 #legend(loc='upper left')
 savefig('output.png', bbox_inches='tight', dpi=300)
 savetxt("posterior_sample.txt", posterior_sample)

### 2015-11-01 Paper/ms.tex
New figure
sentence

@@
 the algorithm terminates. Call the selected particle particle $K$.
-	\State Choose one of the objective functions, and find $K$'s rank in terms
-of that objective function.
+	\State Choose one of the objective functions 
 \EndWhile
@@
 Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
+In Steps 2 and 3, we search for a particle with a low corner count, such as
+one or two (so the fraction of the remaining prior mass to the lower left of
+the chosen particle is approximately $1/N$ or $2/N$ respectively).
 
@@
 \caption{\label{fig:truth}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.75]{figures/output.png}
+\caption{\label{fig:output}}
 \end{figure}

### 2015-11-02 Paper/ms.tex
Random work

--- 2015-10-31 Paper/ms.tex
@@
 \usepackage{algorithm}
+\usepackage{amsmath}
 \usepackage{algorithmicx}
@@
 
+\DeclareUnicodeCharacter{00A0}{ }
 \newcommand{\xx}{\boldsymbol{x}}
@@
 $\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
-example). The equation for the corner count is:
+example). The equation for the corner count of particle $i$ is:
 \begin{eqnarray}
-C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
+C_i &=& \sum_{\substack{j=1\\{j \neq i}}}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
 L_2(\xx_j) < L_2(\xx_i)
@@
 \item The algorithm should seek to explore regions where the values of
-$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
-prior.
+$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
 \item The algorithm should consider a sequence of probability
@@
 \end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.5]{figures/joint2.pdf}
-\caption{\it The same as Figure~\ref{fig:joint1}.
-\label{fig:joint2}}
-\end{figure}
+The standard Nested Sampling algorithm, (as well as alternatives such as
+Diffusive Nested Sampling \citep{dnest}) has these properties.
 
@@
 \begin{algorithmic}
-\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
+\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
+initialise the {\em remaining prior mass} at $M=1$.
 \While{more iterations desired}
@@
 	\State Choose one of the particles which has corner count equal to the
-minimum nonzero value. If all particles have a corner count of zero,
-the algorithm terminates. Call the selected particle particle $K$.
-	\State Choose one of the objective functions, and find $K$'s rank in terms
-of that objective function.
+minimum nonzero value. Label this particle number $K$.
+	\State Choose one of the objective functions at random. Label this
+objective function $J$. Let $L_{J}(\xx_K)$ be the value of this objective
+function for particle $K$.
 \EndWhile
@@
 Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
+In step 2, we try to find a particle with the lowest possible nonzero
+corner count. Often this will be 1. In Figure~\ref{fig:joint2}, particle $K$
+is located at $(2.65, 1.31)$ and has a corner count of 1.
+
+At step 2, if all particles have a corner count of 0, the algorithm terminates.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/joint2.pdf}
+\caption{\it The same as Figure~\ref{fig:joint1}. The selected particle
+at $(2.65, 1.31)$ has a corner count of 1. The particle at
+$(2.50, 0.84)$ can be thought of as a single-particle standard Nested Sampling
+run restricted to $L_1 < 2.65$ which is attempting to increase $L_1$.
+\label{fig:joint2}}
+\end{figure}
 
@@
 \includegraphics[scale=0.75]{figures/truth.pdf}
-\caption{\label{fig:truth}}
+\caption{The true partition function and KL divergence (from the prior to
+the canonical distribution) as a function of two temperatures for the
+test problem.\label{fig:truth}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.75]{figures/output.png}
+\caption{\label{fig:output}}
 \end{figure}

### 2015-11-02 Code/C++/showresults.py
Comment out first plot

@@ -43,13 +43,13 @@ def canonical_properties(temperature1, temperature2):
 
 [logZ, H, logWW, ess, exp1, exp2] = canonical_properties(T1, T2)
 
-print('log(Z) = {logZ}'.format(logZ=logZ))
-print('H = {H} nats'.format(H=H))
+#print('log(Z) = {logZ}'.format(logZ=logZ))
+#print('H = {H} nats'.format(H=H))
 
-plot(exp(logWW))
-ylabel('Weight wrt canonical distribution')
-title('ESS (for purposes of normalising constant calc) = {ess}'.format(ess=ess))
-show()
+#plot(exp(logWW))
+#ylabel('Weight wrt canonical distribution')
+#title('ESS (for purposes of normalising constant calc) = {ess}'.format(ess=ess))
+#show()
 
 # Resample to uniform weight
 N = int(ess)

### 2015-11-02 Code/C++/Models/Atoms.cpp
2D

@@ -54,13 +54,14 @@ void Atoms::calculate_PE()
 
 void Atoms::calculate_PE(int i, int j)
 {
-	double Rmsq = pow(0.05, 2);
-	double rsq = pow(x[i] - x[j], 2) + pow(y[i] - y[j], 2) + pow(z[i] - z[j], 2);
+	constexpr double Rmsq = pow(0.02, 2);
+	double rsq = pow(x[i] - x[j], 2) + pow(z[i] - z[j], 2);
 	terms[i][j] = pow(Rmsq/rsq, 6) - 2.*pow(Rmsq/rsq, 3);
 }
 
 double Atoms::perturb(RNG& rng)
 {
+	top:
 	int which = rng.rand_int(x.size());
 	int what = rng.rand_int(3);
 
@@ -71,6 +72,7 @@ double Atoms::perturb(RNG& rng)
 	}
 	if(what == 1)
 	{
+		goto top;
 		y[which] += rng.randh();
 		wrap(y[which], 0., 1.);
 	}

### 2015-11-02 Paper/ms.tex
Dnest
Random work

@@
 % Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{We present a generalization of the Nested Sampling algorithm to a
-broad class of problems which arise in statistical mechanics.}
+\abstract{We present a generalization of the Nested Sampling algorithm for
+estimating the partition function of a two-dimensional family of
+probability distributions.}
 
@@
 \end{eqnarray}
-NS begins by drawing particles from the
+The main goal of NS is to approximate $Z$.
+This is done by drawing particles from the
 prior $\pi(\xx)$ and successively imposing constraints on the value of
@@
 
-This enables the calculation of the marginal likelihood $Z$
-and properties of the posterior $\pi(\xx)L(\xx)/Z$.
+In addition to the marginal likelihood $Z$, NS also allows us
+to approximate properties of the posterior $\pi(\xx)L(\xx)/Z$.
 NS is required because Equation~\ref{eqn:evidence}
@@
 connection between NS and ideas from rare event simulation \citep{walter}.
-
 NS also allows us to calculate the properties of any other distribution that
@@
 of the ``canonical'' form:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 If we were only interested in a single canonical distribution, for example
@@
 to accurately estimate $Z(\beta_1, \beta_2)$.
-
 \begin{figure}
@@
 This is the two dimensional cumulative distribution function (CDF) of
-$L_1$ and $L_2$.
-
+$L_1$ and $L_2$. The corner count is a related quantity when we have
+a collection of $N$ particles in the $L_1$-$L_2$ plane.
 The corner count of particle $i$ is defined as the number of particles with
-a lower value of {\em both} objective functions compared to particle $i$.
-Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
-particles in the rectangle whose upper right corner is at
-$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
-example). The equation for the corner count of particle $i$ is:
+a lower value of {\em both} objective functions compared to particle $i$:
 \begin{eqnarray}
@@
 \end{eqnarray}
+Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
+particles in the rectangle whose upper right corner is at
+$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
+example).
 
@@
 
+\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,
+\& Cs{\'a}nyi}{2011}]{dnest} Brewer B.~J., P{\'a}rtay L.~B., Cs{\'a}nyi G., 2011,
+Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380
+
 \bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.

### 2015-11-02 Code/C++/showresults.py
Viridis, error is around zero

@@ -1,6 +1,7 @@
 from pylab import *
 import scipy
 from postprocess import logsumexp
+import colormaps as cmaps
 
 rc("font", size=18, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
@@ -100,22 +101,22 @@ for i in xrange(0, N):
 
 figure(figsize=(11, 11))
 subplot(2, 2, 1)
-imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
 title(r'$\log(Z)$')
 ylabel(r'$\log_{10}(T_2)$')
 
 subplot(2, 2, 2)
-imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
 title(r'$H$')
 
 subplot(2, 2, 3)
-imshow(exp1, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+imshow(exp1, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
 title(r'$\left<L_1\right>$')
 xlabel(r'$\log_{10}(T_1)$')
 ylabel(r'$\log_{10}(T_2)$')
 
 subplot(2, 2, 4)
-imshow(exp2, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+imshow(exp2, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
 title(r'$\left<L_2\right>$')
 xlabel(r'$\log_{10}(T_1)$')
 savefig('results.pdf', bbox_inches='tight')
@@ -126,7 +127,7 @@ true_H = loadtxt('true_H.txt')
 subplot(1,2,1)
 err1 = logZ - true_logZ
[…]

### 2015-11-02 Code/C++/truth.py

@@ -1,4 +1,5 @@
 from pylab import *
+import colormaps as cmaps
 
 rc("font", size=18, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
@@ -37,13 +38,13 @@ def grid():
 
   figure(1)
   subplot(1, 2, 1)
-  imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+  imshow(logZ, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
   xlabel(r'$\log_{10}(T_1)$')
   ylabel(r'$\log_{10}(T_2)$')
   title(r'$\ln(Z)$')
 
   subplot(1, 2, 2)
-  imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap='Purples')
+  imshow(H, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', cmap=cmaps.viridis)
   xlabel(r'$\log_{10}(T_1)$')
 #  ylabel(r'$\log_{10}(T_2)$')
   title(r'$H$')

### 2015-11-02 Paper/ms.tex
Updated joint figure

@@
 is located at $(2.65, 1.31)$ and has a corner count of 1.
-
-At step 2, if all particles have a corner count of 0, the algorithm terminates.
+If all particles have a corner count of 0, the algorithm terminates.
 
@@
 at $(2.65, 1.31)$ has a corner count of 1. The particle at
-$(2.50, 0.84)$ can be thought of as a single-particle standard Nested Sampling
+$(2.53, 0.84)$ can be thought of as a single-particle standard Nested Sampling
 run restricted to $L_1 < 2.65$ which is attempting to increase $L_1$.

### 2015-11-02 Paper/figures/joint.py

@@ -39,6 +39,8 @@ plt.axhline(1.3095, linestyle='--', color='k')
 plt.axvline(2.6455, linestyle='--', color='k')
 plt.axhline(0.841, xmin=0., xmax=2.6455/5., linestyle='-.', color='k')
 plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
+plt.plot(2.645, 1.31, 'r*', markersize=15)
+plt.plot(2.525, 0.84, 'gH', markersize=10)
 plt.savefig('joint2.pdf', bbox_inches='tight')
 plt.show()
 

### 2015-11-03 Code/C++/showresults.py
Better temperature choice to show the canonical dist

@@ -7,7 +7,7 @@ rc("font", size=18, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
 
 # Temperatures
-T1, T2 = 1., 1.
+T1, T2 = 200., 100.
 
 # First calculate things about the scalars (e.g. the normalising constant)
 output = loadtxt('sample_info.txt')

### 2015-11-03 Paper/ms.tex
Better temperature choice to show the canonical dist
Updated scalars

@@
 \begin{eqnarray}
-L_1(\xx) &=& -\sum_{i=1}^{100} \left(x_i - 0.5\right)^2\\
+L_1(\xx) &=& -\frac{1}{2}\sum_{i=1}^{100} \left(\frac{x_i - 0.5}{0.01}\right)^2\\
 L_2(\xx) &=& -\sum_{i=1}^{100} \log\left[1 +
@@
 for $L_2$ is a gaussian centered at 0.5 with standard deviation
-$\sqrt{T_1/2}$. At high temperatures for $L_2$ and low temperatures for $L_1$ the
+$0.01\sqrt{T_1}$. At high temperatures for $L_2$ and low temperatures for $L_1$ the
 canonical distributions are Cauchy centered at 0.5 with width
@@
 \includegraphics[scale=0.75]{figures/output.png}
-\caption{\label{fig:output}}
+\caption{The sequence of discarded particles (yellow) and
+points representing the canonical distribution at temperatures
+$(T_1, T_2) = (200, 100)$ (black).\label{fig:output}}
 \end{figure}

### 2015-11-04 Paper/ms.tex
Fixed error

@@
 $(2.53, 0.84)$ can be thought of as a single-particle standard Nested Sampling
-run restricted to $L_1 < 2.65$ which is attempting to increase $L_1$.
+run restricted to $L_1 < 2.65$ which is attempting to increase $L_2$.
 \label{fig:joint2}}

### 2015-11-04 Paper/ms.tex
tweaks

@@
 function, and $Z$ is the normalising constant, usually known as the
-``marginal likelihood'' or the ``evidence'':
+`marginal likelihood' or the `evidence':
 \begin{eqnarray}
@@
 approximately-known factor. Like other related algorithms such as those
-based on ``annealing'', NS
+based on `annealing', NS
 moves through a sequence of probability distributions beginning from the
@@
 between the prior and the posterior. For example, we might be interested in
-a ``power posterior'' where the likelihood is raised to a power $\beta$:
+a `power posterior' where the likelihood is raised to a power $\beta$:
 \begin{eqnarray}
@@
 An example application, in the case of
-a Bayesian inference problem with a ``gaussian noise'' likelihood assumption,
+a Bayesian inference problem with a `gaussian noise' likelihood assumption,
 computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
@@
 
-Alternatives to NS include methods based on ``annealing'', where a sequence of
+Alternatives to NS include methods based on `annealing', where a sequence of
 distributions of the form of Equation~\ref{eqn:power_posterior} is used.
@@
 updated probability distribution that takes into account the constraints is
-of the ``canonical'' form:
+of the `canonical' form:
 \begin{eqnarray}
@@
 \end{eqnarray}
+where $\beta_1$ and $\beta_2$ are two `inverse temperatures'.
 If we were only interested in a single canonical distribution, for example
 with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
-constant and by running standard Nested Sampling with ``likelihood''
+constant and by running standard Nested Sampling with `likelihood'
 $L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
@@
 
+The true partition function (computed numerically) and the Kullback-Leibler
+divergence from the prior to the canonical distribution are shown in
+Figure~\ref{fig:truth} as a function of two temperatures.
+
 \begin{figure}
@@
 value of $\beta_1$ and $\beta_2$ that we're interested in, we can expect
-a (marginal) ``one-sigma'' accuracy in
+a (marginal) `one-sigma' accuracy in
 $\ln\left[Z(\beta_1, \beta_2)\right]$ of $\pm \sqrt{H(\beta_1, \beta_2)/N}$,
@@
 \end{eqnarray}
+where $Z_1 = \int \pi(\xx_1)L(\xx_1) \, d\xx_1$ and
+$Z_2 = \int \pi(\xx_2)L(\xx_2) \, d\xx_2$.
 In large problems, the log of the Bayes Factor is usually more convenient:
@@
 value of $\ln B$. Rather, $\ln B$ has a prior distribution. We can find this
-prior in terms of its CDF:
-\begin{eqnarray}
-P(\ln B \leq R) &=& \int P(D)P(\ln B \leq R | D) \, dD\\
-&=& \int P(D) \mathds{1}\left(\ln B(D) \leq R\right) \, dD\\
-&=& \int \sum_{i\in\{1,2\}}\left[P(M_i)P(D|M_i)\right] \mathds{1}\left(\ln B(D) \leq R\right) \, dD
-\end{eqnarray}
+prior via its CDF:
+
+{\bf Need thought here}
+{\bf Thermodynamic integral might be helpful?}
 

### 2015-11-05 Paper/ms.tex
thermodynamic integral, for inspiration
Some of Ewan's comments

@@
 \abstract{We present a generalization of the Nested Sampling algorithm for
-estimating the partition function of a two-dimensional family of
-probability distributions.}
+estimating the partition function of a family of
+probability distributions indexed by two hyperparameters.}
 
@@
 \citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
-In a Bayesian inference problem with unknown parameters denoted collectively
+In a Bayesian inference framework with unknown parameters denoted collectively
 by a vector $\xx$, the
@@
 \end{eqnarray}
-
+and the marginal likelihood can be computed by numerically integrating the
+function $L(X)$:
+\begin{eqnarray}
+Z &=& \int_0^1 L(X)\, dX.
+\end{eqnarray}
 In addition to the marginal likelihood $Z$, NS also allows us
@@
 connection between NS and ideas from rare event simulation \citep{walter}.
-NS also allows us to calculate the properties of any other distribution that
+Using a single NS run, we can also calculate the properties of any other distribution that
 is in some sense intermediate
@@
 \end{eqnarray}
-where $E(\xx)$ is the energy function, and $\pi(\xx)$ is often uniform over
+where $E(\xx)$ is the energy function (analogous to minus the log likelihood
+in the Bayesian case), and $\pi(\xx)$ is often uniform over
 phase space (the set of possible positions and momenta of a collection of
@@
 the atoms or molecules that make them up \citep[e.g.][]{2009arXiv0906.3544P,
-2014PhRvE..89b2302P, 2015arXiv150303404B}.
+2014PhRvE..89b2302P, 2015arXiv150303404B}. The NS algorithm is invariant under
+monotonic transformations of $L$ or $E$ and therefore there we can discuss the
+algorithm in Bayesian or statistical mechanical terms without loss of
+generality.
 
@@
 we obtain testable information that fixes the expected values of two scalar
-functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$. It is well known that the
+functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$.
+It is well known that the
 updated probability distribution that takes into account the constraints is
@@
 \end{eqnarray}
-where $\beta_1$ and $\beta_2$ are two `inverse temperatures'.
+where $\beta_1$ and $\beta_2$ are two `inverse temperatures'. In a Bayesian
+context, this distribution (when $\beta_1 = \beta_2 = 1$) could be the
+posterior distribution for parameters $\xx$
+given two datasets, where $L_1$ and $L_2$ are the
+likelihood functions for each dataset. In statistical mechanics, $L_1$ might
+be $\exp\left(-\textnormal{energy}\right)$ and $L_2$ might be the total
+angular momentum.
 If we were only interested in a single canonical distribution, for example
@@
 Prior to the data, we don't know the
-value of $\ln B$. Rather, $\ln B$ has a prior distribution. We can find this
-prior via its CDF:
+value of $B$. Rather, $B$ has a prior distribution:
+\begin{eqnarray}
+p(B) &=& P(M_1)p(B|M_1) + P(M_2)p(B|M_2)\\
+&=& P(M_1)\int p(\xx_1, D, B | M_1)\,d\xx_1\,dD
+    + P(M_2)\int p(\xx_2, D, B | M_2)\,d\xx_2\,dD
+\end{eqnarray}
 
@@
 {\bf Thermodynamic integral might be helpful?}
+\begin{eqnarray}
+\ln\left[\int p(\theta)p(D|\theta) \, d\theta\right]
+&=& \int_0^1 \left<\ln p(D|\theta) \right>_\beta \, d\beta
+\end{eqnarray}
 

### 2015-11-05 Paper/ms.tex
Define objective function

@@
 functions, but should generalize to problems with more than two (though perhaps
-not more than a few).
+not more than a few). Throughout this paper we will refer $L_i(\xx)$ or
+$\exp\left[-E_i(\xx)\right]$ as `objective functions',
+by analogy with the terminology used when
+discussing optimizations methods. Although we are interested in sampling rather
+than optimization, we still want to increase the values of $L_i$ or decrease
+the values of $E_i$ relative to what is typical of the prior.
 

### 2015-11-05 Code/C++/display_atoms.py
General number of atoms

@@ -2,12 +2,13 @@
 from pylab import *
 
 output = atleast_2d(loadtxt('sample.txt'))
+num_atoms = (output.shape[1] - 3)//3
 
 ion()
 hold(False)
-for i in xrange(0, output.shape[0]):
+for i in range(0, output.shape[0]):
 	model = output[i, 3:]
-	x, y = model[0:50], model[100:150]
+	x, y = model[0:num_atoms], model[2*num_atoms:3*num_atoms]
 	plot(x, y, 'k.')
 	axis('scaled')
 	axis([0, 1, 0, 1])

### 2015-11-05 Code/C++/showresults.py
xrange --> range for Python3

@@ -56,7 +56,7 @@ show()
 N = int(ess)
 posterior_sample = zeros((N, scalars.shape[1]))
 w = exp(logWW)/max(exp(logWW))
-for i in xrange(0, N):
+for i in range(0, N):
   while True:
     which = randint(scalars.shape[0])
     if rand() <= w[which]:
@@ -88,8 +88,8 @@ exp1 = zeros((N, N))
 exp2 = zeros((N, N))
 depth = output[:,0].max() - output[:,0].min()
 
-for i in xrange(0, N):
-  for j in xrange(0, N):
+for i in range(0, N):
+  for j in range(0, N):
     [logZ[i,j], H[i,j], temp1, temp2, exp1[i, j], exp2[i, j]] = canonical_properties(T1[i, j], T2[i, j])
     # Blank out 'unreliable' results
     if H[i, j] > 0.7*depth:

### 2015-11-05 Code/C++/display_potts.py
xrange --> range

@@ -5,7 +5,7 @@ output = atleast_2d(loadtxt('sample.txt'))
 
 ion()
 hold(False)
-for i in xrange(0, output.shape[0]):
+for i in range(0, output.shape[0]):
 	x = output[i, 3:].reshape((50, 50))
 	imshow(x, interpolation='nearest', cmap='jet')
 	title(str(i+1) + '/' + str(output.shape[0]))

### 2015-11-05 Code/C++/showresults.py
Adapt axes

@@ -67,9 +67,13 @@ plot(scalars[0::10,0], scalars[0::10,1], 'y.', markersize=1, label='Discarded po
 hold(True)
 plot(posterior_sample[:,0], posterior_sample[:,1], 'k.', markersize=1,\
 					alpha=1, label='Canonical distribution')
+# Adaptive axes
+ss1 = sort(scalars[:,0])
+ss2 = sort(scalars[:,1])
+xlim([ss1[int(0.05*len(ss1))], ss1.max()])
+ylim([ss2[int(0.05*len(ss2))], ss2.max()])
 xlabel("$L_1$")
 ylabel("$L_2$")
-#xlim([-10, 0])
 legend(loc='upper left', markerscale=10, numpoints=1)
 savefig('output.png', bbox_inches='tight', dpi=300)
 savetxt("posterior_sample.txt", posterior_sample)

### 2015-11-22 Paper/ms.tex
Upper corner count

@@
 Our proposed algorithm (introduced in Section~\ref{sec:algorithm}) makes use
-of a quantity we call the corner count. For any values $(\ell_1, \ell_2)$,
-the prior mass for which both objective functions are lower is
+of the (two dimensional) cumulative distribution of the objective functions.
+For any values $(\ell_1, \ell_2)$ of the objective functions,
+the prior mass for which both $L_1 > \ell_1$ {\em and} $L_2 > \ell_2$ is:
 \begin{eqnarray}
-X(\ell_1, \ell_2) &=& \int \pi(L_1, L_2) \mathds{1}\left(L_1 < \ell_1, L_2 < \ell_2 \right)
+X(\ell_1, \ell_2) &=& \int \pi(L_1, L_2)
+\mathds{1}\left[L_1 > \ell_1, L_2 > \ell_2 \right]
  \, dL_1 \, dL_2
 \end{eqnarray}
-This is the two dimensional cumulative distribution function (CDF) of
-$L_1$ and $L_2$. The corner count is a related quantity when we have
-a collection of $N$ particles in the $L_1$-$L_2$ plane.
-The corner count of particle $i$ is defined as the number of particles with
-a lower value of {\em both} objective functions compared to particle $i$:
+When the prior $\pi$ is approximated with a set of $N$
+particles $\{\xx_i\}_{i=1}^N$
+(Monte Carlo samples ``drawn from'' $\pi$),
+$X(\ell_1, \ell_2)$ can be approximated by:
 \begin{eqnarray}
-C_i &=& \sum_{\substack{j=1\\{j \neq i}}}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
-L_2(\xx_j) < L_2(\xx_i)
-\right).\label{eqn:corner_count}
+\hat{X}(\ell_1, \ell_2) &=&
+\frac{1}{N}
+\sum_{i=1}^N \mathds{1}\left[L_1(\xx_i) > \ell_1,
+L_2(\xx_i) > \ell_2\right].\label{eqn:corner_count}
 \end{eqnarray}
-Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
-particles in the rectangle whose upper right corner is at
-$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
+Graphically, on a plot of $L_2$ vs. $L_1$, $\hat{X}(\ell_1, \ell_2)$
+is the fraction of particles in the rectangle whose upper right corner is at
+$\left(\ell_1, \ell_2\right)$ (see Figure~\ref{fig:joint2} for an
 example).

### 2015-11-23 Code/C++/main.cpp
Vector of RNGs

@@ -10,24 +10,30 @@
 
 using namespace std;
 
-int main()
+int main(int argc, char** argv)
 {
-	// Make an RNG
-	RNG rng;
-	rng.set_seed(time(0));
+	// Number of threads is 1 or a number given on the command line
+	int num_threads = (argc==1)?(1):(atoi(argv[1]));
 
-	constexpr int num_particles = 3001;
-	constexpr int num_mcmc_steps = 1000;
-	constexpr double depth = 1000.;
-	constexpr int steps = depth/log(2.);
+	// Make (num_threads) RNGs: one for the main process
+	// and num_threads for use whenever we send out threads
+	vector<RNG> rngs(num_threads+1);
+	auto seed = time(0);
+	for(RNG& r: rngs)
+		r.set_seed(++seed);
 
-	// Create a sampler
-	Sampler<SimpleExample> sampler(rng, num_particles, num_mcmc_steps, 1);
-	sampler.initialise();
+//	constexpr int num_particles = 3001;
+//	constexpr int num_mcmc_steps = 1000;
+//	constexpr double depth = 1000.;
+//	constexpr int steps = depth/log(2.);
 
-	// Do NS indefinitely
-	for(int i=0; i<steps; ++i)
-		sampler.do_iteration();
+//	// Create a sampler
+//	Sampler<SimpleExample> sampler(rng, num_particles, num_mcmc_steps, 1);
+//	sampler.initialise();
+
+//	// Do NS indefinitely
+//	for(int i=0; i<steps; ++i)
+//		sampler.do_iteration();
 
 	return 0;
 }

### 2015-11-23 Code/C++/Sampler.h
Use rngs[0]

@@ -16,7 +16,7 @@ class Sampler
 {
 	private:
 		// The random number generator to use
-		RNG rng;
+		std::vector<RNG> rngs;
 
 		// The particles, and information about them
 		int num_particles;
@@ -60,11 +60,8 @@ class Sampler
 
 	public:
 		// Constructor
-		Sampler(const RNG& rng, int num_particles, int mcmc_steps,
+		Sampler(const std::vector<RNG>& rngs, int num_particles, int mcmc_steps,
 														int saves_per_iteration);
-
-		// Set RNG seed
-		void set_rng_seed(unsigned int seed);
 
 		// Call from_prior on all the particles
 		void initialise();
@@ -82,12 +79,6 @@ class Sampler
 		// the output files. If not called, it will append
 		// to whatever's already in them
 		void clear_output_files();
-
-		// Setter and getter for the RNG
-		void set_rng(const RNG& rng)
-		{ this->rng = rng; }
-		RNG get_rng() const
-		{ return rng; }
 };
 
 #include "SamplerImpl.h"

### 2015-11-23 Code/C++/SamplerImpl.h

@@ -8,9 +8,9 @@
 #include "Utils.h"
 
 template<class MyModel>
-Sampler<MyModel>::Sampler(const RNG& rng, int num_particles, int mcmc_steps,
-							int saves_per_iteration)
-:rng(rng)
+Sampler<MyModel>::Sampler(const std::vector<RNG>& rngs, int num_particles,
+							int mcmc_steps,	int saves_per_iteration)
+:rngs(rngs)
 ,num_particles(num_particles)
 ,particles(num_particles)
 ,scalars(num_particles)
@@ -31,23 +31,17 @@ Sampler<MyModel>::Sampler(const RNG& rng, int num_particles, int mcmc_steps,
 }
 
 template<class MyModel>
-void Sampler<MyModel>::set_rng_seed(unsigned int seed)
-{
-	rng.set_seed(seed);
-}
-
-template<class MyModel>
 void Sampler<MyModel>::initialise()
 {
 	for(int i=0; i<num_particles; i++)
 	{
-		particles[i].from_prior(rng);
+		particles[i].from_prior(rngs[0]);
 		const std::vector<double>& s = particles[i].get_scalars();
 		scalars[i].clear();
 		for(size_t j=0; j<s.size(); j++)
 			scalars[i].push_back(ScalarType(s[j]));
 		for(size_t j=0; j<scalars[i].size(); j++)
-			scalars[i][j].from_prior(rng);
+			scalars[i][j].from_prior(rngs[0]);
 	}
 }
 
@@ -148,7 +142,7 @@ void Sampler<MyModel>::do_iteration()
 			int ii;
 			do
 			{
-				ii = rng.rand_int(num_particles);
+				ii = rngs[0].rand_int(num_particles);
 			}while(!dying[ii]);
 			save[ii] = true;
 		}
@@ -227,7 +221,7 @@ int Sampler<MyModel>::refresh_particle(int which)
 	int copy;
 	do
 	{
-		copy = rng.rand_int(num_particles);
+		copy = rngs[0].rand_int(num_particles);
 	}
 	while(!is_okay(scalars[copy]));
 
@@ -245,12 +239,12 @@ int Sampler<MyModel>::refresh_particle(int which)
 		proposal = particles[which];
 		s_proposal = scalars[which];
 
-		logH = proposal.perturb(rng);
+		logH = proposal.perturb(rngs[0]);
 		for(size_t j=0; j<s_proposal.size(); j++)
[…]

### 2015-11-23 Code/C++/main.cpp

@@ -22,18 +22,18 @@ int main(int argc, char** argv)
 	for(RNG& r: rngs)
 		r.set_seed(++seed);
 
-//	constexpr int num_particles = 3001;
-//	constexpr int num_mcmc_steps = 1000;
-//	constexpr double depth = 1000.;
-//	constexpr int steps = depth/log(2.);
+	constexpr int num_particles = 3001;
+	constexpr int num_mcmc_steps = 1000;
+	constexpr double depth = 1000.;
+	constexpr int steps = depth/log(2.);
 
-//	// Create a sampler
-//	Sampler<SimpleExample> sampler(rng, num_particles, num_mcmc_steps, 1);
-//	sampler.initialise();
+	// Create a sampler
+	Sampler<SimpleExample> sampler(rngs, num_particles, num_mcmc_steps, 1);
+	sampler.initialise();
 
-//	// Do NS indefinitely
-//	for(int i=0; i<steps; ++i)
-//		sampler.do_iteration();
+	// Do NS indefinitely
+	for(int i=0; i<steps; ++i)
+		sampler.do_iteration();
 
 	return 0;
 }

### 2015-11-23 Code/C++/Sampler.h
Writing the method needed for multi-threading

@@ -44,7 +44,9 @@ class Sampler
 		// Do MCMC to equilibrate a particle
 		int refresh_particle(int which);
 
-		// Check s against rects
+		// Do MCMC to equilibrate a set of particles
+		int refresh_particles(const std::vector<int>& indices);
+
 	public:
 		bool is_okay(const std::vector<ScalarType>& s);
 

### 2015-11-23 Code/C++/SamplerImpl.h

@@ -202,16 +202,33 @@ void Sampler<MyModel>::do_iteration()
 
 	// Replace dead particles
 	int accepted = 0;
+	int num_threads = rngs.size() - 1;
+
+	// Assign particles to threads
+	std::vector< std::vector<int> > which_particles(num_threads);
+	int k=0;
 	for(int i=0; i<num_particles; ++i)
 	{
 		if(dying[i])
-		{
-			// Do MCMC to generate a new particle
-			accepted += refresh_particle(i);
-		}
+			which_particles[k++].push_back(i);
 	}
+
+	// Do MCMC to generate a new particle
+	// accepted += refresh_particle(i);
+
 	std::cout<<"# Accepted "<<accepted<<"/"<<num_dying*mcmc_steps<<". "<<std::endl<<std::endl;
 	iteration++;
+}
+
+
+// Refresh all the particles in indices
+template<class MyModel>
+int Sampler<MyModel>::refresh_particles(const std::vector<int>& indices)
+{
+	int accepts = 0;
+	for(int i: indices)
+		accepts += refresh_particle(i);
+	return accepts;
 }
 
 template<class MyModel>

### 2015-11-23 Code/C++/Sampler.h
Seems to be working!

@@ -42,10 +42,11 @@ class Sampler
 		long double log_prior_mass;
 
 		// Do MCMC to equilibrate a particle
-		int refresh_particle(int which);
+		int refresh_particle(int which, int which_rng);
 
 		// Do MCMC to equilibrate a set of particles
-		int refresh_particles(const std::vector<int>& indices);
+		void refresh_particles(const std::vector<int>& indices, int which_rng,
+										int& accepts);
 
 	public:
 		bool is_okay(const std::vector<ScalarType>& s);

### 2015-11-23 Code/C++/SamplerImpl.h

@@ -5,6 +5,7 @@
 #include <fstream>
 #include <algorithm>
 #include <cassert>
+#include <thread>
 #include "Utils.h"
 
 template<class MyModel>
@@ -201,8 +202,7 @@ void Sampler<MyModel>::do_iteration()
 	std::cout<<log_prior_mass<<"."<<std::endl;
 
 	// Replace dead particles
-	int accepted = 0;
-	int num_threads = rngs.size() - 1;
+	int num_threads = rngs.size();
 
 	// Assign particles to threads
 	std::vector< std::vector<int> > which_particles(num_threads);
@@ -210,11 +210,23 @@ void Sampler<MyModel>::do_iteration()
 	for(int i=0; i<num_particles; ++i)
 	{
 		if(dying[i])
-			which_particles[k++].push_back(i);
+			which_particles[(k++)%num_threads].push_back(i);
 	}
+	// Store acceptance counts
+	std::vector<int> accepts(num_threads);
 
-	// Do MCMC to generate a new particle
-	// accepted += refresh_particle(i);
+	// Do MCMC to generate new particles
+	std::vector<std::thread> threads;
+	for(int i=0; i<num_threads; i++)
+	{
+		threads.push_back(std::thread(std::bind(&Sampler<MyModel>::refresh_particles, this, which_particles[i], i, std::ref(accepts[i]))));
+	}
+	for(int i=0; i<num_threads; i++)
+		threads[i].join();
+	// Sum acceptance counts
+	int accepted = 0;
+	for(const int& c: accepts)
+		accepted += c;
 
 	std::cout<<"# Accepted "<<accepted<<"/"<<num_dying*mcmc_steps<<". "<<std::endl<<std::endl;
 	iteration++;
@@ -223,22 +235,23 @@ void Sampler<MyModel>::do_iteration()
 
 // Refresh all the particles in indices
 template<class MyModel>
-int Sampler<MyModel>::refresh_particles(const std::vector<int>& indices)
+void Sampler<MyModel>::refresh_particles(const std::vector<int>& indices,
+														int which_rng,
+														int& accepts)
 {
-	int accepts = 0;
+	accepts = 0;
 	for(int i: indices)
-		accepts += refresh_particle(i);
-	return accepts;
+		accepts += refresh_particle(i, which_rng);
 }
 
 template<class MyModel>
-int Sampler<MyModel>::refresh_particle(int which)
[…]

### 2015-11-23 Code/C++/main.cpp

@@ -17,7 +17,7 @@ int main(int argc, char** argv)
 
 	// Make (num_threads) RNGs: one for the main process
 	// and num_threads for use whenever we send out threads
-	vector<RNG> rngs(num_threads+1);
+	vector<RNG> rngs(num_threads, RNG());
 	auto seed = time(0);
 	for(RNG& r: rngs)
 		r.set_seed(++seed);

### 2015-11-23 Paper/figures/joint.py
New version of Figure

@@ -32,15 +32,20 @@ plt.plot(xx, yy, 'ko')
 plt.xlabel(r'$L_1$')
 plt.ylabel(r'$L_2$')
 plt.title(r'Prior $\pi(L_1, L_2)$')
-plt.savefig('joint1.pdf', bbox_inches='tight')
-plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
-plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
-plt.axhline(1.3095, linestyle='--', color='k')
-plt.axvline(2.6455, linestyle='--', color='k')
-plt.axhline(0.841, xmin=0., xmax=2.6455/5., linestyle='-.', color='k')
-plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
-plt.plot(2.645, 1.31, 'r*', markersize=15)
-plt.plot(2.525, 0.84, 'gH', markersize=10)
-plt.savefig('joint2.pdf', bbox_inches='tight')
+plt.axhline(1.5, xmin=3./5., xmax=5./5., color='k')
+plt.axvline(3., ymin=1.5/5., ymax=5./5., color='k')
+plt.fill_between(x[0, :][x[0, :] > 3.], 1.5, 5., color=[0.6, 0.6, 0.6], alpha=0.2)
 plt.show()
+
+#plt.savefig('joint1.pdf', bbox_inches='tight')
+#plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
+#plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
+#plt.axhline(1.3095, linestyle='--', color='k')
+#plt.axvline(2.6455, linestyle='--', color='k')
+#plt.axhline(0.841, xmin=0., xmax=2.6455/5., linestyle='-.', color='k')
+#plt.fill_between(x[0, :][x[0, :] < 2.6455], 0., 1.3095, color=[0.6, 0.6, 0.6], alpha=0.2)
+#plt.plot(2.645, 1.31, 'r*', markersize=15)
+#plt.plot(2.525, 0.84, 'gH', markersize=10)
+#plt.savefig('joint2.pdf', bbox_inches='tight')
+#plt.show()
 

### 2015-11-23 Paper/figures/joint.py
Save the figure

@@ -35,9 +35,9 @@ plt.title(r'Prior $\pi(L_1, L_2)$')
 plt.axhline(1.5, xmin=3./5., xmax=5./5., color='k')
 plt.axvline(3., ymin=1.5/5., ymax=5./5., color='k')
 plt.fill_between(x[0, :][x[0, :] > 3.], 1.5, 5., color=[0.6, 0.6, 0.6], alpha=0.2)
+plt.savefig('joint1.pdf', bbox_inches='tight')
 plt.show()
 
-#plt.savefig('joint1.pdf', bbox_inches='tight')
 #plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
 #plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
 #plt.axhline(1.3095, linestyle='--', color='k')

### 2015-11-23 Paper/ms.tex
Updated caption
refer to correct figure

@@
 \caption{\it An example of a prior for $L_1$ and $L_2$, which is implied by
-the prior $\pi(\xx)$. We have overplotted some points drawn from this prior.
+the prior $\pi(\xx)$. We have overplotted 15 points drawn from this prior.
+The estimated prior mass in the shaded rectangle is
+$\hat{X}(3, 1.5) = 13/15 \approx 0.867$.
 \label{fig:joint1}}
@@
 is the fraction of particles in the rectangle whose upper right corner is at
-$\left(\ell_1, \ell_2\right)$ (see Figure~\ref{fig:joint2} for an
+$\left(\ell_1, \ell_2\right)$ (see Figure~\ref{fig:joint1} for an
 example).

### 2015-11-23 Paper/figures/joint.py
updated figures

@@ -8,7 +8,7 @@ Make a plot of \pi(L1, L2)
 Based on a similar plot from ABCNS
 """
 
-rng.seed(0)
+rng.seed(3)
 rc("font", size=18, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
 
@@ -16,13 +16,13 @@ rc("text", usetex=True)
 N = 256
 [x, y] = np.meshgrid(np.linspace(0., 5., N), np.linspace(5., 0., N))
 
-f = np.exp(-0.5*(x-3.5)**2/1.**2)*np.exp(-0.5*((y - 5*(x/5)**2)**2)/0.3**2)
+f = np.exp(-0.5*(x-1.5)**2/1.**2)*np.exp(-0.5*((y - 5*(x/5)**2)**2)/1.**2)
 f /= f.sum()
 
 # Generate samples
-M = 20
-xx = 3.5 + 1.*rng.randn(M)
-yy = 5*(xx/5)**2 + 0.3*rng.randn(M)
+M = 40
+xx = 1.5 + 1.*rng.randn(M)
+yy = 5*(xx/5)**2 + 1.*rng.randn(M)
 keep = (xx > 0.) & (xx < 5.) & (yy > 0.) & (yy < 5.)
 xx = xx[keep]
 yy = yy[keep]
@@ -32,20 +32,26 @@ plt.plot(xx, yy, 'ko')
 plt.xlabel(r'$L_1$')
 plt.ylabel(r'$L_2$')
 plt.title(r'Prior $\pi(L_1, L_2)$')
-plt.axhline(1.5, xmin=3./5., xmax=5./5., color='k')
-plt.axvline(3., ymin=1.5/5., ymax=5./5., color='k')
-plt.fill_between(x[0, :][x[0, :] > 3.], 1.5, 5., color=[0.6, 0.6, 0.6], alpha=0.2)
+plt.fill_between(x[0, :][x[0, :] > 2.], 1.5, 5., color=[0, 0, 0], alpha=0.1)
 plt.savefig('joint1.pdf', bbox_inches='tight')
 plt.show()
 
-#plt.axhline(1.3095, xmin=0., xmax=2.6455/5., color='k')
-#plt.axvline(2.6455, ymin=0., ymax=1.3095/5., color='k')
-#plt.axhline(1.3095, linestyle='--', color='k')
-#plt.axvline(2.6455, linestyle='--', color='k')
[…]

### 2015-11-23 Paper/figures/joint.py
fixed bug

@@ -51,7 +51,6 @@ plt.plot(xx, yy, 'ko')
 plt.xlabel(r'$L_1$')
 plt.ylabel(r'$L_2$')
 plt.title(r'Prior $\pi(L_1, L_2)$')
-#plt.fill_between(x[0, :][x[0, :] > 3.], 1.5, 5., color=[0, 0, 0], alpha=0.1)
-plt.savefig('joint1.pdf', bbox_inches='tight')
+plt.savefig('joint2.pdf', bbox_inches='tight')
 plt.show()
 

### 2015-11-23 Paper/ms.tex
captions up to date

@@
 \caption{\it An example of a prior for $L_1$ and $L_2$, which is implied by
-the prior $\pi(\xx)$. We have overplotted 15 points drawn from this prior.
+the prior $\pi(\xx)$. We have overplotted 25 points drawn from this prior.
 The estimated prior mass in the shaded rectangle is
-$\hat{X}(3, 1.5) = 13/15 \approx 0.867$.
+$\hat{X}(2, 1.5) = 5/25 = 0.2$.
 \label{fig:joint1}}
@@
 \includegraphics[scale=0.5]{figures/joint2.pdf}
-\caption{\it The same as Figure~\ref{fig:joint1}. The selected particle
-at $(2.65, 1.31)$ has a corner count of 1. The particle at
-$(2.53, 0.84)$ can be thought of as a single-particle standard Nested Sampling
-run restricted to $L_1 < 2.65$ which is attempting to increase $L_2$.
+\caption{\it The joint prior for the objective functions, and Monte Carlo
+samples, as in Figure~\ref{fig:joint1}. The shaded region imposes a
+constraint on the value of $\hat{X}$. Its boundary approximates a contour
+of $X(L_1, L_2)$.
 \label{fig:joint2}}

### 2015-11-23 Paper/ms.tex
algorithm

@@
 distributions proportional to $\pi$, but restricted to smaller and smaller
-domains for which we can approximately measure the enclosed prior mass.
+domains for which the enclosed prior mass shrinks by a (roughly) constant and
+known factor each iteration.
 \item The algorithm should be invariant to monotonic transformations of
@@
 \State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
-initialise the {\em remaining prior mass} at $M=1$.
+initialise the {\em remaining prior mass} at $M=1$. The number of particles
+$N$ should be odd and quite large (1001 or greater).
 \While{more iterations desired}
-	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
-of each particle.
-	\State Choose one of the particles which has corner count equal to the
-minimum nonzero value. Label this particle number $K$.
-	\State Choose one of the objective functions at random. Label this
-objective function $J$. Let $L_{J}(\xx_K)$ be the value of this objective
-function for particle $K$.
+	\State Calculate $\hat{X}(\ell_1, \ell_2)$ at the position of each particle
+			and sort the $\hat{X}$ values observed.
+	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$, where $k = (N-1)/2$.
+	\While{$k > 1$ and $\hat{X}_{(k-1)} = \hat{X}_{(k)}$}
+		\State Set $k \leftarrow (k-1)$.
+	\EndWhile
+	\State Save all particles with $\hat{X} \geq \hat{X}_{(k)}$ to disk,
+			assigning each a prior weight of $M/N$.
+	\State Create a region in the $L_1$-$L_2$ plane defined by
+				$\hat{X} \geq \hat{X}_{(k)}$, to be excluded at all subsequent
+				steps.
+	\State Update the remaining prior mass to $M \leftarrow (M - M/N)$.
+	\State Replace all particles with $\hat{X} \geq \hat{X}_{(k)}$ by new
+			particles generated from $\pi(\xx)$, subject to the constraint
+			of not being in the excluded region.
 \EndWhile
@@
 Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
-In step 2, we try to find a particle with the lowest possible nonzero
-corner count. Often this will be 1. In Figure~\ref{fig:joint2}, particle $K$
-is located at $(2.65, 1.31)$ and has a corner count of 1.
-If all particles have a corner count of 0, the algorithm terminates.
 

### 2015-11-23 Paper/ms.tex
remaining prior mass correct
more algorithm

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{algorithmicx}
@@
 \begin{algorithmic}
-\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
-initialise the {\em remaining prior mass} at $M=1$. The number of particles
+\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
 $N$ should be odd and quite large (1001 or greater).
+\State Initialise the {\em remaining prior mass} to be $M=1$.
+\State Initialise the {\em excluded region} $\mathcal{R}$ to be the empty set
+$\emptyset$.
 \While{more iterations desired}
@@
 			assigning each a prior weight of $M/N$.
-	\State Create a region in the $L_1$-$L_2$ plane defined by
-				$\hat{X} \geq \hat{X}_{(k)}$, to be excluded at all subsequent
-				steps.
-	\State Update the remaining prior mass to $M \leftarrow (M - M/N)$.
+	\State Exclude the region in the $L_1$-$L_2$ plane where
+$\hat{X} \geq \hat{X}_{(k)}$. That is, set
+	\State \hspace{20pt} $\mathcal{R} \leftarrow \mathcal{R} \cup \{(\ell_1, \ell_2) \in \mathbb{R}^2: \hat{X}(\ell_1, \ell_2) \geq \hat{X}_{(k)}\}$.
+	\State Update the remaining prior mass to $M \leftarrow M(1 - k/N)$.
 	\State Replace all particles with $\hat{X} \geq \hat{X}_{(k)}$ by new

### 2015-11-23 Code/C++/SamplerImpl.h
terminate

@@ -178,6 +178,12 @@ void Sampler<MyModel>::do_iteration()
 		}
 	}
 
+	if(num_dying == num_particles)
+	{
+		std::cout<<"# Cannot continue."<<std::endl;
+		exit(0);
+	}
+
 	// Place forbidding rectangles anywhere ucc >= threshold
 	for(int i=0; i<num_particles; i++)
 	{

### 2015-11-23 Paper/ms.tex

@@
 \mathds{1}\left[L_1 > \ell_1, L_2 > \ell_2 \right]
- \, dL_1 \, dL_2
+ \, dL_1 \, dL_2.
 \end{eqnarray}
+Since $X$ is a probability, it must satisfy a product rule:
+\begin{eqnarray}
+X(\ell_1, \ell_2) = X(\ell_1)X(\ell_2 | \ell_1) = X(\ell_2)X(\ell_1 | \ell_2).
+\end{eqnarray}
+
 When the prior $\pi$ is approximated with a set of $N$
@@
 is the fraction of particles in the rectangle whose upper right corner is at
-$\left(\ell_1, \ell_2\right)$ (see Figure~\ref{fig:joint1} for an
-example).
+$\left(\ell_1, \ell_2\right)$. See Figure~\ref{fig:joint1} for an
+example.
 
@@
 Diffusive Nested Sampling \citep{dnest}) has these properties.
+
+Consider running standard NS using the objective function $L_1(\xx)$. After some
+time, we'd know the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
 

### 2015-11-23 Code/C++/Models/Atoms.cpp
Attract to center of box

@@ -41,7 +41,7 @@ void Atoms::compute_scalars()
 	scalars[0] = -PE;
 	scalars[1] = 0.;
 	for(size_t i=0; i<x.size(); i++)
-		scalars[1] -= z[i];
+		scalars[1] += -pow(x[i] - 0.5, 2) - pow(z[i] - 0.5, 2);
 }
 
 void Atoms::calculate_PE()

### 2015-11-23 Code/C++/main.cpp
More mcmc_steps now that it's parallelized

@@ -23,7 +23,7 @@ int main(int argc, char** argv)
 		r.set_seed(++seed);
 
 	constexpr int num_particles = 3001;
-	constexpr int num_mcmc_steps = 1000;
+	constexpr int num_mcmc_steps = 5000;
 	constexpr double depth = 1000.;
 	constexpr int steps = depth/log(2.);
 

### 2015-11-25 Paper/ms.tex
Updated formula and fig

@@
 \end{eqnarray}
-The two objective functions are:
+The first objective function is quadratic in form, such that the canonical
+distribution is an iid gaussian centered at $(0.5, 0.5, ..., 0.5)$:
 \begin{eqnarray}
-L_1(\xx) &=& -\frac{1}{2}\sum_{i=1}^{100} \left(\frac{x_i - 0.5}{0.01}\right)^2\\
-L_2(\xx) &=& -\sum_{i=1}^{100} \log\left[1 +
-\left(\frac{x_i - 0.5}{0.01}\right)^2
+L_1(\xx) &=& -\sum_{i=1}^{100} \left(x_i - 0.5\right)^2.
+\end{eqnarray}
+
+The second objective function upweights states where coordinates $\{x_i\}$
+are close to 0, 0.25, 0.5, 0.75, or 1:
+\begin{eqnarray} 
+L_2(\xx) &=& -\sum_{i=1}^{100} sin^2(4\pi x)
 \right].
 \end{eqnarray}
-The canonical distribution at low temperature for $L_1$ and high temperature
-for $L_2$ is a gaussian centered at 0.5 with standard deviation
-$0.01\sqrt{T_1}$. At high temperatures for $L_2$ and low temperatures for $L_1$ the
-canonical distributions are Cauchy centered at 0.5 with width
-$0.01\sqrt{T_2}$.
 
@@
 points representing the canonical distribution at temperatures
-$(T_1, T_2) = (200, 100)$ (black).\label{fig:output}}
+$(T_1, T_2) = (0.1, 1)$ (black).\label{fig:output}}
 \end{figure}

### 2015-11-27 Paper/ms.tex
Fixed problem in equation
KL divergence

@@
 \end{eqnarray}
-The main goal of NS is to approximate $Z$.
-This is done by drawing particles from the
+The main goal of NS is to approximate $Z$, but it can also be used to make
+Monte Carlo approximations of the posterior, and calculate the ``information'',
+or Kullback-Leibler divergence from the prior to the posterior:
+\begin{eqnarray}
+\mathcal{I} &=& \int p(\xx | \data, \II) \ln
+\left[\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right] \, d\xx \\
+&=& \int \frac{\pi(\xx) L(\xx)}{Z} \ln
+\left[\frac{L(\xx)}{Z}\right] \, d\xx.
+\end{eqnarray}
+$\mathcal{I}$ quantifies how compressed the posterior distribution is with
+respect to the prior. For example, $\mathcal{I} = 10$ implies the posterior
+occupies a fraction $e^{-10}$ of the prior mass.
+
+NS works by drawing particles from the
 prior $\pi(\xx)$ and successively imposing constraints on the value of
@@
 \begin{eqnarray} 
-L_2(\xx) &=& -\sum_{i=1}^{100} sin^2(4\pi x)
-\right].
+L_2(\xx) &=& -\sum_{i=1}^{100} sin^2(4\pi x_i)
 \end{eqnarray}

### 2015-11-28 Paper/ms.tex
gibbs entropy
\sin
overloaded notation

@@
 \end{eqnarray}
+Throughout this paper we use the popular `overloaded' notation for some
+functions and probability distributions.
+For example, $L(\xx)$ is the likelihood function, and $L(X)$ is the likelihood
+as a function of the enclosed prior mass $X$. These cannot literally be the
+same function as the inputs are not even necessarily of the same dimensionality.
+Therefore the symbol $L$ represents the output of the function.
+
 In addition to the marginal likelihood $Z$, NS also allows us
@@
 compute $Z(\beta)$ as a function of $\beta$, which is called the
-{\it partition function}. NS can achieve this, allowing the study of
+{\it partition function}. The quantity $\mathcal{I}(\beta)$ is, up to a sign
+change and an additive constant, the Gibbs entropy.
+
+NS can achieve this, allowing the study of
 the properties of materials from first principles, based on hypotheses about
@@
 \begin{eqnarray} 
-L_2(\xx) &=& -\sum_{i=1}^{100} sin^2(4\pi x_i)
+L_2(\xx) &=& -\sum_{i=1}^{100} \sin^2(4\pi x_i)
 \end{eqnarray}

### 2015-11-28 Paper/ms.tex
More writing

@@
 
-Consider running standard NS using the objective function $L_1(\xx)$. After some
-time, we'd know the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
+Consider running standard NS with $N$ particles
+using the objective function $L_1(\xx)$.
+After $M$ iterations, we'd be able to approximate
+the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
+from $X\approx e^{-M/N}$ to $X=1$. Defining $Y=-\ln(X)$, we could also
+construct an approximation to the function
+$L(Y)$, a non-decreasing function of $Y$, from $Y=0$ to $Y\approx M/N$.
 

### 2015-11-29 Code/C++/Models/Atoms.cpp
200 particles

@@ -7,8 +7,8 @@ using namespace std;
 using namespace DNest3;
 
 Atoms::Atoms()
-:x(50), y(50), z(50)
-,terms(50, vector<double>(50))
+:x(200), y(200), z(200)
+,terms(200, vector<double>(200))
 ,scalars(2)
 {
 

### 2015-11-29 Paper/ms.tex

@@
 
+\section{Rationale}\label{sec:rationale}
+
+{\bf Here is where I'll attempt to provide a justify the
+algorithm}.
+
 Consider running standard NS with $N$ particles
@@
 
-Here we describe the proposed algorithm, which took three years of
+The proposed algorithm is given in Algorithm~\ref{alg:algorithm}
+and took three years of
 experimentation to develop. Please contact the first author for a list of
 failed ideas, some of which might lead to potential alternatives to this
-algorithm.
+algorithm. The basic idea is to try to mimic a standard NS run, where the
+objective function being increased is
+\begin{eqnarray}
+Y(\xx) &=& -\ln\left[X\left(L_1(\xx), L_2(\xx)\right)\right].
+\end{eqnarray}
 
-\begin{algorithm}
+\begin{algorithm}[ht!]
 \begin{algorithmic}
@@
 \end{algorithmic}
+\caption{\label{alg:algorithm}}
 \end{algorithm}
 
-Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
+%Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
 

### 2015-11-29 Code/C++/Models/CambridgeLJ.cpp
Working on the constructors and stuff

@@ -17,9 +17,10 @@ extern "C"
 					double* univar_30k, double* nvar_60k, bool* flat_v_prior);
 }
 
+const int CambridgeLJ::N = 100;
+
 CambridgeLJ::CambridgeLJ()
-:N(100)
-,scalars(2)
+:scalars(2)
 {
 	s = new double[3*N];
 }
@@ -32,8 +33,6 @@ CambridgeLJ::CambridgeLJ(const CambridgeLJ& other)
 
 CambridgeLJ& CambridgeLJ::operator = (const CambridgeLJ& other)
 {
-	assert(N == other.N);
-	N = other.N;
 	for(int i=0; i<N; i++)
 		s[i] = other.s[i];
 	scalars = other.scalars;

### 2015-11-29 Code/C++/Models/CambridgeLJ.h

@@ -8,9 +8,8 @@
 class CambridgeLJ
 {
 	private:
-		int N;
+		static const int N;
 		double* s;
-
 		std::vector<double> scalars;
 
 		void compute_scalars();

### 2015-12-03 Paper/ms.tex
Editing intro

@@
 estimating the partition function of a family of
-probability distributions indexed by two hyperparameters.}
+probability distributions indexed by two hyperparameters. These
+problems arise frequently in statistical mechanics.}
 
@@
 
-Nested Sampling \citep[NS][]{skilling} is an effective and popular
-Monte Carlo algorithm which can solve a wide range of problems in Bayesian
+Nested Sampling \citep[NS][]{skilling} is popular general
+Monte Carlo algorithm that can solve a wide range of problems in Bayesian
 computation and statistical mechanics
@@
 prior $\pi(\xx)$ and successively imposing constraints on the value of
-the likelihood $L(\xx)$ that compress the prior mass by an
-approximately-known factor. Like other related algorithms such as those
-based on `annealing', NS
+the likelihood $L(\xx)$ that compress the prior mass by a (roughly) known
+and constant factor. Like related algorithms such as those
+based on `annealing' or `tempering', NS
 moves through a sequence of probability distributions beginning from the
@@
 parallel tempering method \citep{pt}.
-However, unlike annealing methods, NS continues
-to work when the problem contains a phase transition.
+However, unlike annealing methods, NS requires only a small number of
+tuning parameters, and still works on problems involvin phase transitions.
 
@@
 change and an additive constant, the Gibbs entropy.
-
 NS can achieve this, allowing the study of
@@
 
+Throughout this paper we will refer to $L(\xx)$ or
+$-E(\xx)$ as `objective functions',
+by analogy with the terminology used when
+discussing optimization methods. Although we are interested in sampling rather
+than optimization, we still want to increase the values of $L$ or decrease
+the values of $E$ relative to what is typical of the prior.
+
 In some inference and
@@
 $\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
-we obtain testable information that fixes the expected values of two scalar
+we obtain information that fixes the expected values of two scalar
 functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$.
@@
 updated probability distribution that takes into account the constraints is
-of the `canonical' form:
+of the canonical form:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $\beta_1$ and $\beta_2$ are two `inverse temperatures'. In a Bayesian
+where $\beta_1$ and $\beta_2$ are two `inverse temperatures' (the temperatures
+themselves are $T_1 = \beta_1^{-1}$ and $T_2 = \beta_2^{-1}$).
+In a Bayesian
 context, this distribution (when $\beta_1 = \beta_2 = 1$) could be the
@@
 be $\exp\left(-\textnormal{energy}\right)$ and $L_2$ might be the total
-angular momentum.
-If we were only interested in a single canonical distribution, for example
-with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
-constant and by running standard Nested Sampling with `likelihood'
+angular momentum, or be a term depending on an external field (so $\beta_2$
+would then describe the strength of the external field).
+
+If we are only interested in a single canonical distribution, for example
+with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we can estimate its normalising
+constant and by running standard Nested Sampling with objective function
 $L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
 are interested in a range of values for $\beta_1$ and $\beta_2$, and we
-want to know the normalisation $Z(\beta_1, \beta_2)$, called the
-partition function. This work describes
+want to know the entire partition function $Z(\beta_1, \beta_2)$.
+This work describes
 progress towards solving this class of problems while maintaining the benefits
@@
 transitions. The algorithm is designed for problems with two objective
-functions, but should generalize to problems with more than two (though perhaps
-not more than a few). Throughout this paper we will refer $L_i(\xx)$ or
-$\exp\left[-E_i(\xx)\right]$ as `objective functions',
-by analogy with the terminology used when
-discussing optimizations methods. Although we are interested in sampling rather
-than optimization, we still want to increase the values of $L_i$ or decrease
-the values of $E_i$ relative to what is typical of the prior.
+functions, but may generalize to problems with more than two (though perhaps
+not more than a few).
 

### 2015-12-04 Paper/ms.tex
A few more words

@@
 $Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
-run. To be considered a variant of Nested Sampling, the method should satisfy
+run. To be a variant of Nested Sampling, the algorithm should satisfy
 the following properties:
 \begin{enumerate}
-\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
-\item The algorithm should seek to explore regions where the values of
-$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
+\item It should begin with $N$ points drawn from the prior $\pi(\xx)$.
+\item It should seek to explore regions where the values of
+$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior $\pi(L_1, L_2)$
+would typically imply.
 \item The algorithm should consider a sequence of probability
@@
 The standard Nested Sampling algorithm, (as well as alternatives such as
-Diffusive Nested Sampling \citep{dnest}) has these properties.
+Diffusive Nested Sampling \citep{dnest}) has these properties but for a
+single objective function.
 
@@
 
-Consider running standard NS with $N$ particles
+Consider the following proposal. Imagine running standard NS with $N$ particles
 using the objective function $L_1(\xx)$.
-After $M$ iterations, we'd be able to approximate
+After $M_1$ iterations, we'd be able to approximate
 the function $X(\ell_1) = \int \pi(\xx) \, d\xx$
-from $X\approx e^{-M/N}$ to $X=1$. Defining $Y=-\ln(X)$, we could also
+from $X\approx e^{-M_1/N}$ to $X=1$. Defining $Y=-\ln(X)$, we could also
 construct an approximation to the function
-$L(Y)$, a non-decreasing function of $Y$, from $Y=0$ to $Y\approx M/N$.
+$L_1(Y)$, a non-decreasing function of $Y$, from $Y=0$ to $Y\approx M_1/N$.
+After $M$ iterations, we keep the last value of $L_1(\xx)$ obtained
+(call it $L_1^*$), and continue to impose the constraint $L_1(\xx) > L_1^*$.
+
+Subject to the constraint $L_1(\xx) > L_1^*$, standard NS can continue but
+attempting to increase $L_2$ instead of $L_1$. After $M_2$ such iterations,
+the algorithm would have compressed the prior by a factor
+$e^{(M_1 + M_2)/N}$ and therefore ended at a point in the $(L_1, L_2)$ plane
+where $X(L_1, L_2) \approx e^{-(M_1 + M_2)/N}$. Any such run consisting of
+$M_1 + M_2$ iterations will therefore appoximately finish on a contour line
+of the function $X(L_1, L_2)$.
 
@@
 
-The proposed algorithm is given in Algorithm~\ref{alg:algorithm}
-and took three years of
-experimentation to develop. Please contact the first author for a list of
-failed ideas, some of which might lead to potential alternatives to this
-algorithm. The basic idea is to try to mimic a standard NS run, where the
+The basic idea of our algorithm is to try to mimic a standard NS run, where the
 objective function being increased is
@@
 \end{eqnarray}
+Standard NS cannot be applied simply because we cannot evaluate $X(L_1, L_2)$ and
+therefore cannot restrict the parameter space based on the value of $X$ (or
+equivalently $Y$. However, given a number of Monte Carlo samples, we
+can approximate this.
 

### 2015-12-04 Code/Julia/idea.jl
Same distribution from the paper

@@ -4,15 +4,15 @@ using PyCall
 include("Utils.jl")
 
 # Number of pixels
-(ni, nj) = (1024, 1024)
+(ni, nj) = (256, 256)
 
 # Prior distribution
 log_prior = Array(Float64, (ni, nj))
 for(j in 1:nj)
-	x = (j - 0.5)/nj
+	x = 5.0*(j - 1)/(nj - 1)
 	for(i in 1:ni)
-		y = 1.0 - (i - 0.5)/ni
-		log_prior[i, j] = -0.5*(x - 0.1)^2/0.03^2 - 0.5*(y - x^2)^2/0.03^2
+		y = 5.0 - 5*(i - 1)/(ni - 1)
+		log_prior[i, j] = -0.5*(x - 1.5)^2/1.0^2 - 0.5*((y - 5*(x/5)^2)^2)/1.0^2
 	end
 end
 log_prior -= logsumexp(vec(log_prior))

### 2015-12-04 Code/Julia/idea.jl
contours

@@ -38,23 +38,15 @@ for(j in nj:-1:1)
 	end
 end
 
-plt.figure(figsize=(14, 6))
-plt.subplot(1, 2, 1)
-plt.imshow(exp(log_prior), cmap=colormaps.viridis, interpolation="nearest")
+plt.rc("font", size=18, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=true)
+plt.imshow(exp(log_prior), extent=[0, 5, 0, 5], cmap="Blues", interpolation="nearest")
 plt.title("\$\\pi(L_1, L_2)\$", fontsize=16)
 plt.xlabel("\$L_1\$", fontsize=16)
 plt.ylabel("\$L_2\$", fontsize=16)
-plt.gca()[:set_xticklabels]([])
-plt.gca()[:set_yticklabels]([])
-
-plt.subplot(1, 2, 2)
-p = exp(log_prior - logG)
-upper = maximum(sort(vec(p))[1:Int64(floor(0.95*ni*nj))])
-plt.imshow(p, vmax=upper, cmap=colormaps.viridis, interpolation="nearest")
-plt.title("\$\\pi(L_1, L_2)/G(L_1, L_2)\$", fontsize=16)
-plt.xlabel("\$L_1\$", fontsize=16)
-plt.ylabel("\$L_2\$", fontsize=16)
-plt.gca()[:set_xticklabels]([])
-plt.gca()[:set_yticklabels]([])
+plt.contour(logG, extent=[0, 5, 0, 5], origin="upper", levels=-log(2.0)*(1:15), colors="k", linestyles="solid", alpha=0.5)
+plt.gca()[:set_xticklabels]([0, 1, 2, 3, 4, 5])
+plt.gca()[:set_yticklabels]([0, 1, 2, 3, 4, 5])
+plt.savefig("contours.pdf", bbox_inches="tight")
 plt.show()
 

### 2015-12-04 Paper/ms.tex
New figure

@@
 
-{\bf Here is where I'll attempt to provide a justify the
-algorithm}.
+{\bf Here is where I'll attempt to justify the algorithm}.
 
@@
 of the function $X(L_1, L_2)$.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{figures/contours.pdf}
+\caption{\it The joint prior for the objective functions, and fifteen
+contours of the cumulative distribution $X(L_1, L_2)$, each enclosing half
+the prior mass of the previous one. \label{fig:contours}}
+\end{figure}
 

### 2015-12-04 Paper/figures/contours.jl

@@ -1,0 +1,52 @@
+using PyCall
+@pyimport matplotlib.pyplot as plt
+@pyimport colormaps
+include("../../Code/Julia/Utils.jl")
+
+# Number of pixels
+(ni, nj) = (256, 256)
+
+# Prior distribution
+log_prior = Array(Float64, (ni, nj))
+for(j in 1:nj)
+	x = 5.0*(j - 1)/(nj - 1)
+	for(i in 1:ni)
+		y = 5.0 - 5*(i - 1)/(ni - 1)
+		log_prior[i, j] = -0.5*(x - 1.5)^2/1.0^2 - 0.5*((y - 5*(x/5)^2)^2)/1.0^2
+	end
+end
+log_prior -= logsumexp(vec(log_prior))
+
+# Upper corner mass
+logG = Array(Float64, (ni, nj))
+for(j in nj:-1:1)
+	for(i in 1:ni)
+		up = -Inf
+		right = -Inf
+		up_and_right = -Inf
+		if(i != 1)
+			up = logG[i-1, j]
+		end
+		if(j != nj)
+			right = logG[i, j+1]
+		end
+		if((i != 1) && (j != nj))
+			up_and_right = logG[i-1, j+1]
+		end
+		logG[i, j] = logsumexp([log_prior[i, j], up, right])
+		logG[i, j] = logdiffexp(logG[i, j], up_and_right)
+	end
+end
+
+plt.rc("font", size=18, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=true)
+plt.imshow(exp(log_prior), extent=[0, 5, 0, 5], cmap="Blues", interpolation="nearest")
+plt.title("\$\\pi(L_1, L_2)\$", fontsize=16)
+plt.xlabel("\$L_1\$", fontsize=16)
+plt.ylabel("\$L_2\$", fontsize=16)
+plt.contour(logG, extent=[0, 5, 0, 5], origin="upper", levels=-log(2.0)*(1:15), colors="k", linestyles="solid", alpha=0.5)
+plt.gca()[:set_xticklabels]([0, 1, 2, 3, 4, 5])
+plt.gca()[:set_yticklabels]([0, 1, 2, 3, 4, 5])
+plt.savefig("contours.pdf", bbox_inches="tight")
+plt.show()
+

### 2015-12-04 Code/OldC++/Model.cpp
Removed old stuff

@@ -1,44 +1,0 @@
-#include "Model.h"
-#include <RandomNumberGenerator.h>
-#include <Utils.h>
-
-using namespace DNest3;
-using namespace std;
-
-Model::Model(int num_scalars)
-:num_scalars(num_scalars), scalars(num_scalars), tiebreakers(num_scalars)
-{
-
-}
-
-Model::~Model()
-{
-
-}
-
-void Model::from_prior_tiebreakers()
-{
-	for(size_t i=0; i<tiebreakers.size(); i++)
-		tiebreakers[i] = randomU();
-}
-
-double Model::perturb_tiebreakers()
-{
-	int which = randInt(tiebreakers.size());
-	tiebreakers[which] += randh();
-	wrap(tiebreakers[which], 0., 1.);
-	return 0.;
-}
-
-bool Model::is_above(const vector< vector<double> >& threshold) const
-{
-	assert(scalars.size() == threshold.size());
-	for(size_t i=0; i<scalars.size(); i++)
-	{
-		if((scalars[i] < threshold[i][0]) ||
-			((scalars[i] == threshold[i][0] && tiebreakers[i] < threshold[i][1])))
-			return false;
-	}
-	return true;
-}
-

### 2015-12-04 Code/OldC++/Model.h

@@ -1,55 +1,0 @@
-#ifndef _Model_
-#define _Model_
-
-#include <vector>
-#include <ostream>
-/*
-* Abstract class defining what is needed to implement a model.
-* This will be like DNest. An object of a subclass represents a
-* Model in parameter space.
-*/
-
-class Model
-{
-	protected:
-		int num_scalars;
-		std::vector<double> scalars;
-		std::vector<double> tiebreakers;
-
-	public:
-		// Pass in the number of scalars
-		Model(int num_scalars);
-
-		virtual ~Model();
-
-		// Generate the Model from the prior
-		virtual void from_prior() = 0;
-
-		// Do a Metropolis-Hastings proposal that satisfies detailed
-		// balance wrt the prior. Return log of any ratios needed to
-		// achieve this.
-		virtual double perturb() = 0;
-
-		// Methods for tiebreakers
-		void from_prior_tiebreakers();
-		double perturb_tiebreakers();
-
-		// Are all scalars greater than or equal to the threshold
-		// values?
-		bool is_above(const std::vector< std::vector<double> >&
-					threshold) const;
-
-		// Getters for scalars and tiebreakers
-		int get_num_scalars() const { return num_scalars; }
-		const std::vector<double>& get_scalars() const
-		{
-			return scalars;
-		}
-		const std::vector<double>& get_tiebreakers() const
-		{
-			return tiebreakers;
-		}
-};
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/Atoms.cpp

@@ -1,105 +1,0 @@
-#include "Atoms.h"
-#include "RandomNumberGenerator.h"
-#include "Utils.h"
-#include <cmath>
-
-using namespace std;
-using namespace DNest3;
-
-Atoms::Atoms()
-:Model(2), x(50), y(50), z(50)
-,terms(50, vector<double>(50))
-{
-
-}
-
-Atoms::~Atoms()
-{
-
-}
-
-void Atoms::from_prior()
-{
-	for(size_t i=0; i<x.size(); i++)
-	{
-		x[i] = randomU();
-		y[i] = randomU();
-		z[i] = randomU();
-	}
-
-	for(size_t i=0; i<terms.size(); i++)
-		for(size_t j=(i+1); j<terms[i].size(); j++)
-			calculate_PE(i, j);
-
-	calculate_PE();
-	compute_scalars();
-}
-
-void Atoms::compute_scalars()
-{
-	scalars[0] = -PE;
-	scalars[1] = 0.;
-	for(size_t i=0; i<x.size(); i++)
-		scalars[1] -= z[i];
-}
-
-void Atoms::calculate_PE()
-{
-	PE = 0.;
-	for(size_t i=0; i<x.size(); i++)
-		for(size_t j=(i+1); j<x.size(); j++)
-			PE += terms[i][j];
-}
-
-void Atoms::calculate_PE(int i, int j)
-{
-	double Rmsq = pow(0.05, 2);
-	double rsq = pow(x[i] - x[j], 2) + pow(y[i] - y[j], 2) + pow(z[i] - z[j], 2);
-	terms[i][j] = pow(Rmsq/rsq, 6) - 2.*pow(Rmsq/rsq, 3);
-}
-
-double Atoms::perturb()
-{
-	int which = randInt(x.size());
-	int what = randInt(3);
-
-	if(what == 0)
-	{
-		x[which] += randh();
-		wrap(x[which], 0., 1.);
-	}
-	if(what == 1)
-	{
-		y[which] += randh();
-		wrap(y[which], 0., 1.);
-	}
-	if(what == 2)
-	{
-		z[which] += randh();
-		wrap(z[which], 0., 1.);
-	}
-
-	for(int i=0; i<which; i++)
-		calculate_PE(i, which);
-
-	for(int j=(which+1); j<int(x.size()); j++)
-		calculate_PE(which, j);
-
-
-	calculate_PE();
-	compute_scalars();
-
[…]

### 2015-12-04 Code/OldC++/Models/Atoms.h

@@ -1,42 +1,0 @@
-#ifndef _Atoms_
-#define _Atoms_
-
-#include "Model.h"
-#include <vector>
-
-class Atoms:public Model
-{
-	private:
-		// Positions
-		std::vector<double> x, y, z;
-
-		std::vector< std::vector<double> > terms;
-
-		// Energy
-		long double PE;
-
-		// From scratch, total
-		void calculate_PE();
-
-		// Just a single pair
-		void calculate_PE(int i, int j);
-		void compute_scalars();
-
-	public:
-		Atoms();
-		~Atoms();
-
-		// Generate the point from the prior
-		void from_prior();
-
-		// Metropolis-Hastings proposals
-		double perturb();
-
-		friend std::ostream& operator << (std::ostream& out, const
-								Atoms& a);
-};
-
-std::ostream& operator << (std::ostream& out, const Atoms& a);
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/Gravity.cpp

@@ -1,130 +1,0 @@
-#include "Gravity.h"
-#include "RandomNumberGenerator.h"
-#include "Utils.h"
-
-#include <cmath>
-
-using namespace DNest3;
-using namespace std;
-
-Gravity::Gravity()
-:Model(2)
-,x(1000), y(1000), z(1000), vx(1000), vy(1000), vz(1000)
-,staleness(0)
-{
-
-}
-
-void Gravity::from_prior()
-{
-	for(size_t i=0; i<x.size(); i++)
-	{
-		x[i] = -10. + 20.*randomU();
-		y[i] = -10. + 20.*randomU();
-		z[i] = -10. + 20.*randomU();
-		vx[i] = -10. + 20.*randomU();
-		vy[i] = -10. + 20.*randomU();
-		vz[i] = -10. + 20.*randomU();
-	}
-
-	refresh();
-	compute_scalars();
-}
-
-void Gravity::increment(int i, int sign)
-{
-	KE += sign*0.5*(vx[i]*vx[i] + vy[i]*vy[i] + vz[i]*vz[i]);
-	L += sign*(x[i]*vy[i] - y[i]*vx[i]);
-
-	double rsq;
-	for(size_t j=0; j<x.size(); j++)
-	{
-		if(i != static_cast<int>(j))
-		{
-			rsq = pow(x[i] - x[j], 2) + pow(y[i] - y[j], 2)
-				+ pow(z[i] - z[j], 2);
-			if(rsq <= 0.01)
-				rsq = 0.01;
-			PE += -1./sqrt(rsq)*sign;
-		}
-	}
-}
-
-double Gravity::perturb()
-{
-	int reps = 1;
-	if(randomU() <= 0.5)
-		reps += 1 + randInt(9);
-
-	int which;
-	for(int i=0; i<reps; i++)
-	{
-		which = randInt(x.size());
-
-		// Remove effect of this particle
-		increment(which, -1);
-
-		x[which] += 20.*randh();
-		x[which] = mod(x[which] + 10., 20.) - 10.;
-
-		y[which] += 20.*randh();
-		y[which] = mod(y[which] + 10., 20.) - 10.;
-
-		z[which] += 20.*randh();
-		z[which] = mod(z[which] + 10., 20.) - 10.;
-
-		vx[which] += 20.*randh();
-		vx[which] = mod(vx[which] + 10., 20.) - 10.;
-
[…]

### 2015-12-04 Code/OldC++/Models/Gravity.h

@@ -1,28 +1,0 @@
-#ifndef _Gravity_
-#define _Gravity_
-
-#include "Model.h"
-#include <vector>
-#include <ostream>
-
-class Gravity:public Model
-{
-	private:
-		std::vector<double> x, y, z, vx, vy, vz;
-
-		long double PE, KE, L;
-		int staleness;
-
-		void refresh();
-		void increment(int i, int sign);
-		void compute_scalars();
-	public:
-		Gravity();
-		void from_prior();
-		double perturb();
-
-	friend std::ostream& operator << (std::ostream& out, const Gravity& e);
-};
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/ImageEntropy.cpp

@@ -1,118 +1,0 @@
-#include "ImageEntropy.h"
-#include <cassert>
-#include <cmath>
-#include <RandomNumberGenerator.h>
-#include <Utils.h>
-#include <fstream>
-
-using namespace std;
-using namespace DNest3;
-
-vector< vector<double> > ImageEntropy::data(100, vector<double>(100));
-PSF ImageEntropy::psf(21);
-PSF ImageEntropy::preblur(5);
-
-void ImageEntropy::load_data()
-{
-	fstream fin("data.txt", ios::in);
-	for(size_t i=0; i<data.size(); i++)
-		for(size_t j=0; j<data[i].size(); j++)
-			fin>>data[i][j];
-	fin.close();
-
-	psf.load("psf.txt");
-	psf.calculate_fft(100, 100);
-
-	preblur.load("preblur.txt");
-	preblur.calculate_fft(100, 100);
-}
-
-ImageEntropy::ImageEntropy()
-:Model(2)
-,image(100, vector<double>(100))
-{
-
-}
-
-ImageEntropy::~ImageEntropy()
-{
-
-}
-
-void ImageEntropy::from_prior()
-{
-	for(size_t i=0; i<image.size(); i++)
-		for(size_t j=0; j<image[i].size(); j++)
-			image[i][j] = randomU();
-	compute_scalars();
-}
-
-double ImageEntropy::perturb()
-{
-	int reps = 1;
-	if(randomU() <= 0.9)
-		reps = (int)pow(10., 4.*randomU());
-
-	int ii, jj;
-	for(int i=0; i<reps; i++)
-	{
-		ii = randInt(image.size());
-		jj = randInt(image[ii].size());
-		image[ii][jj] += randh();
-		wrap(image[ii][jj], 0., 1.);
-	}
-
-	compute_scalars();
-	return 0.;
-}
-
-void ImageEntropy::compute_scalars()
-{
-	// Find image total (use entropy of normalised image)
-	double tot = 0.;
-	for(size_t i=0; i<image.size(); i++)
-		for(size_t j=0; j<image[i].size(); j++)
-			tot += image[i][j];
-
-	// Find image entropy
-	double S = 0.;
-	for(size_t i=0; i<image.size(); i++)
-		for(size_t j=0; j<image[i].size(); j++)
-			S += -(image[i][j]/tot)*log(image[i][j]/tot + 1E-300);
-	scalars[0] = S;
-
[…]

### 2015-12-04 Code/OldC++/Models/ImageEntropy.h

@@ -1,41 +1,0 @@
-#ifndef _ImageEntropy_
-#define _ImageEntropy_
-
-/*
-* An object of this class represents a
-* point in the parameter space.
-*/
-
-#include <vector>
-#include <ostream>
-#include "Model.h"
-#include "PSF.h"
-
-class ImageEntropy:public Model
-{
-	private:
-		// Data and PSF
-		static PSF psf;
-		static std::vector< std::vector<double> > data;
-		static PSF preblur;
-
-		std::vector< std::vector<double> > image;
-		void compute_scalars();	
-
-	public:
-		ImageEntropy();
-		~ImageEntropy();
-
-		void from_prior();
-		double perturb();
-
-		static void load_data();
-
-		friend std::ostream& operator << (std::ostream& out,
-							const ImageEntropy& m);
-};
-
-std::ostream& operator << (std::ostream& out, const ImageEntropy& m);
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/PSF.cpp

@@ -1,176 +1,0 @@
-#include "PSF.h"
-#include "Utils.h"
-#include <cassert>
-#include <iostream>
-#include <fstream>
-
-using namespace std;
-using namespace DNest3;
-using namespace arma;
-
-PSF::PSF(int size)
-:size(size)
-,pixels(size, vector<double>(size, 0.))
-,fft_ready(false)
-{
-	assert(size%2 == 1);
-	pixels[size/2][size/2] = 1.;
-}
-
-void PSF::set_size(int new_size)
-{
-	size = new_size;
-	pixels.assign(size, vector<double>(size, 0.));
-	pixels[size/2][size/2] = 1.;
-}
-
-void PSF::load(const char* filename)
-{
-	fstream fin(filename, ios::in);
-	if(!fin)
-	{
-		cerr<<"# ERROR: couldn't open file "<<filename<<"."<<endl;
-		return;
-	}
-	for(int i=0; i<size; i++)
-		for(int j=0; j<size; j++)
-			fin>>pixels[i][j];
-	fin.close();
-	normalise();
-}
-
-void PSF::normalise()
-{
-	double sum = 0.;
-	for(int i=0; i<size; i++)
-		for(int j=0; j<size; j++)
-			sum += pixels[i][j];
-
-	for(int i=0; i<size; i++)
-		for(int j=0; j<size; j++)
-			pixels[i][j] /= sum;
-}
-
-void PSF::calculate_fft(int Ni, int Nj)
-{
-	// Make the psf the same size as the image
-	mat psf(Ni, Nj);
-	psf.zeros();
-
-	int ni = pixels.size();
-	int nj = pixels[0].size();
-
-	int m, n;
-	for(int i=0; i<ni; i++)
-	{
-		m = mod(i - ni/2, Ni);
-		for(int j=0; j<nj; j++)
-		{
-			n = mod(j - nj/2, Nj);
-			psf(m, n) = pixels[i][j];
-		}
-	}
-
-	fft_of_psf = fft2(psf);
-	fft_ready = true;
-}
-
-void PSF::blur_image(vector< vector<double> >& img) const
-{
-	// Make result image. Assume img is rectangular...
-	vector< vector<double> > result(img.size(),
-					vector<double>(img[0].size(), 0.));
-
-	int h = size/2;
-	int ii, jj;
-	int M = static_cast<int>(img.size());
[…]

### 2015-12-04 Code/OldC++/Models/PSF.h

@@ -1,38 +1,0 @@
-#ifndef _PSF_
-#define _PSF_
-
-#include <vector>
-#include <armadillo>
-
-class PSF
-{
-	private:
-		int size;
-		std::vector< std::vector<double> > pixels;
-
-		// FFT of the PSF
-		arma::cx_mat fft_of_psf;
-		bool fft_ready;
-
-	public:
-		PSF(int size);
-		// Uses loopy method
-		void blur_image(std::vector< std::vector<double> >& img) const;
-		// Uses fft method
-		void blur_image2
-			(std::vector< std::vector<double> >& img) const;
-		void normalise();
-		void calculate_fft(int Ni, int Nj);
-		void load(const char* filename);
-		void set_size(int new_size);
-
-		// Getter for pixels
-		const std::vector< std::vector<double> >& get_pixels() const
-		{ return pixels; }
-
-		// Unit test
-		static void test();
-};
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/Potts.cpp

@@ -1,142 +1,0 @@
-#include "Potts.h"
-#include "RandomNumberGenerator.h"
-#include "Utils.h"
-
-#include <cmath>
-
-using namespace DNest3;
-using namespace std;
-
-Potts::Potts()
-:Model(2)
-,x(50, vector<int>(50))
-{
-
-}
-
-void Potts::from_prior()
-{
-	for(size_t i=0; i<x.size(); i++)
-		for(size_t j=0; j<x[i].size(); j++)
-			x[i][j] = randInt(num_colors);
-
-	compute_score();
-	compute_scalars();
-}
-
-void Potts::compute_score()
-{
-	score = 0;
-	score2 = 0;
-
-	// Coordinates of neighbouring cells
-	vector<int> ii(4); vector<int> jj(4);
-
-	for(size_t i=0; i<x.size(); i++)
-	{
-		for(size_t j=0; j<x[i].size(); j++)
-		{
-			for(int k=0; k<4; k++)
-			{
-				ii[k] = i;
-				jj[k] = j;
-			}
-			// Down, up, right, left
-			ii[0] = mod(i + 1, static_cast<int>(x.size()));
-			ii[1] = mod(i - 1, static_cast<int>(x.size()));
-			jj[2] = mod(j + 1, static_cast<int>(x[i].size()));
-			jj[3] = mod(j - 1, static_cast<int>(x[i].size()));
-
-			for(int k=0; k<4; k++)
-			{
-				if(x[i][j] == x[ii[k]][jj[k]])
-				{
-					score++;
-					if(k >= 2)
-						score2++;
-				}
-			}
-		}
-	}
-}
-
-#include <iostream>
-
-double Potts::perturb()
-{
-	int reps = 1;
-	if(rand() <= 0.5)
-		reps += 1 + randInt(9);
-
-	// Which cell is being perturbed
-	int i, j;
-
-	// Value of proposal
-	int proposal;
-
-	// Coordinates of neighbouring cells
-	vector<int> ii(4); vector<int> jj(4);
-	for(int z=0; z<reps; z++)
-	{
-		i = randInt(x.size());
-		j = randInt(x[0].size());
-
-		for(int k=0; k<4; k++)
-		{
-			ii[k] = i;
-			jj[k] = j;
-		}
-		// Down, up, right, left
-		ii[0] = mod(i + 1, static_cast<int>(x.size()));
[…]

### 2015-12-04 Code/OldC++/Models/Potts.h

@@ -1,30 +1,0 @@
-#ifndef _Potts_
-#define _Potts_
-
-#include "Model.h"
-#include <vector>
-#include <ostream>
-
-class Potts:public Model
-{
-	private:
-		static const int num_colors = 5;
-
-		std::vector< std::vector<int> > x;
-
-		// So we can just update on the fly
-		int score;
-		int score2;
-
-		void compute_score();
-		void compute_scalars();
-	public:
-		Potts();
-		void from_prior();
-		double perturb();
-
-	friend std::ostream& operator << (std::ostream& out, const Potts& e);
-};
-
-#endif
-

### 2015-12-04 Code/OldC++/Models/SimpleExample.cpp

@@ -1,76 +1,0 @@
-#include "SimpleExample.h"
-#include <cassert>
-#include <cmath>
-#include <RandomNumberGenerator.h>
-#include <Utils.h>
-
-using namespace std;
-using namespace DNest3;
-
-SimpleExample::SimpleExample()
-:Model(2)
-,params(20)
-{
-
-}
-
-SimpleExample::~SimpleExample()
-{
-
-}
-
-void SimpleExample::from_prior()
-{
-	for(size_t i=0; i<params.size(); i++)
-		params[i] = randomU();
-	compute_scalars();
-}
-
-double SimpleExample::perturb()
-{
-	int reps = 1 + ((randomU() <= 0.5)?(0):(1 + randInt(9)));
-	vector<bool> change(params.size(), false);
-	for(int i=0; i<reps; i++)
-		change[randInt(params.size())] = true;
-
-	for(size_t i=0; i<params.size(); i++)
-	{
-		if(change[i])
-		{
-			params[i] += randh();
-			wrap(params[i], 0., 1.);
-		}
-	}
-	compute_scalars();
-
-	return 0.;
-}
-
-void SimpleExample::compute_scalars()
-{
-	double u = 0.01;
-	double v = 0.1;
-	double C = log(1.0/sqrt(2*M_PI));
-	double logl1 = 0;
-	double logl2 = 0;
-
-	for(size_t i=0; i<params.size(); i++)
-	{
-		logl1 += C - log(u) - 0.5*pow((params[i] - 0.0)/u, 2);
-		logl2 += C - log(v) - 0.5*pow(params[i]/v, 2);
-	}
-	logl1 += log(100.0);
-
-	scalars[0] = logsumexp(logl1, logl2);
-	scalars[1] = 0.;
-	for(size_t i=0; i<params.size(); i++)
-		scalars[1] += -pow(sin(4.*M_PI*params[i]), 2);
-}
-
-ostream& operator << (ostream& out, const SimpleExample& m)
-{
-	for(size_t i=0; i<m.params.size(); i++)
-		out<<m.params[i]<<' ';
-	return out;
-}
-

### 2015-12-04 Code/OldC++/Models/SimpleExample.h

@@ -1,34 +1,0 @@
-#ifndef _SimpleExample_
-#define _SimpleExample_
-
-/*
-* An object of this class represents a
-* point in the parameter space.
-*/
-
-#include <vector>
-#include <ostream>
-#include "Model.h"
-
-class SimpleExample:public Model
-{
-	private:
-		std::vector<double> params;
-
-		void compute_scalars();	
-
-	public:
-		SimpleExample();
-		~SimpleExample();
-
-		void from_prior();
-		double perturb();
-
-		friend std::ostream& operator << (std::ostream& out,
-							const SimpleExample& m);
-};
-
-std::ostream& operator << (std::ostream& out, const SimpleExample& m);
-
-#endif
-

### 2015-12-04 Code/OldC++/Runs/gather.cpp

@@ -1,138 +1,0 @@
-#include <iostream>
-#include <fstream>
-#include <vector>
-#include <iomanip>
-#include <cmath>
-#include <string>
-#include <algorithm>
-
-using namespace std;
-
-
-double logsumexp(double x1, double x2)
-{
-	if(x1 > x2)
-		return x1 + log(1. + exp(x2 - x1));
-	return x2 + log(1. + exp(x1 - x2));
-}
-
-
-int main()
-{
-	// Temperatures
-	double T1 = 0.1; double T2 = 1.;
-
-	// Normalising constant for prior weights
-	double logC = -1E300;
-
-	// Prior weights (relative) and 'likelihoods'
-	vector<double> logw, logL;
-
-	bool save = true;
-
-	vector<string> directories;
-	directories.push_back(string(".."));
-
-	fstream fout1, fout2;
-	if(save)
-	{
-		fout1.open("logw.txt", ios::out);
-		fout2.open("scalars.txt", ios::out);
-	}
-
-	int k = 0;
-	for(size_t i=0; i<directories.size(); i++)
-	{
-		string filename1 = directories[i] + string("/logw.txt");
-		fstream fin1(filename1.c_str(), ios::in);
-
-		string filename2 = directories[i] + string("/scalars.txt");
-		fstream fin2(filename2.c_str(), ios::in);
-
-		double temp1, temp2, temp3;
-		while(fin1>>temp1 && fin2>>temp2 && fin2>>temp3)
-		{
-			if(save)
-			{
-				fout1<<temp1<<endl;
-				fout2<<temp2<<' '<<temp3<<endl;
-			}
-
-			logC = logsumexp(logC, temp1);
-			logw.push_back(temp1);
-			logL.push_back(temp2/T1 + temp3/T2);
-
-			k++;
-		}
-
-		fin1.close();
-		fin2.close();
-
-		cout<<"# Loaded "<<k<<" points."<<endl;
-	}
-
-	if(save)
-	{
-		fout1.close();
-		fout2.close();
-	}
-
-	// Calculate log(Z)
-	double logZ = -1E300;
-	for(size_t i=0; i<logL.size(); i++)
-		logZ = logsumexp(logZ, (logw[i] - logC) + logL[i]);
-
-	// Calculate partial log(Z)s from each run
-	double logw_max = *max_element(logw.begin(), logw.end());
[…]

### 2015-12-04 Code/OldC++/Sampler.h

@@ -1,62 +1,0 @@
-#ifndef _Sampler_
-#define _Sampler_
-
-#include <vector>
-#include <cmath>
-#include <iostream>
-#include <iomanip>
-#include <fstream>
-#include <boost/thread.hpp>
-#include "Utils.h"
-
-template<class Type>
-class Sampler
-{
-	private:
-		const int num_threads, num_particles, steps, thin;
-		const double peel_factor;
-
-		// Thread-specific rngs
-		std::vector<gsl_rng*> rngs;
-
-		int iterations;
-		std::vector<Type> particles;
-
-		std::vector< std::vector<double> > thresholds;
-		std::vector< std::vector<double> > thresholds_tiebreakers;
-		double log_prior_mass;
-
-		// How many of the thresholds is the particle below?
-		int badness(const Type& particle) const;
-		// Is s1 below s2?
-		bool is_below(const std::vector<double>& s1,
-				const std::vector<double>& s2,
-				const std::vector<double>& tb1,
-				const std::vector<double>& tb2) const;
-
-		void create_threshold(const std::vector< std::vector<double> >&
-						keep,
-					const std::vector< std::vector<double> >&
-						tiebreakers);
-		void remove_redundant_thresholds();
-		void runThread(int thread, const std::vector<int>& which_particles,
-				std::vector<int>& bad, int& accepts);
-
-		double compute_frac_below(int i, const std::vector< std::vector<double> >& keep,
-						const std::vector< std::vector<double> >& tiebreakers) const;
-
-
-	public:
-		Sampler(int num_threads, int num_particles, int steps, double peel_factor,
-				int thin);
-		~Sampler();
-
-		void initialise();
-		void explore();
-		void refresh();
-};
-
-#include "SamplerImpl.h"
-
-#endif
-

### 2015-12-04 Code/OldC++/SamplerImpl.h

@@ -1,319 +1,0 @@
-
-/*********************************************************************
- *			IMPLEMENTATIONS BEGIN			     *
- *********************************************************************/
-
-template<class Type>
-Sampler<Type>::Sampler(int num_threads, int num_particles, int steps, 					double peel_factor, int thin)
-:num_threads(num_threads)
-,num_particles(num_particles)
-,steps(steps)
-,thin(thin)
-,peel_factor(peel_factor)
-,rngs(num_threads)
-,iterations(0)
-,particles(num_particles)
-,log_prior_mass(0.)
-{
-	// Initialise rngs
-	for(int i=0; i<num_threads; i++)
-		rngs[i] = gsl_rng_alloc(gsl_rng_mt19937);
-	std::cout<<std::setprecision(8);
-}
-
-template<class Type>
-int Sampler<Type>::badness(const Type& particle) const
-{
-	int count = 0;
-	for(int i=int(thresholds.size())-1; i >= 0; i--)
-	{
-		if(is_below(particle.get_scalars(), thresholds[i],
-				particle.get_tiebreakers(), thresholds_tiebreakers[i]))
-		{
-			count++;
-			if(count == 2)
-				return count;
-		}
-	}
-	return count;
-}
-
-template<class Type>
-Sampler<Type>::~Sampler()
-{
-	for(int i=0; i<num_threads; i++)
-		gsl_rng_free(rngs[i]);
-}
-
-template<class Type>
-void Sampler<Type>::initialise()
-{
-	for(size_t i=0; i<particles.size(); i++)
-	{
-		particles[i].from_prior();
-		particles[i].from_prior_tiebreakers();
-	}
-
-	thresholds.clear();
-	thresholds_tiebreakers.clear();
-
-	// Open output files
-	std::fstream fout("output.txt", std::ios::out);
-	fout.close();
-	fout.open("sample.txt", std::ios::out);
-	fout.close();
-
-	// Set rng seeds
-	for(int i=0; i<num_threads; i++)
-		gsl_rng_set(rngs[i], time(0) + 10*(i+1));
-}
-
-template<class Type>
-void Sampler<Type>::runThread(int thread, const std::vector<int>& which_particles, std::vector<int>& bad, int& accepts)
-{
-	// Initialise thread-specific rng
-	DNest3::RandomNumberGenerator::initialise_instance();
[…]

### 2015-12-04 Code/OldC++/Utils.jl

@@ -1,15 +1,0 @@
-# Heavy tailed proposal
-function randh()
-  return 10.^(1.5 - 6.*rand())*randn()
-end
-
-function logsumexp(x::Array{Float64, 1})
-  top = maximum(x)
-  y = exp(x - top)
-  return log(sum(y)) + top
-end
-
-function logdiffexp(a::Float64, b::Float64)
-  return b + log(exp(a - b) - 1.)
-end
-

### 2015-12-04 Code/OldC++/display_atoms.py

@@ -1,19 +1,0 @@
-# Display file for Gravity
-from pylab import *
-
-output = atleast_2d(loadtxt('sample.txt'))
-
-ion()
-hold(False)
-for i in xrange(0, output.shape[0]):
-	model = output[i, 1:]
-	x, y = model[0:50], model[100:150]
-	plot(x, y, 'k.')
-	axis('scaled')
-	axis([0, 1, 0, 1])
-	title(i+1)
-	draw()
-
-ioff()
-show()
-

### 2015-12-04 Code/OldC++/display_gravity.py

@@ -1,24 +1,0 @@
-# Display file for Gravity
-from pylab import *
-
-output = atleast_2d(loadtxt('sample.txt'))
-
-figure(figsize=(13, 6))
-ion()
-hold(False)
-for i in xrange(0, output.shape[0]):
-	model = output[i, 0:-1]
-	x, y = model[0:1000], model[1000:2000]
-	vx, vy = model[3000:4000], model[4000:5000]
-	subplot(1,2,1)
-	plot(x, y, 'k.')
-	axis([-10, 10, -10, 10])
-	subplot(1,2,2)
-	plot(vx, vy, 'k.')
-	axis([-10, 10, -10, 10])
-	title(i+1)
-	draw()
-
-ioff()
-show()
-

### 2015-12-04 Code/OldC++/display_imageentropy.py

@@ -1,16 +1,0 @@
-# Display file for Potts
-from pylab import *
-
-output = atleast_2d(loadtxt('sample.txt'))
-
-ion()
-hold(False)
-for i in xrange(0, output.shape[0]):
-	x = output[i, 1:-1].reshape((100, 100))
-	imshow(x, interpolation='nearest', cmap='gray')
-	title(str(i+1) + '/' + str(output.shape[0]))
-	print(output[i, -1])
-	draw()
-ioff()
-show()
-

### 2015-12-04 Code/OldC++/display_potts.py

@@ -1,15 +1,0 @@
-# Display file for Potts
-from pylab import *
-
-output = atleast_2d(loadtxt('sample.txt'))
-
-ion()
-hold(False)
-for i in xrange(0, output.shape[0]):
-	x = output[i, 1:-2].reshape((50, 50))
-	imshow(x, interpolation='nearest')
-	title(str(i+1) + '/' + str(output.shape[0]))
-	draw()
-ioff()
-show()
-

### 2015-12-04 Code/OldC++/main.cpp

@@ -1,31 +1,0 @@
-#include <iostream>
-#include <ctime>
-#include <cstdlib>
-#include <RandomNumberGenerator.h>
-#include "Sampler.h"
-#include "Models/SimpleExample.h"
-
-using namespace std;
-using namespace DNest3;
-
-int main()
-{
-	// Initialise RNG and seed with time
-	RandomNumberGenerator::initialise_instance();
-	RandomNumberGenerator::get_instance().set_seed(time(0));
-
-//	ImageEntropy::load_data();
-//	Sampler<ImageEntropy> sampler(8, 2, 1000, 0.5, 10);
-
-	Sampler<SimpleExample> sampler(8, 1000, 5000, 0.001, 1000);
-	sampler.initialise();
-
-	while(true)
-	{
-		sampler.explore();
-		sampler.refresh();
-	}
-
-	return 0;
-}
-

### 2015-12-04 Code/OldC++/scenarios.py

@@ -1,58 +1,0 @@
-from pylab import *
-import pylab
-
-logw = loadtxt('logw.txt')
-scalars = loadtxt('scalars.txt')
-
-plot(scalars[:,0], scalars[:,1], 'bo', alpha=0.1)
-show()
-
-# Truncate so we only use completed runs
-steps = 200
-runs = logw.size//steps
-
-scalars = scalars[0:runs*steps, :]
-
-# Generate logX values that have a valid ordering.
-# Strictly decreasing within each run, and any point
-# with BOTH scalars higher than another point must have
-# lower logX.
-
-logX = zeros(scalars.shape[0])
-
-# Make an arbitrary assignment for run 1
-for i in xrange(0, steps):
-  logX[i] = -(i+1)
-
-# Go through other runs
-for i in xrange(steps, runs*steps):
-  lower_limit, upper_limit = -1E300, 0.
-
-  if i % steps != 0:
-    upper_limit = logX[i-1]
-
-  # Go through all assigned values in previous runs, and find
-  # all distributions known to be exterior to the current one
-  # we're trying to assign. If any have an assignment below
-  # upper_limit, lower the upper_limit.
-  for j in xrange(0, (i//steps)*steps):
-    # If exterior
-    if pylab.all(scalars[j, :] < scalars[i, :]) and logX[j] < upper_limit:
-      upper_limit = logX[j]
-
-  # Do the same for the lower limit. Look for any distributions
-  # known to be interior to the current point. Raise lower_limit
-  # if necessary.
-  for j in xrange(0, (i//steps)*steps):
-    # If exterior
-    if pylab.all(scalars[j, :] > scalars[i, :]) and logX[j] > lower_limit:
-      lower_limit = logX[j]
-
-  if lower_limit == -1E300:
-    logX[i] = logX[0:i].min() - 1
-  else:
-    logX[i] = 0.5*(lower_limit + upper_limit)
-
-plot(logX, scalars[:,0], 'bo')
-show()
-

### 2015-12-04 Code/OldC++/showresults.py

@@ -1,132 +1,0 @@
-from pylab import *
-import scipy
-from postprocess import logsumexp
-
-# Temperatures
-T1, T2 = 0.1, 1.
-
-# First calculate things about the scalars (e.g. the normalising constant)
-output = loadtxt('output.txt')
-scalars = output[:,1:]
-logw = output[:,0]
-smallest = min([scalars.shape[0], logw.size])
-scalars = scalars[0:smallest, :]
-logw = logw[0:smallest]
-
-# Prior weights, normalised
-logw = logw - logsumexp(logw)
-
-# Evaluate normalising constant at any temperature
-def canonical_properties(temperature1, temperature2):
-  # Posterior weights, unnormalised
-  logW = logw + scalars[:,0]/temperature1 + scalars[:,1]/temperature2
-
-  # Normaliser
-  logZ = logsumexp(logW)
-
-  # Posterior weights, normalised
-  logWW = logW - logZ
-  WW = exp(logWW)
-  ess = exp(-sum(WW*logWW))
-
-  # Information
-  H = sum(WW*(logWW - logw))
-
-  # Expected values of scalars
-  exp1 = sum(WW*scalars[:,0])
-  exp2 = sum(WW*scalars[:,1])
-
-  return [logZ, H, logWW, ess, exp1, exp2]
-
-[logZ, H, logWW, ess, exp1, exp2] = canonical_properties(T1, T2)
-
-print('log(Z) = {logZ}'.format(logZ=logZ))
-print('H = {H} nats'.format(H=H))
-
-plot(exp(logWW))
-ylabel('Weight wrt canonical distribution')
-title('ESS (for purposes of normalising constant calc) = {ess}'.format(ess=ess))
-show()
-
-# Resample to uniform weight
-N = int(ess)
-posterior_sample = zeros((N, scalars.shape[1]))
-w = exp(logWW)/max(exp(logWW))
-for i in xrange(0, N):
-  while True:
-    which = randint(scalars.shape[0])
-    if rand() <= w[which]:
-      break
-  posterior_sample[i,:] = scalars[which,:]
-
-plot(posterior_sample[:,0], posterior_sample[:,1], 'b.', markersize=5, alpha=0.5)
-hold(True)
-plot(scalars[:,0], scalars[:,1], 'r.', markersize=1, alpha=0.2)
[…]

### 2015-12-04 Code/OldC++/truth.py

@@ -1,58 +1,0 @@
-from pylab import *
-
-x = linspace(0., 1., 2001)
-
-def truth(T1, T2, do_plot=False):
-  p = exp(-(x - 0.5)**2/T1 - sin(2.*pi*x/0.5)**2/T2)
-  Z = trapz(p, x=x)
-  H = trapz(p/Z*log(p/Z), x=x)
-
-  logZ = 100*log(Z)
-  H *= 100
-
-  if do_plot:
-    plot(x, p)
-    title([logZ, H])
-    show()
-
-  return [logZ, H]
-  
-
-def grid():
-  # Calculate log(Z) and H for some canonical distributions
-  T1 = 10.**(linspace(-2.5, 5., 51))
-  T2 = T1.copy()
-  [T1, T2] = meshgrid(T1, T2)
-  T2 = T2[::-1, :]
-  logZ = T1.copy()
-  H = T1.copy()
-
-  for i in xrange(0, 51):
-    for j in xrange(0, 51):
-      [logZ[i, j], H[i, j]] = truth(T1[i, j], T2[i, j])
-    print(i+1)
-
-  figure(1)
-  subplot(1, 2, 1)
-  imshow(logZ, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest')
-  xlabel('ln($T_1$)')
-  ylabel('ln($T_2$)')
-  title('log(Z)')
-
-  subplot(1, 2, 2)
-  imshow(H, extent=(log(T1.min()), log(T1.max()), log(T2.min()), log(T2.max())), interpolation='nearest')
-  xlabel('ln($T_1$)')
-  ylabel('ln($T_2$)')
-  title('H, max={m}'.format(m=H.max()))
-
-  show()
-  return [logZ, H]
-
-if __name__ == '__main__':
-  truth(0.1, 1., do_plot=True)
-  [logZ, H] = grid()
-
-  savetxt('true_logZ.txt', logZ)
-  savetxt('true_H.txt', H)
-
-

### 2015-12-04 Code/OldC++/use_jags.r

@@ -1,86 +1,0 @@
-model = "model
-{
-  # Coefficients
-  A ~ dunif(-1000, 1000)
-  B ~ dunif(-1000, 1000)
-  C ~ dunif(-1000, 1000)
-
-  for(i in 1:N)
-  {
-    log_sigma[i] <- A*log(walkers[i]) + B*log(reps[i]) + C
-    sigma[i] <- exp(log_sigma[i])
-    # log(Z) as output variable
-    y[i] ~ dnorm(-209.053, 1/sigma[i]^2)
-    # H as output variable
-    #y[i] ~ dnorm(48.4834, 1/sigma[i]^2)
-  }
-
-  # Predict the results of an accurate run
-  log_sigma_new <- A*log(100) + B*log(1000) + C
-  sigma_new <- exp(log_sigma_new)
-  y_new ~ dnorm(-209.053, 1/sigma_new^2)
-}
-"
-
-data = as.matrix(read.table('results.txt'))
-
-# Cull runs with < 3 walkers or reps
-#data = data[data[,1] >= 3 & data[,2] >= 3, ]
-
-# log(Z) as output variable
-data = list(walkers=data[,1], reps=data[,2], y=data[,4], N=length(data[,4]))
-# H as output variable
-# data = list(walkers=data[,1], reps=data[,2], y=data[,5], N=length(data[,4]))
-
-# Variables to monitor
-variable_names = c('A', 'B', 'C', 'y_new')
-
-# How many burn-in steps?
-burn_in = 1000
-
-# How many proper steps?
-steps = 100000
-
-# Thinning?
-thin = 10
-
-# Random number seed
-seed = 42
-
-
-# NO NEED TO EDIT PAST HERE!!!
-# Just run it all and use the results list.
-
-library('rjags')
-
-# Write model out to file
-fileConn=file("model.temp")
-writeLines(model, fileConn)
-close(fileConn)
-
-if(all(is.na(data)))
-{
-	m = jags.model(file="model.temp", inits=list(.RNG.seed=seed, .RNG.name="base::Mersenne-Twister"))
-} else
-{
-	m = jags.model(file="model.temp", data=data, inits=list(.RNG.seed=seed, .RNG.name="base::Mersenne-Twister"))
-}
-update(m, burn_in)
-draw = jags.samples(m, steps, thin=thin, variable.names = variable_names)
-# Convert to a list
-make_list <- function(draw)
-{
-	results = list()
-	for(name in names(draw))
[…]

### 2015-12-04 Code/Python/Particle.py

@@ -1,65 +1,0 @@
-import numpy as np
-import numpy.random as rng
-
-class Particle:
-	"""
-	Class defining a particle, and the scalar functions
-	(log-likelihood or -energy) that you want to increase
-	"""
-	def __init__(self):
-		"""
-		Constructor: does nothing.
-		"""
-		pass
-
-	def from_prior(self):
-		"""
-		Generate coordinates/parameters from the prior
-		"""
-		self.x = rng.rand(100)
-
-	def proposal(self):
-		"""
-		Do something to generate a Metropolis proposal
-		"""
-		# Choose a probability from log p ~ uniform
-		p = 10.**(-3.*rng.rand())
-
-		# Move each coordinate with probability p
-		which = rng.rand(len(self.x)) < p
-
-		# Make sure we're at least moving one
-		num_moving = which.sum()
-		if num_moving == 0:
-			which[rng.randint(len(self.x))] = True
-			num_moving = 1
-
-		# Move using a heavy-tailed proposal
-		self.x[which] += 10.**(1.5 - 6.*rng.rand(num_moving))\
-								*rng.randn(num_moving)
-
-		# Mod into prior range
-		self.x = np.mod(self.x, 1.)
-
-		# Return zero -- in general return log of
-		# a factor that enforces detailed balance
-		# wrt the prior.
-		return 0.
-
-	@property
-	def scalars(self):
-		"""
-		Calculate the scalar functions of interest
-		and return them in the form of a numpy array.
-		Current demo problem: the first scalar wants
-		the coordinates to cluster in a gaussian in
-		the center of the domain, and the second wants
-		them to have periodic density. The coordinates
-		are IID in this problem so it's easy to numerically
-		compute the true partition function.
-		"""
-		s = np.empty(2)
-		s[0] = -np.sum((self.x - 0.5)**2)
-		s[1] = -np.sum(np.sin(4.*np.pi*self.x)**2)
-		return s
-

### 2015-12-04 Code/Python/Sampler.py

@@ -1,164 +1,0 @@
-import numpy as np
-import numpy.random as rng
-import matplotlib.pyplot as plt
-from copy import deepcopy
-
-from Particle import Particle
-
-class Sampler:
-	"""
-	An object of this class is a sampler.
-	"""
-
-	def __init__(self, num_particles):
-		"""
-		Constructor: pass in the number of particles
-		"""
-		self.num_particles = num_particles
-		self.particles = [Particle() for i in range(0, num_particles)]
-
-		# Open and close output file to clear it
-		f = open('output.txt', 'w')
-		f.close()
-
-		# Rectangles that have been forbidden so far
-		self.forbidden_rectangles = None
-
-		# Fraction of prior mass still in play
-		self.log_prior_mass = 0.
-
-	def initialise(self):
-		"""
-		Generate all the particles from the prior
-		"""
-		self.all_scalars = []	# Values of scalars for all particles
-		for particle in self.particles:
-			particle.from_prior()
-			self.all_scalars.append(particle.scalars)
-		self.all_scalars = np.array(self.all_scalars)
-		self.forbidden_rectangles = np.empty((0, self.all_scalars.shape[1]))
-		self.iteration = 0
-		return
-
-	def do_iteration(self):
-		"""
-		Do an NS iteration
-		"""
-		counts = self.corner_counts
-		ii = 
-		counts = counts[counts != 0]
-		if len(counts) == 0:
-			exit()
-
-		# Choose one with minimum nonzero count
-		temp = np.nonzero(counts == counts.min())[0]
-		choice = temp[rng.randint(len(temp))]
-
-		# Choose one of the two scalars and find its minimum within
-		# particle[choice]'s rectangle
-		which_scalar = rng.randint(2)
-		inside = [Sampler.is_in_rectangle(self.all_scalars, self.all_scalars[choice, :]) \
-						for i in range(self.num_particles)]
-		inside = np.nonzero(inside)[0]
-		print(inside)
-		ss = self.all_scalars[inside, which_scalar]
-		which = np.nonzero(ss==ss.min())[0]
-
-		forbid = self.all_scalars[choice, :].copy()
-		forbid[which_scalar] = self.all_scalars[which, which_scalar]
-
-		# Estimate the fraction of *remaining* prior mass being eliminated
-		frac = float(1 + counts[which])/self.num_particles
-
[…]

### 2015-12-04 Code/Python/Utils.py

@@ -1,19 +1,0 @@
-"""
-Various useful functions
-"""
-
-import numpy as np
-
-def logsumexp(values):
-	biggest = np.max(values)
-	x = values - biggest
-	result = np.log(np.sum(np.exp(x))) + biggest
-	return result
-
-def logdiffexp(x1, x2):
-	biggest = x1
-	xx1 = x1 - biggest
-	xx2 = x2 - biggest
-	result = np.log(np.exp(xx1) - np.exp(xx2)) + biggest
-	return result
-

### 2015-12-04 Code/Python/logZ.py

@@ -1,28 +1,0 @@
-"""
-Load output.txt and compute logZ
-"""
-
-import numpy as np
-from Utils import logsumexp
-
-# Load the data
-output = np.loadtxt('output.txt')
-
-def logZ(temperatures):
-	"""
-	Compute the normalising constant. Only works for two scalars.
-	"""
-	return logsumexp(output[:,0] + output[:,1]/temperatures[0] + output[:,2]/temperatures[1])
-
-def H(temperatures):
-	"""
-	Compute the information. Only works for two scalars.
-	"""
-	# Log posterior weights (normalised)
-	logp = output[:,0] + output[:,1]/temperatures[0] + output[:,2]/temperatures[1]\
-				- logZ(temperatures)
-	return np.sum(np.exp(logp)*(logp - output[:,0]))
-
-print('log(Z) = {a}'.format(a=logZ([0.1, 1.])))
-print('H = {h}'.format(h=H([0.1, 1.])))
-

### 2015-12-07 Code/C++/Sampler.h
Fixed oopsies (copy particles to use for initial conditions)

@@ -23,6 +23,10 @@ class Sampler
 		std::vector<MyModel> particles;
 		std::vector< std::vector<ScalarType> > scalars;
 
+		// Backup
+		std::vector<MyModel> backup_particles;
+		std::vector< std::vector<ScalarType> > backup_scalars;
+
 		// Forbidden rectangles
 		std::list< std::vector<ScalarType> > rects;
 

### 2015-12-07 Code/C++/SamplerImpl.h

@@ -208,6 +208,10 @@ void Sampler<MyModel>::do_iteration()
 	// Replace dead particles
 	int num_threads = rngs.size();
 
+	// Backup
+	backup_particles = particles;
+	backup_scalars = scalars;
+
 	// Assign particles to threads
 	std::vector< std::vector<int> > which_particles(num_threads);
 	int k=0;
@@ -262,11 +266,11 @@ int Sampler<MyModel>::refresh_particle(int which, int which_rng)
 	{
 		copy = rngs[which_rng].rand_int(num_particles);
 	}
-	while(!is_okay(scalars[copy]));
+	while(!is_okay(backup_scalars[copy]));
 
 	// Clone it
-	particles[which] = particles[copy];
-	scalars[which] = scalars[copy];
+	particles[which] = backup_particles[copy];
+	scalars[which] = backup_scalars[copy];
 
 	// Do the MCMC
 	MyModel proposal;

### 2015-12-14 Code/C++/Sampler.h
Fixed problems by distinguishing interior, boundary, and exterior particles

@@ -23,12 +23,12 @@ class Sampler
 		std::vector<MyModel> particles;
 		std::vector< std::vector<ScalarType> > scalars;
 
+		// Particle status
+		std::vector<int> status;
+
 		// Backup
 		std::vector<MyModel> backup_particles;
 		std::vector< std::vector<ScalarType> > backup_scalars;
-
-		// Dying flags
-		std::vector<bool> dying;
 
 		// Forbidden rectangles
 		std::list< std::vector<ScalarType> > rects;

### 2015-12-14 Code/C++/SamplerImpl.h

@@ -16,7 +16,6 @@ Sampler<MyModel>::Sampler(const std::vector<RNG>& rngs, int num_particles,
 ,num_particles(num_particles)
 ,particles(num_particles)
 ,scalars(num_particles)
-,dying(num_particles)
 ,mcmc_steps(mcmc_steps)
 ,saves_per_iteration(saves_per_iteration)
 ,initialised(false)
@@ -120,21 +119,46 @@ void Sampler<MyModel>::do_iteration()
 	// Make a ucc threshold (particles on the threshold die too)
 	int threshold = particle_uccs_sorted[num_particles/2];
 
-	// Assign particles to die
-	dying.assign(num_particles, false);
-	int num_dying = 0;
+	// Place forbidding rectangles anywhere ucc >= threshold
 	for(int i=0; i<num_particles; i++)
 	{
-		if(particle_uccs[i] >= threshold)
+		int j = num_particles-1;
+		while(j > 0 && ucc[i][j] < threshold)
+			j--;
+
+		if(ucc[i][j] >= threshold)
 		{
-			dying[i] = true;
-			num_dying++;
+			std::vector<ScalarType> latest{s1[j], s2[num_particles-i-1]};
+			prune_rectangles(latest);
+			rects.push_front(latest);
 		}
 	}
 
+	// -1 <===> interior
+	//  0 <===> boundary
+	// +1 <===> exterior
+	status.assign(num_particles, 0);
+	int num_interior = 0;
+	int num_exterior = 0;
+	int num_boundary = 0;
+	for(int i=0; i<num_particles; i++)
+	{
+		if(particle_uccs[i] > threshold)
+		{
+			status[i] = -1;
+			++num_interior;
+		}
+		if(particle_uccs[i] < threshold)
+		{
+			status[i] = +1;
+			++num_exterior;
+		}
+	}
+	num_boundary = num_particles - (num_interior + num_exterior);
+
 	// Select some particles to save in their entirety
 	std::vector<bool> save(num_particles, false);
-	if(num_dying != 0)
+	if(num_interior != 0)
 	{
 		for(int i=0; i<saves_per_iteration; i++)
 		{
@@ -142,18 +166,18 @@ void Sampler<MyModel>::do_iteration()
 			do
 			{
 				ii = rngs[0].rand_int(num_particles);
-			}while(!dying[ii]);
+			}while(status[ii] != -1);
 			save[ii] = true;
 		}
[…]

### 2015-12-14 Code/C++/main.cpp

@@ -22,8 +22,8 @@ int main(int argc, char** argv)
 	for(RNG& r: rngs)
 		r.set_seed(++seed);
 
-	constexpr int num_particles = 3001;
-	constexpr int num_mcmc_steps = 2000;
+	constexpr int num_particles = 1001;
+	constexpr int num_mcmc_steps = 1000;
 	constexpr double depth = 1000.;
 	constexpr int steps = depth/log(2.);
 

### 2015-12-14 Paper/figures/joint.py
classify particles

@@ -8,7 +8,7 @@ Make a plot of \pi(L1, L2)
 Based on a similar plot from ABCNS
 """
 
-rng.seed(3)
+rng.seed(1)
 rc("font", size=18, family="serif", serif="Computer Sans")
 rc("text", usetex=True)
 
@@ -20,7 +20,7 @@ f = np.exp(-0.5*(x-1.5)**2/1.**2)*np.exp(-0.5*((y - 5*(x/5)**2)**2)/1.**2)
 f /= f.sum()
 
 # Generate samples
-M = 40
+M = 41
 xx = 1.5 + 1.*rng.randn(M)
 yy = 5*(xx/5)**2 + 1.*rng.randn(M)
 keep = (xx > 0.) & (xx < 5.) & (yy > 0.) & (yy < 5.)
@@ -36,6 +36,7 @@ plt.fill_between(x[0, :][x[0, :] > 2.], 1.5, 5., color=[0, 0, 0], alpha=0.1)
 plt.savefig('joint1.pdf', bbox_inches='tight')
 plt.show()
 
+
 ucc = np.zeros((N, N))
 for i in range(N):
 	for j in range(N):
@@ -45,9 +46,23 @@ for i in range(N):
 				ucc[i, j] += 1
 Xhat = ucc/len(xx)
 
+particle_uccs = np.zeros(len(xx))
+for i in range(len(xx)):
+	particle_uccs[i] = 0
+	for k in range(len(xx)):
+		if (xx[k] >= xx[i]) and (yy[k] >= yy[i]):
+			particle_uccs[i] += 1
+
+boundary = particle_uccs == np.median(particle_uccs)
+interior = particle_uccs > np.median(particle_uccs)
+exterior = particle_uccs < np.median(particle_uccs)
+
 plt.imshow(f, extent=[x.min(), x.max(), y.min(), y.max()], cmap='Blues')
 plt.imshow(-(Xhat>0.25), extent=[x.min(), x.max(), y.min(), y.max()], cmap='gray', alpha=0.2)
-plt.plot(xx, yy, 'ko')
+plt.plot(xx[exterior], yy[exterior], 'go', markersize=6, label='Exterior')
+plt.plot(xx[boundary], yy[boundary], 'y*', markersize=8, label='Boundary')
+plt.plot(xx[interior], yy[interior], 'rs', markersize=6, label='Interior')
[…]

### 2015-12-14 Paper/ms.tex
Polson and Scott
geq

@@
 For any values $(\ell_1, \ell_2)$ of the objective functions,
-the prior mass for which both $L_1 > \ell_1$ {\em and} $L_2 > \ell_2$ is:
+the prior mass for which both $L_1 \geq \ell_1$ {\em and} $L_2 \geq \ell_2$ is:
 \begin{eqnarray}
 X(\ell_1, \ell_2) &=& \int \pi(L_1, L_2)
-\mathds{1}\left[L_1 > \ell_1, L_2 > \ell_2 \right]
+\mathds{1}\left[L_1 \geq \ell_1, L_2 \geq \ell_2 \right]
  \, dL_1 \, dL_2.
@@
 \frac{1}{N}
-\sum_{i=1}^N \mathds{1}\left[L_1(\xx_i) > \ell_1,
-L_2(\xx_i) > \ell_2\right].\label{eqn:corner_count}
+\sum_{i=1}^N \mathds{1}\left[L_1(\xx_i) \geq \ell_1,
+L_2(\xx_i) \geq \ell_2\right].\label{eqn:corner_count}
 \end{eqnarray}
@@
 \end{figure}
+
+Try to relate this to \citet{vertical}.
 
@@
 
+\bibitem[Polson and Scott(2014)]{vertical}
+Polson, Nicholas G., and James G. Scott. ``Vertical-likelihood Monte Carlo
+Integration.'' arXiv preprint arXiv:1409.3601 (2014).
+
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.

### 2015-12-14 Paper/ms.tex
update algorithm

@@
 	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$, where $k = (N-1)/2$.
-	\While{$k > 1$ and $\hat{X}_{(k-1)} = \hat{X}_{(k)}$}
-		\State Set $k \leftarrow (k-1)$.
-	\EndWhile
-	\State Save all particles with $\hat{X} \geq \hat{X}_{(k)}$ to disk,
-			assigning each a prior weight of $M/N$.
+	\State Classify all particles as either ``interior'', ``boundary'', or
+			``exterior'', based on whether their $\hat{X}$ values are
+			greater than, equal to, or less than $\hat{X}_{(k)}$ respectively.
+	\State Save all ``interior'' particles to disk, assigning each a prior weight 				of $n_{\rm interior}/(n_{\rm interior} + n_{\rm exterior})$.
 	\State Exclude the region in the $L_1$-$L_2$ plane where
@@
 	\State \hspace{20pt} $\mathcal{R} \leftarrow \mathcal{R} \cup \{(\ell_1, \ell_2) \in \mathbb{R}^2: \hat{X}(\ell_1, \ell_2) \geq \hat{X}_{(k)}\}$.
-	\State Update the remaining prior mass to $M \leftarrow M(1 - k/N)$.
-	\State Replace all particles with $\hat{X} \geq \hat{X}_{(k)}$ by new
+	\State Update the remaining prior mass to $M \leftarrow M n_{\rm exterior}/N$.
+	\State Replace all interior and boundary particles by new
 			particles generated from $\pi(\xx)$, subject to the constraint

### 2016-01-05 Code/C++/showresults.py
New results

@@ -134,6 +134,7 @@ good = logical_not(isnan(err1))
 imshow(err1, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', vmin=-(abs(err1[good])).max(), vmax=(abs(err1[good])).max(), cmap='coolwarm')
 print("H_max = " + str(H[good].max()))
 print(err1[good].min(), err1[good].max())
+print(sqrt(mean(err1[good]**2)))
 print(4.*H[good].max()/(abs(err1[good]).max())**2)
 
 subplot(1,2,2)
@@ -142,5 +143,4 @@ good = logical_not(isnan(err1))
 imshow(err2, extent=(log10(T1.min()), log10(T1.max()), log10(T2.min()), log10(T2.max())), interpolation='nearest', vmin=-(abs(err2[good])).max(), vmax=(abs(err2[good])).max(), cmap='coolwarm')
 print(err2[good].min(), err2[good].max())
 show()
-
 

### 2016-01-05 Paper/ms.tex
Extra writing

@@
 \abstract{We present a generalization of the Nested Sampling algorithm for
-estimating the partition function of a family of
-probability distributions indexed by two hyperparameters. These
-problems arise frequently in statistical mechanics.}
+estimating the partition function $Z(\beta_1, \beta_2)$ of a family of
+probability distributions indexed by two hyperparameters $\beta_1$ and
+$\beta_2$. These
+problems arise frequently in statistical mechanics and occasionally in
+Bayesian inference. We demonstrate the algorithm on a toy example where
+$Z(\beta_1, \beta_2)$ is known, and show empirically that the accuracy
+improves as XX as a function of the number of particles $N$. We also
+calculate the partition function for a Lennard-Jones system in an external
+field, as a function of the external field strength, in a single run.}
 
@@
 \citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
+Its key strength its ability to cope with phase transitions which cause
+many other methods to fail.
 In a Bayesian inference framework with unknown parameters denoted collectively
@@
 \end{eqnarray}
+
+$Z$ is useful because it is the probability of the data given $\xx$ is one of
+the possibilities in the space being integrated over, and thus plays the role
+of a likelihood for the parameter space as a whole.
 The main goal of NS is to approximate $Z$, but it can also be used to make
-Monte Carlo approximations of the posterior, and calculate the ``information'',
+Monte Carlo approximations of the posterior distribution, and to
+calculate the `information',
 or Kullback-Leibler divergence from the prior to the posterior:
 \begin{eqnarray}
-\mathcal{I} &=& \int p(\xx | \data, \II) \ln
+\mathcal{H} &=& \int p(\xx | \data, \II) \ln
 \left[\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right] \, d\xx \\
@@
 \end{eqnarray}
-$\mathcal{I}$ quantifies how compressed the posterior distribution is with
-respect to the prior. For example, $\mathcal{I} = 10$ implies the posterior
-occupies a fraction $e^{-10}$ of the prior mass.
+$\mathcal{H}$ quantifies how compressed the posterior distribution is with
+respect to the prior. For example, $\mathcal{H} = 10$ implies (loosely speaking)
+that the posterior occupies a fraction $e^{-10}$ of the prior mass.
 
@@
 p(\xx; \ell) &=& \frac{1}{X(\ell)}\pi(\xx)\mathds{1}\left(L(\xx) > \ell\right).
+\label{eq:constrained_prior}
 \end{eqnarray}
@@
 
-In addition to the marginal likelihood $Z$, NS also allows us
-to approximate properties of the posterior $\pi(\xx)L(\xx)/Z$.
+Naively, Equation~\ref{eqn:evidence} appears straightforward to compute because
+it is a simple expectation of $L$ with respect to the distribution $\pi$.
+However,
 NS is required because Equation~\ref{eqn:evidence}
@@
 (the prior for the $L$ value implied by $\pi(\xx)$). Hence, there is a strong
-connection between NS and ideas from rare event simulation \citep{walter}.
+connection between NS and ideas from rare event simulation, which has been
+explicitly investigated by \citet{walter}.
+
 Using a single NS run, we can also calculate the properties of any other distribution that
@@
 computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
-posterior distribution would have been if the noise variance had been greater.
+posterior distribution would have been if the noise variance had been greater,
+without having to re-run the algorithm.
 This is different from including an extra parameter to allow the noise variance
@@
 However, unlike annealing methods, NS requires only a small number of
-tuning parameters, and still works on problems involvin phase transitions.
+tuning parameters (just the number of particles $N$ is inherent to NS, although
+most NS implementations also have some tuning parameters related to their
+methods for sampling Equation~\ref{eq:constrained_prior}). Annealing methods
+also tend to fail on phase change problems \citep{skilling} which can arise
+in data analysis \citep{rjobject, exoplanet}.
 
@@
 
+\bibitem[Brewer and Donovan(2015)]{exoplanet} Brewer, B.~J., 
+Donovan, C.~P.\ 2015.\ Fast Bayesian inference for exoplanet discovery in 
+radial velocity data.\ Monthly Notices of the Royal Astronomical Society 
+448, 3206-3214.
+
+\bibitem[Brewer(2014)]{rjobject} Brewer, B.~J.\ 2014.\ Inference 
+for Trans-dimensional Bayesian Models with Diffusive Nested Sampling.\ 
+ArXiv e-prints arXiv:1411.3921.
+
 \bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,

### 2016-01-06 Paper/ms.tex
Improving descriptions of the algorithm and its motivation

@@
 \end{enumerate}
-The standard Nested Sampling algorithm, (as well as alternatives such as
-Diffusive Nested Sampling \citep{dnest}) has these properties but for a
-single objective function.
+The standard Nested Sampling algorithm has these properties but for a
+single objective function. The same is true of variants such as
+Diffusive Nested Sampling \citep{dnest}.
 
@@
 
-{\bf Here is where I'll attempt to justify the algorithm}.
-
-Consider the following proposal. Imagine running standard NS with $N$ particles
+Consider the following proposal, which we label the `turn-the-corner idea'.
+Imagine running standard NS with $N$ particles
 using the objective function $L_1(\xx)$.
@@
 Subject to the constraint $L_1(\xx) > L_1^*$, standard NS can continue but
-attempting to increase $L_2$ instead of $L_1$. After $M_2$ such iterations,
+attempting to increase $L_2$ instead of $L_1$. This is what we mean by
+`turning the corner'. After $M_2$ such iterations,
 the algorithm would have compressed the prior by a factor
@@
 of the function $X(L_1, L_2)$.
+
+Figure~\ref{fig:contours} shows the contour lines of $X(L_1, L_2)$ for the
+prior in Figure~\ref{fig:joint1}. Each contour line encloses half of the
+prior mass of the previous one. The fact that contours of $X(L_1, L_2)$ are
+relevant suggests that a standard NS run with objective function
+$X(L_1(\xx), L_2(\xx))$ would achieve a similar effect to the turn-the-corner
+idea.
 
@@
 
-Try to relate this to \citet{vertical}.
+{\bf Try to relate this to \citet{vertical}}.
 
@@
 \begin{eqnarray}
-Y(\xx) &=& -\ln\left[X\left(L_1(\xx), L_2(\xx)\right)\right].
+Y(\xx) &=& 1/X\left(L_1(\xx), L_2(\xx)\right).
 \end{eqnarray}
@@
 therefore cannot restrict the parameter space based on the value of $X$ (or
-equivalently $Y$. However, given a number of Monte Carlo samples, we
-can approximate this.
+equivalently, $Y$). However, given a number of Monte Carlo samples, we
+can {\em approximate} $X$ by $\hat{X}(L_1, L_2)$, and restrict the parameter
+space based on a contour of that approximation.
+The algorithm is presented in detail below.
 
@@
 \State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
-$N$ should be odd and quite large (1001 or greater).
+$N$ should quite large (typically 1000 or greater), and the algorithm is
+slightly more elegant if $N$ is odd.
 \State Initialise the {\em remaining prior mass} to be $M=1$.
@@
 			and sort the $\hat{X}$ values observed.
-	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$, where $k = (N-1)/2$.
+	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$.
 	\State Classify all particles as either ``interior'', ``boundary'', or
@@
 			greater than, equal to, or less than $\hat{X}_{(k)}$ respectively.
+	\State If the number of interior particles $n_{\rm interior}$ or the number
+			of exterior particles $n_{\rm exterior}$ is zero, replace $k$ by
+			$(k-1)$ (to try a percentile other than the median)
+			and go back to the previous step (if $k=1$ then set $k=N$ instead of zero).
 	\State Save all ``interior'' particles to disk, assigning each a prior weight of $n_{\rm interior}/(n_{\rm interior} + n_{\rm exterior})$.

### 2016-01-07 Paper/ms.tex
New references

@@
 \usepackage{algpseudocode}
+\usepackage{natbib}
 
@@
 of the function $X(L_1, L_2)$.
+We have explored the idea of using multiple turn-the-corner runs, with
+the corner being turned at different times (i.e. using different values of
+$M_1$), and combining the results using biased sampling
+\citep{vardi, cameron} in order to estimate $Z(\beta_1, \beta_2)$. However,
+this was very computationally intensive. We presented the turn-the-corner idea
+here in order to provide an intuitive rationale for our actual algorithm given in
+Section~\ref{sec:algorithm}.
 
@@
 
+\bibitem[Cameron and Pettitt(2014)]{cameron} Cameron, E., and Pettitt, A., 2014,
+``Recursive pathways to marginal likelihood estimation with
+prior-sensitivity analysis.'' Statistical Science 29.3 (2014): 397-419.
+
 \bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.
@@
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.
+
+\bibitem[Vardi (1985)]{vardi}
+Vardi, Yehuda. ``Empirical distributions in selection bias models.'', The Annals of Statistics (1985): 178-203.
 

### 2016-01-07 Paper/ms.tex
Numbers in the algorithm, and failure modes

@@
 \begin{algorithmic}
-\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
+\State 1) Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
 $N$ should quite large (typically 1000 or greater), and the algorithm is
 slightly more elegant if $N$ is odd.
-\State Initialise the {\em remaining prior mass} to be $M=1$.
-\State Initialise the {\em excluded region} $\mathcal{R}$ to be the empty set
+\State 2) Initialise the {\em remaining prior mass} to be $M=1$.
+\State 3) Initialise the {\em excluded region} $\mathcal{R}$ to be the empty set
 $\emptyset$.
 \While{more iterations desired}
-	\State Calculate $\hat{X}(\ell_1, \ell_2)$ at the position of each particle
+	\State i) Calculate $\hat{X}(\ell_1, \ell_2)$ at the position of each particle
 			and sort the $\hat{X}$ values observed.
-	\State Find the median $\hat{X}$ value, $\hat{X}_{(k)}$.
-	\State Classify all particles as either ``interior'', ``boundary'', or
+	\State ii) Find the median $\hat{X}$ value, $\hat{X}_{(k)}$.
+	\State iii) Classify all particles as either ``interior'', ``boundary'', or
 			``exterior'', based on whether their $\hat{X}$ values are
 			greater than, equal to, or less than $\hat{X}_{(k)}$ respectively.
-	\State If the number of interior particles $n_{\rm interior}$ or the number
+	\State iv) If the number of interior particles $n_{\rm interior}$ or the number
 			of exterior particles $n_{\rm exterior}$ is zero, replace $k$ by
 			$(k-1)$ (to try a percentile other than the median)
-			and go back to the previous step (if $k=1$ then set $k=N$ instead of zero).
-	\State Save all ``interior'' particles to disk, assigning each a prior weight of $n_{\rm interior}/(n_{\rm interior} + n_{\rm exterior})$.
-	\State Exclude the region in the $L_1$-$L_2$ plane where
+			and go to step iii) (if $k=1$ then set $k=N$ instead of zero).
+	\State v) Save all ``interior'' particles to disk, assigning each a prior weight of $n_{\rm interior}/(n_{\rm interior} + n_{\rm exterior})$.
+	\State vi) Exclude the region in the $L_1$-$L_2$ plane where
 $\hat{X} \geq \hat{X}_{(k)}$. That is, set
 	\State \hspace{20pt} $\mathcal{R} \leftarrow \mathcal{R} \cup \{(\ell_1, \ell_2) \in \mathbb{R}^2: \hat{X}(\ell_1, \ell_2) \geq \hat{X}_{(k)}\}$.
-	\State Update the remaining prior mass to $M \leftarrow M n_{\rm exterior}/N$.
-	\State Replace all interior and boundary particles by new
+	\State vii) Update the remaining prior mass to $M \leftarrow M n_{\rm exterior}/N$.
+	\State viii) Replace all interior and boundary particles by new
 			particles generated from $\pi(\xx)$, subject to the constraint
@@
 
+\section{Failure modes}
+
+If $L_1(\xx)$ and $L_2(\xx)$ are inversely and strongly dependent...
+
 %Use Skilling's argument about the prior distribution of a log Bayes factor.
@@
 \authorcontributions{Author Contributions}
+
+BJB initiated the project and explored various ideas which did not work.
 

### 2016-01-07 Paper/ms.tex
contributions
removed instructions

@@
 
-\section{Failure modes}
+\section{Failure modes}\label{sec:failure_modes}
 
@@
 BJB initiated the project and explored various ideas which did not work.
-
-Required if more than one author.
-Authorship must include and be strictly limited to researchers who have substantially contributed to the reported work.
-Please carefully review our criteria regarding the Qualification for Authorship: \web /instructions.
-
+EC worked on the biased sampling suggestion in Section~\ref{sec:rationale},
+discovered the failure mode in Section~\ref{sec:failure_modes}, ...
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-01-13 Paper/ms.tex
2016
Handley
Info theory probably better way for that argument

@@
 \pubvolume{xx}
-\pubyear{2015}
+\pubyear{2016}
 %\externaleditor{Academic Editor: xx}
@@
 \begin{eqnarray}
-\ln B = \ln Z_1 - \ln Z_2.
+\Delta := \ln B = \ln Z_1 - \ln Z_2.
 \end{eqnarray}
-Naively, we want $\ln B$ to an accuracy of about 0.1, so that we can reliably
+Naively, we want $\Delta$ to an accuracy of about 0.1, so that we can reliably
 quantify the strength of the evidence when the two models are similarly
 supported by the data. But is this a realistic situation?
-Prior to the data, we don't know the
-value of $B$. Rather, $B$ has a prior distribution:
+
+Prior to knowing the data $D$, we don't know how much information we will get
+about the question of $M_1$ vs. $M_2$, quantified by the KL divergence:
 \begin{eqnarray}
-p(B) &=& P(M_1)p(B|M_1) + P(M_2)p(B|M_2)\\
-&=& P(M_1)\int p(\xx_1, D, B | M_1)\,d\xx_1\,dD
-    + P(M_2)\int p(\xx_2, D, B | M_2)\,d\xx_2\,dD
+\mathcal{H} &=& \sum_{i=1}^2 P(M_i | D)
+\ln\left[
+\frac{P(M_i | D)}{P(M_i)}
+\right].
 \end{eqnarray}
 
-{\bf Need thought here}
-{\bf Thermodynamic integral might be helpful?}
+The expected value of $\mathcal{H}$ (with respect to the prior over datasets)
+is the mutual information:
 \begin{eqnarray}
-\ln\left[\int p(\theta)p(D|\theta) \, d\theta\right]
-&=& \int_0^1 \left<\ln p(D|\theta) \right>_\beta \, d\beta
+\mathcal{I} &=& \mathds{E}\left[\mathcal{H}\right]\\
+&=&
+\sum_{i=1}^2 \int P(M_i, D)
+\ln\left[
+\frac{P(M_i, D)}{P(M_i)P(D)}
+\right] \, dD.
 \end{eqnarray}
@@
 It is a pleasure to thank Gábor Csányi (Cambridge), Livia B. Pártay (Cambridge),
-and Robert Baldock (Cambridge) for valuable conversations.
+Robert Baldock (Cambridge), and Will Handley (Cambridge)
+for valuable conversations.
 

### 2016-01-27 Code/C++/canonical.py
Sample from canonical distributions

@@ -1,0 +1,78 @@
+"""
+Code to load results and resample to generate a sample
+from a canonical distribution.
+"""
+
+from pylab import *
+import scipy
+import colormaps as cmaps
+
+rc("font", size=18, family="serif", serif="Computer Sans")
+rc("text", usetex=True)
+
+def logsumexp(values):
+	biggest = np.max(values)
+	x = values - biggest
+	result = np.log(np.sum(np.exp(x))) + biggest
+	return result
+
+
+# Temperatures
+T1, T2 = 2., 10000.
+
+# First calculate things about the scalars (e.g. the normalising constant)
+sample = loadtxt('sample.txt')
+scalars = sample[:,1:3]
+logw = sample[:,0]
+smallest = min([scalars.shape[0], logw.size])
+scalars = scalars[0:smallest, :]
+logw = logw[0:smallest]
+
+# Prior weights, normalised
+logw = logw - logsumexp(logw)
+
+# Evaluate normalising constant at any temperature
+def canonical_properties(temperature1, temperature2):
+  # Posterior weights, unnormalised
+  logW = logw + scalars[:,0]/temperature1 + scalars[:,1]/temperature2
+
+  # Normaliser
+  logZ = logsumexp(logW)
+
+  # Posterior weights, normalised
+  logWW = logW - logZ
+  WW = exp(logWW)
+  ess = exp(-sum(WW*logWW))
+
+  # Information
+  H = sum(WW*(logWW - logw))
+
+  # Expected values of scalars
+  exp1 = sum(WW*scalars[:,0])
+  exp2 = sum(WW*scalars[:,1])
+  var1 = sum(WW*(scalars[:,0] - exp1)**2)
+  var2 = sum(WW*(scalars[:,1] - exp2)**2)
+
+  return [logZ, H, logWW, ess, exp1, exp2, var1, var2]
+
+[logZ, H, logWW, ess, exp1, exp2, var1, var2] = canonical_properties(T1, T2)
+WW = exp(logWW - logWW.max())
+
+plot(WW)
+ylabel('Weight wrt canonical distribution')
+title('ESS (for purposes of normalising constant calc) = {ess}'.format(ess=ess))
+show()
+
+# Resample to uniform weight
+N = int(ess)
+posterior_sample = zeros((N, sample.shape[1]))
+
+for i in range(0, N):
+	while True:
+		which = randint(sample.shape[0])
+		if rand() <= WW[which]:
+			break
[…]

### 2016-01-27 Paper/ms.tex
Rob

@@
 % Authors (Add full first names)
-\Author{Brendon J. Brewer$^{1,}$* and Ewan Cameron$^{2}$}
+\Author{Brendon J. Brewer$^{1,}$*, Ewan Cameron$^{2}$, and
+Robert J. N. Baldock$^{3}$}
 
@@
 $^{2}$ Spatial Ecology and Epidemiology Group, Tinbergen Building, Department
-of Zoology, University of Oxford, South Parks Road, Oxford, UK}
+of Zoology, University of Oxford, South Parks Road, Oxford, UK\\
+$^{3}$ Cavendish, University of Cambridge, Cambridge, UK}
 
@@
 
-\section{Test problem}
-
+\section{Example 1: Known solution}
 We tested the algorithm on a simple 100-dimensional example, where the
@@
 \end{figure}
+
+\section{Example 2: Lennard-Jones/Polymer Example}
+
+\section{Example 3: ``MaxEnt'' image deconvolution example}
 
@@
 It is a pleasure to thank Gábor Csányi (Cambridge), Livia B. Pártay (Cambridge),
-Robert Baldock (Cambridge), and Will Handley (Cambridge)
+and Will Handley (Cambridge)
 for valuable conversations.

### 2016-02-06 Paper/ms.tex
Image model

@@
 
+The unknown parameters are a set of pixel intensities
+$\{m_i\}_{i=1}^n$ where $n$ is the number of pixels and $m_i \geq 0$.
+The dataset is also a collection of pixel intensities $\{D_i\}_{i=1}^n$.
+The prior for the dataset given the parameters is a normal distribution
+with mean given by the underlying image $\{m_i\}$ convolved by
+a point spread function. The standard deviation of each pixel is
+$\sigma$, assumed known. This is usually written as
+
+\begin{eqnarray}
+\boldsymbol{D} = \boldsymbol{P}\boldsymbol{m} + \boldsymbol{n}
+\end{eqnarray}
+where $\boldsymbol{n}$ is the noise in each pixel and $\boldsymbol{P}$ is
+the blurring operator written in matrix form.
+
 \section{What kind of accuracy can we expect?}

### 2016-02-06 Paper/ms.tex
equations
murray

@@
 
+The posterior for the underlying image given the data is
+\begin{eqnarray}
+p(\boldsymbol{m} | \boldsymbol{D})
+&=& \frac{p(\boldsymbol{m})p(\boldsymbol{D}|\boldsymbol{m})}
+{\int p(\boldsymbol{m})p(\boldsymbol{D}|\boldsymbol{m}) \,d^n\boldsymbol{m}}\\
+&=& \frac{\exp\left[-\lambda \sum_i m_i\ln m_i
+-\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}m_{j}\right)^2\right]}
+{\int \exp\left[-\lambda \sum_i m_i\ln m_i
+-\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}m_{j}\right)^2\right]\, d^n\boldsymbol{m}}
+\end{eqnarray}
+
+Therefore, this is a `doubly intractable' problem, closely related to
+those studied (and solved) by \citep{murray}.
+
 \section{What kind of accuracy can we expect?}
@@
 
+\bibitem[Murray(2012)]{murray} Murray, Iain, Ghahramani, Zoubin, and MacKay, David.
+``MCMC for doubly-intractable distributions.'' arXiv preprint arXiv:1206.6848 (2012).
+
 \bibitem[P{\'a}rtay et al.(2014)]{2014PhRvE..89b2302P} P{\'a}rtay, L.~B., 

### 2016-02-06 Paper/figures/sweep.py
idea for an intuitive basis for the algorithm

@@ -1,0 +1,17 @@
+"""
+Experimenting with some animations to provide intuition about
+the TwinPeaks algorithm.
+\pi(X1, X2) ~ Uniform([0, 1]^2)
+Constraint on X1X2 < const
+const decreases.
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Let x=X1, y=X2|X1
+
+dt = 0.01
+
+for i in range(0, steps):
+

### 2016-02-08 Paper/ms.tex
Maxent talk

@@
 \section{Example 3: ``MaxEnt'' image deconvolution example}
+Problems with more than one objective function can also arise in data
+analysis. Consider, for example, the problem of deblurring a noisy
+image. A common approach to these problems is regularized inversion,
+such as the ``maximum entropy'' technique of \citet{gull}. It is important
+to distinguish between this method and the maximum entropy principle used
+to update or assign probabilities \citep{caticha}.
 
@@
 
+\bibitem[Caticha(2008)]{caticha} Caticha, A.\ 2008.\ Lectures 
+on Probability, Entropy, and Statistical Physics.\ ArXiv e-prints 
+arXiv:0808.0012.
+
+\bibitem[Gull and Daniell(1978)]{gull} Gull, Stephen F., and Daniell, G., 1978
+Image reconstruction from incomplete and noisy data. Nature (1978): 686-690.
+
 \bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.

### 2016-02-08 Paper/ms.tex
Cite gregory, also some tweaks in intro

@@
 
-$Z$ is useful because it is the probability of the data given $\xx$ is one of
+$Z$ is useful because it is the probability of the data given that
+$\xx$ is {\em one of}
 the possibilities in the space being integrated over, and thus plays the role
@@
 that the posterior occupies a fraction $e^{-10}$ of the prior mass.
+$\mathcal{H}$ can be interpreted quite literally as a measure of how much
+the data $\data$ told you about $\xx$.
 
@@
 same function as the inputs are not even necessarily of the same dimensionality.
-Therefore the symbol $L$ represents the output of the function.
+Therefore the symbol $L$ implies that the output of the function is a
+likelihood value; so $L(\xx)$ represents a likelihood value computed from
+the parameters $\xx$, and $L(X)$ represents a likelihood value that
+corresponds to an amount $X$ of prior mass.
 
@@
 There are many different methods based on this idea, such as the popular
-parallel tempering method \citep{pt}.
+parallel tempering method \citep{pt, gregory}.
 However, unlike annealing methods, NS requires only a small number of
@@
 
+\bibitem[\protect\citeauthoryear{Gregory}{2005}]{gregory}
+Gregory, Phil. Bayesian Logical Data Analysis for the Physical Sciences: A Comparative Approach with Mathematica® Support. Cambridge University Press, 2005.
+
 \bibitem[Gull and Daniell(1978)]{gull} Gull, Stephen F., and Daniell, G., 1978

### 2016-02-08 Paper/figures/sweep.py
save figures

@@ -16,7 +16,7 @@ def deriv(state):
 x = np.linspace(1E-6, 1, 1001)
 
 # Timestep
-t_final, dt = 10., 0.003
+t_final, dt = 10., 0.01
 steps = int(t_final/dt)
 
 # Initial condition for threshold
@@ -48,6 +48,7 @@ for i in range(0, steps):
 
 	plt.axhline(X, linestyle='--', color='g')
 	plt.axvline(X, linestyle='--', color='r')
+	plt.savefig('%0.4d' % (i + 1) + '.png', bbox_inches='tight')
 	plt.draw()
 
 # Finalise plotting

### 2016-02-08 Code/C++/Models/CambridgeLJ.cpp
Removed Cambridge model

@@ -1,79 +1,0 @@
-#include "CambridgeLJ.h"
-#include "Utils.h"
-#include <cmath>
-#include <cassert>
-
-using namespace std;
-
-// To call Rob's Fortran functions
-extern "C"
-{
-	void __twin_peaks_routines_MOD_initialise_config(double* s, double* h0,
-					double* V, double* ener, double* min_height,
-					double* Vmax, int* N, double* cutoff,
-					double* univar_3Nplus10001, double* univar_60k,
-					double* univar_30k, double* nvar_60k, bool* flat_v_prior);
-}
-
-CambridgeLJ::CambridgeLJ()
-:N(100)
-,s(0., 3*N)
-,scalars(2)
-{
-}
-
-void CambridgeLJ::from_prior(RNG& rng)
-{
-	// "Outputs" (along with s)
-	std::valarray<double> h0(9);
-	double V, ener;
-
-	// "constants"
-	double min_height = 0.1;
-	double Vmax = 10.;
-	double cutoff = 20.;
-	bool flat_v_prior = false;
-
-	// Random numbers
-	std::vector<double> univar_3Nplus10001(3*N+10001);
-	std::vector<double> univar_60k(60000);
-	std::vector<double> univar_30k(30000);
-	std::vector<double> nvar_60k(60000);
-	for(double& u: univar_3Nplus10001)
-		u = rng.rand();
-	for(double& u: univar_60k)
-		u = rng.rand();
-	for(double& u: univar_30k)
-		u = rng.rand();
-	for(double& n: nvar_60k)
-		n = rng.rand();
-
-	// Call Rob's code
-	__twin_peaks_routines_MOD_initialise_config(&s[0], &h0[0], &V,
-					&ener, &min_height, &Vmax, &N, &cutoff,
-					&univar_3Nplus10001[0], &univar_60k[0],
-					&univar_30k[0], &nvar_60k[0], &flat_v_prior);
-
-	compute_scalars();
-}
-
-void CambridgeLJ::compute_scalars()
-{
-	scalars[0] = 0.;
-	scalars[1] = 0.;
-}
-
-double CambridgeLJ::perturb(RNG& rng)
-{
-	double logH = 0.;
-	logH += rng.rand();
-
-	compute_scalars();
-	return 0.;
-}
-
-void CambridgeLJ::write_text(std::ostream& out) const
-{
-	out<<' ';
-}
-

### 2016-02-08 Code/C++/Models/CambridgeLJ.h

@@ -1,37 +1,0 @@
-#ifndef _CambridgeLJ_
-#define _CambridgeLJ_
-
-#include "RNG.h"
-#include <vector>
-#include <valarray>
-#include <ostream>
-
-class CambridgeLJ
-{
-	private:
-		int N;
-		std::valarray<double> s;
-		std::vector<double> scalars;
-
-		void compute_scalars();
-
-	public:
-		CambridgeLJ();
-
-		// Generate the point from the prior
-		void from_prior(RNG& rng);
-
-		// Metropolis-Hastings proposals
-		double perturb(RNG& rng);
-
-		void write_text(std::ostream& out) const;
-
-		// Getter
-		const std::vector<double>& get_scalars() const
-		{ return scalars; }
-
-		CambridgeLJ& operator = (const CambridgeLJ& other);
-};
-
-#endif
-

### 2016-02-08 Code/C++/display_simpleexample.py
xrange --> range

@@ -5,7 +5,7 @@ output = atleast_2d(loadtxt('sample.txt'))
 figure(figsize=(13, 6))
 ion()
 hold(False)
-for i in xrange(0, output.shape[0]):
+for i in range(0, output.shape[0]):
 	model = output[i, 3:]
 	hist(model, 300, alpha=0.2)
 	axis([0., 1., 0., 10.])

### 2016-02-09 Code/C++/canonical.py
Temperatures (1, 1) by default

@@ -18,7 +18,7 @@ def logsumexp(values):
 
 
 # Temperatures
-T1, T2 = 2., 10000.
+T1, T2 = 1., 1.
 
 # First calculate things about the scalars (e.g. the normalising constant)
 sample = loadtxt('sample.txt')

### 2016-02-09 Paper/ms.tex
Increased figure size
LJ potential

@@
 \centering
-\includegraphics[scale=0.75]{figures/output.png}
+\includegraphics[scale=1]{figures/output.png}
 \caption{The sequence of discarded particles (yellow) and
@@
 \section{Example 2: Lennard-Jones/Polymer Example}
+
+Consider $n$ Lennard-Jones atoms confined to the unit cube $[0, 1]^3$.
+The position of each atom is given by a vector
+$\boldsymbol{r}_i = (x_i, y_i, z_i)$.
+The potential energy of any configuration of the atoms is given by
+a sum over all pairs of atoms:
+\begin{eqnarray}
+U_{\rm LJ} &=& \sum_{i=1}^n\sum_{j=(i+1)}^n U_{\rm LJ}
+(\boldsymbol{r}_i, \boldsymbol{r}_j)
+\end{eqnarray}
+where the inter-atomic potential energy is given by:
+
+\begin{eqnarray}
+U_{\rm LJ}(\boldsymbol{r}_i, \boldsymbol{r}_j)
+&=& 4\epsilon
+\left[
+\left(\frac{\sigma}{|\boldsymbol{r}_i - \boldsymbol{r}_j|}\right)^12 -
+\left(\frac{\sigma}{|\boldsymbol{r}_i - \boldsymbol{r}_j|}\right)^6
+\right].
+\end{eqnarray}
+We adopted the Lennard Jones units by setting $\epsilon=1$ and $\sigma=1$.
 
@@
 
-Therefore, this is a `doubly intractable' problem, closely related to
-those studied (and solved) by \citep{murray}.
+Therefore, this is a `doubly intractable' problem \citep{murray}. An integral, whose
+result is a function of the parameters, appears in the {\em denominator} of the
+target density. However, in our case the integral in the
+denominator arises from the prior for the parameters, rather than from the
+likelihood function.
 
@@
 EC worked on the biased sampling suggestion in Section~\ref{sec:rationale},
-discovered the failure mode in Section~\ref{sec:failure_modes}, ...
+discovered the failure mode in Section~\ref{sec:failure_modes}, and was involved
+in many other discussions about the project. RB suggested the statistical
+mechanics example and helped BJB to get it working and to interpret the results.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-09 Paper/ms.tex
Rob's words
Fixed LJ subscript
entropy prior talk

@@
 
+In terms of the Lennard Jones parameters $\epsilon$ and $\sigma$, the
+``polymer'' potential is
+\begin{eqnarray}
+U_{\rm polymer} &=&
+\sum_{i=1}^{n-1} \phi^{(1)}_{i,i+1} + \sum_{i=2}^{n-1} \phi^{(2)}_{i-1, i, i+1}
+\end{eqnarray}
+where
+
+\begin{eqnarray}
+\phi^{(1)}_{i,i+1} &=& \frac{k}{2}\left(|\boldsymbol{r}_i - \boldsymbol{r}_{i+1}|- 2^{1/6}\sigma\right)^2,\\
+\phi^{(2)}_{i-1, i, i+1} &=& \frac{k}{2}
+\left((\boldsymbol{r}_i - \boldsymbol{r}_{i-1})\cdot(\boldsymbol{r}_{i-1} - \boldsymbol{r}_{i}) - 1\right)^2,
+\end{eqnarray}
+and $k=\frac{36 \times 2^{2/3}\epsilon}{\sigma^2}$ has been chosen to match
+the curvature of $U_{\rm LJ}(\boldsymbol{r}_i, \boldsymbol{r}_j)$.
+
 \section{Example 3: ``MaxEnt'' image deconvolution example}
@@
 the blurring operator written in matrix form.
+
+The prior for the pixel intensities $\{m_i\}$ has an ``entropic'' form,
+where images are upweighted (relative to a uniform prior) if the pixel
+intensities have a low dynamic range. The prior density is
+\begin{eqnarray}
+p(\boldsymbol{m}) \propto \exp
+\left[
+-\lambda \sum_{i=1}^n m_i \ln m_i
+\right]
+\end{eqnarray}
+where $\lambda$ is a hyperparameter.
+
+In the 1980s, this prior was thought to have a fundamental status, but this
+is no longer the case \citep{massinf}. Instead, we can consider the configurational
+entropy $-\sum_{i=1}^n m_i \ln m_i$ as simply a function of an image that may
+be relevant somehow, and that we are uncertain about.
 
@@
 
+\bibitem[\protect\citeauthoryear{Skilling}{1998}]{massinf}
+Skilling J., 1998, Massive Inference and Maximum Entropy, in Maximum Entropy
+and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14
+
 \bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.

### 2016-02-09 Paper/ms.tex
Condition on lambda

@@
 \begin{eqnarray}
-p(\boldsymbol{m}) \propto \exp
+p(\boldsymbol{m} | \lambda) \propto \exp
 \left[
@@
 where $\lambda$ is a hyperparameter.
-
 In the 1980s, this prior was thought to have a fundamental status, but this
@@
 \begin{eqnarray}
-p(\boldsymbol{m} | \boldsymbol{D})
-&=& \frac{p(\boldsymbol{m})p(\boldsymbol{D}|\boldsymbol{m})}
-{\int p(\boldsymbol{m})p(\boldsymbol{D}|\boldsymbol{m}) \,d^n\boldsymbol{m}}\\
+p(\boldsymbol{m} | \boldsymbol{D}, \lambda)
+&=& \frac{p(\boldsymbol{m} | \lambda)p(\boldsymbol{D}|\boldsymbol{m}, \lambda)}
+{\int p(\boldsymbol{m}|\lambda)p(\boldsymbol{D}|\boldsymbol{m}, \lambda) \,d^n\boldsymbol{m}}\\
 &=& \frac{\exp\left[-\lambda \sum_i m_i\ln m_i
@@
 
-Therefore, this is a `doubly intractable' problem \citep{murray}. An integral, whose
+This is a `doubly intractable' problem \citep{murray}, since an integral, whose
 result is a function of the parameters, appears in the {\em denominator} of the

### 2016-02-09 Paper/ms.tex
Updated abstract

@@
 problems arise frequently in statistical mechanics and occasionally in
-Bayesian inference. We demonstrate the algorithm on a toy example where
-$Z(\beta_1, \beta_2)$ is known, and show empirically that the accuracy
-improves as XX as a function of the number of particles $N$. We also
-calculate the partition function for a Lennard-Jones system in an external
-field, as a function of the external field strength, in a single run.}
+Bayesian inference. We demonstrate the algorithm on three example problems.
+The first is a simple toy example where
+$Z(\beta_1, \beta_2)$ is known, where we show empirically how the accuracy
+improves as a function of the number of particles $N$. We also
+calculate the partition function for a collection of atoms with a potential
+that is `intermediate' between a Lennard-Jones form and a polymer form, in
+a single run. Finally, we present a Bayesian image reconstruction example,
+where we can infer the value of a hyperparameter even though its prior
+distribution contains an `intractable' normalizing constant.}
 

### 2016-02-10 Paper/ms.tex
ht!
Some notational fixes

@@
 \begin{eqnarray}
-\mathcal{H} &=& \int p(\xx | \data, \II) \ln
+H &=& \int p(\xx | \data, \II) \ln
 \left[\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right] \, d\xx \\
@@
 \end{eqnarray}
-$\mathcal{H}$ quantifies how compressed the posterior distribution is with
-respect to the prior. For example, $\mathcal{H} = 10$ implies (loosely speaking)
+$H$ quantifies how compressed the posterior distribution is with
+respect to the prior. For example, $H = 10$ implies (loosely speaking)
 that the posterior occupies a fraction $e^{-10}$ of the prior mass.
-$\mathcal{H}$ can be interpreted quite literally as a measure of how much
+$H$ can be interpreted quite literally as a measure of how much
 the data $\data$ told you about $\xx$.
@@
 to accurately estimate $Z(\beta_1, \beta_2)$.
-\begin{figure}
+\begin{figure}[ht!]
 \centering
@@
 
-\begin{figure}
+\begin{figure}[ht!]
 \centering
@@
 \section{Example 3: ``MaxEnt'' image deconvolution example}
+
 Problems with more than one objective function can also arise in data
@@
 The unknown parameters are a set of pixel intensities
-$\{m_i\}_{i=1}^n$ where $n$ is the number of pixels and $m_i \geq 0$.
+$\{x_i\}_{i=1}^n$ where $n$ is the number of pixels and $x_i \geq 0$.
 The dataset is also a collection of pixel intensities $\{D_i\}_{i=1}^n$.
 The prior for the dataset given the parameters is a normal distribution
-with mean given by the underlying image $\{m_i\}$ convolved by
+with mean given by the underlying image $\{xi\}$ convolved by
 a point spread function. The standard deviation of each pixel is
@@
 \begin{eqnarray}
-\boldsymbol{D} = \boldsymbol{P}\boldsymbol{m} + \boldsymbol{n}
+\boldsymbol{D} = \boldsymbol{M}\boldsymbol{x} + \boldsymbol{n}
 \end{eqnarray}
-where $\boldsymbol{n}$ is the noise in each pixel and $\boldsymbol{P}$ is
+where $\boldsymbol{n}$ is the noise in each pixel and $\boldsymbol{M}$ is
 the blurring operator written in matrix form.
 
-The prior for the pixel intensities $\{m_i\}$ has an ``entropic'' form,
+The prior for the pixel intensities $\{x_i\}$ has an ``entropic'' form,
 where images are upweighted (relative to a uniform prior) if the pixel
@@
 \begin{eqnarray}
-p(\boldsymbol{m} | \lambda) \propto \exp
+p(\boldsymbol{x} | \lambda) \propto \exp
 \left[
--\lambda \sum_{i=1}^n m_i \ln m_i
+-\lambda \sum_{i=1}^n x_i \ln x_i
 \right]
@@
 is no longer the case \citep{massinf}. Instead, we can consider the configurational
-entropy $-\sum_{i=1}^n m_i \ln m_i$ as simply a function of an image that may
+entropy $-\sum_{i=1}^n x_i \ln x_i$ as simply a function of an image that may
 be relevant somehow, and that we are uncertain about.
@@
 \begin{eqnarray}
-p(\boldsymbol{m} | \boldsymbol{D}, \lambda)
-&=& \frac{p(\boldsymbol{m} | \lambda)p(\boldsymbol{D}|\boldsymbol{m}, \lambda)}
-{\int p(\boldsymbol{m}|\lambda)p(\boldsymbol{D}|\boldsymbol{m}, \lambda) \,d^n\boldsymbol{m}}\\
-&=& \frac{\exp\left[-\lambda \sum_i m_i\ln m_i
--\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}m_{j}\right)^2\right]}
-{\int \exp\left[-\lambda \sum_i m_i\ln m_i
--\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}m_{j}\right)^2\right]\, d^n\boldsymbol{m}}
+p(\boldsymbol{x} | \boldsymbol{D}, \lambda)
+&=& \frac{p(\boldsymbol{x} | \lambda)p(\boldsymbol{D}|\boldsymbol{x}, \lambda)}
+{\int p(\boldsymbol{x}|\lambda)p(\boldsymbol{D}|\boldsymbol{x}, \lambda) \,d^n\boldsymbol{m}}\\
+&=& \frac{\exp\left[-\lambda \sum_i x_i\ln x_i
+-\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}x_{j}\right)^2\right]}
+{\int \exp\left[-\lambda \sum_i x_i\ln x_i
+-\frac{1}{2\sigma^2}\left(D_i - \sum_{j} P_{ij}x_{j}\right)^2\right]\, d^n\boldsymbol{m}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathcal{H} &=& \sum_{i=1}^2 P(M_i | D)
+H &=& \sum_{i=1}^2 P(M_i | D)
 \ln\left[
@@
 
-The expected value of $\mathcal{H}$ (with respect to the prior over datasets)
+The expected value of $H$ (with respect to the prior over datasets)
 is the mutual information:
 \begin{eqnarray}
-\mathcal{I} &=& \mathds{E}\left[\mathcal{H}\right]\\
+\mathcal{I} &=& \mathds{E}\left[H\right]\\
 &=&

### 2016-02-10 Paper/ms.tex
working on image section

@@
 where images are upweighted (relative to a uniform prior) if the pixel
-intensities have a low dynamic range. The prior density is
+intensities have a low dynamic range. The prior density is defined conditional
+on a hyperparameter $\lambda$:
 \begin{eqnarray}
-p(\boldsymbol{x} | \lambda) \propto \exp
-\left[
--\lambda \sum_{i=1}^n x_i \ln x_i
-\right]
+p(\boldsymbol{x} | \lambda) &=&
+\frac{\exp\left[\lambda S(\xx)\right]}{\int \exp\left[\lambda S(\xx)\right] \, d\xx}
+\label{eq:entropy_prior}
 \end{eqnarray}
-where $\lambda$ is a hyperparameter.
+where
+\begin{eqnarray}
+S(\xx) &=& -\sum_{i=1}^n x_i \ln x_i
+\end{eqnarray}
+is the ``configurational entropy'' of the image (crucially, this is a function of
+an image, and not a function of a probability distribution).
+The denominator in Equation~\ref{eq:entropy_prior} ensures the conditional prior
+is normalized.
+
 In the 1980s, this prior was thought to have a fundamental status, but this
-is no longer the case \citep{massinf}. Instead, we can consider the configurational
-entropy $-\sum_{i=1}^n x_i \ln x_i$ as simply a function of an image that may
-be relevant somehow, and that we are uncertain about.
+is no longer the case \citep{massinf}, although the function $S(\xx)$ does
+have special properties when its argument is a measure \citep{knuth_skilling}.
+However, it is still a prior whose
+consequences we might be interested in.
+
+Instead, we can consider the configurational
+entropy $S(\xx)$ as simply a function of an image that may
+be relevant somehow, and that we are uncertain about. The joint posterior
+for $\xx$ and $\lambda$ is:
+\begin{eqnarray}
+p(\lambda, \xx | \data) &=&
+\frac{p(\lambda)p(\xx | \lambda)p(\data | \xx)}
+{p(\data)}\\
+&=&
+\frac{p(\lambda)p(\xx | \lambda)p(\data | \xx)}
+{p(\data)}
+\end{eqnarray}
 
@@
 
+\bibitem[Knuth and Skilling(2010)]{knuth_skilling} Knuth, K.~H., 
+Skilling, J.\ 2010.\ Foundations of Inference.\
+Axioms 1.1 (2012): 38-73.\
+ArXiv e-prints arXiv:1008.4831.
+
 \bibitem[Murray(2012)]{murray} Murray, Iain, Ghahramani, Zoubin, and MacKay, David.

### 2016-02-18 Paper/ms.tex
An appendix describing how to calculate uccs

@@
 
-%\appendix
-%\section{Appendix Title}
-%
-%Main text.
+\appendix
+\section{Appendix 1: Fast computation of upper corner counts}
+The upper corner count of particle $i$ is defined as the value of
+$\hat{X}$ at the position of particle $i$. Less formally, it is the number
+of other particles that are `up and to the right' of particle $i$ in
+a plot like Figure~\ref{fig:joint1}.
+
+\begin{eqnarray}
+C_i = \sum_{\substack{j=1\\j \neq i}}^N
+\mathds{1}\left[L_1(\xx_j) \geq L_1(\xx_i), L_2(\xx_j) \geq L_2(\xx_i)\right]
+\end{eqnarray}
 

### 2016-02-23 Paper/ms.tex
Working on uccs

--- 2016-02-10 Paper/ms.tex
@@
 
-%\appendix
-%\section{Appendix Title}
-%
-%Main text.
+\appendix
+\section{Appendix 1: Fast computation of upper corner counts}
+The upper corner count of particle $i$ is defined as the value of
+$\hat{X}$ at the position of particle $i$. Less formally, it is the number
+of other particles that are `up and to the right' of particle $i$ in
+a plot like Figure~\ref{fig:joint1}, plus one.
+
+\begin{eqnarray}
+C_i &=& \sum_{j=1}^N
+\mathds{1}\left[L_1(\xx_j) \geq L_1(\xx_i), L_2(\xx_j) \geq L_2(\xx_i)\right]\\
+&=& \sum_{j=1}^N
+\mathds{1}\left[L_1(\xx_j) \geq L_1(\xx_i)\right]
+\mathds{1}\left[L_2(\xx_j) \geq L_2(\xx_i)\right]
+\end{eqnarray}
 

### 2016-02-23 Code/Julia/Sampler.jl

@@ -57,3 +57,12 @@ function sort_scalars!(sampler::Sampler)
 	return nothing
 end
 
+@doc """
+Calculate the ucc as a function of rank wrt two objective functions.
+""" ->
+function calculate_uccs(sampler::Sampler)
+	uccs = zeros(UInt16, (sampler.num_particles, sampler.num_particles))
+
+	return uccs
+end
+

### 2016-02-23 Code/Julia/main.jl

@@ -19,3 +19,7 @@ plt.scatter(sampler.ranks[:,1], sampler.ranks[:,2], marker="o", alpha=0.1)
 plt.axis([0, 1+sampler.num_particles+1, 0, 1+sampler.num_particles])
 plt.show()
 
+uccs = calculate_uccs(sampler)
+plt.imshow(uccs, interpolation="nearest", cmap="Blues")
+plt.show()
+

### 2016-03-10 Code/Julia/Models/Example.jl
Removed Julia implementation (was very far from complete0

@@ -1,66 +1,0 @@
-include("../Utils.jl")
-
-@doc """
-An object of this class represents a point in parameter space.
-There are functions defined to evaluate the log likelihood and
-move around.
-""" ->
-type Particle
-	params::Vector{Float64}
-end
-
-@doc """
-A constructor taking no parameters.
-""" ->
-function Particle()
-	return Particle(Array(Float64, (2, )))
-end
-
-@doc """
-Generate params from the prior
-""" ->
-function from_prior!(particle::Particle)
-	particle.params[1] = randn()
-	particle.params[2] = particle.params[1] + randn()
-	return nothing
-end
-
-@doc """
-Do a metropolis proposal. Return log(hastings factor for prior sampling)
-""" ->
-function perturb!(particle::Particle)
-	log_H = 0.0
-
-	which = rand(1:2)
-	if(which == 1)
-		log_H -= -0.5*particle.params[1]^2
-		particle.params[1] += randh()
-		log_H += -0.5*particle.params[1]^2
-	else
-		particle.params[2] -= particle.params[1]
-		log_H -= -0.5*particle.params[2]^2
-		particle.params[2] += randh()
-		log_H += -0.5*particle.params[2]^2
-		particle.params[2] += particle.params[1]
-	end
-	return log_H
-end
-
-@doc """
-Evaluate the two objective functions
-""" ->
-function calculate_scalars(particle::Particle)
-	s = Array(Float64, (2, ))
-	s[1] = particle.params[1]
-	s[2] = particle.params[2]
-	return s
-end
-
-@doc """
-Convert to string, for output to sample.txt
-"""
-import Base.string
-function string(particle::Particle)
-	return join([string(signif(x, 6), " ") for(x in particle.params)])
-end
-

### 2016-03-10 Code/Julia/Particle.jl

@@ -1,14 +1,0 @@
-@doc """
-Particle class
-""" ->
-type Particle
-	params::Array{Float64, 1}
-end
-
-@doc """
-Constructor taking no arguments
-""" ->
-function Particle()
-	return Particle()
-end
-

### 2016-03-10 Code/Julia/Sampler.jl

@@ -1,133 +1,0 @@
-@doc """
-An object of this class is a Sampler, obviously.
-""" ->
-type Sampler
-    iteration::UInt64
-
-	# The walkers/particles
-	num_particles::Int64
-	particles::Array{Particle, 1}
-	scalars::Array{Float64, 2}
-
-    # Break ucc ties with these
-    tiebreakers::Array{Float64, 1}
-
-	# Stuff related to the ordering of the scalars
-	# by scalar 1 (first column) and then by scalar two (second column)
-	indices::Array{Int64, 2}
-	ranks::Array{Int64, 2}
-    uccs::Array{UInt16, 2}
-    particle_uccs::Array{UInt16, 1}
-
-    # Forbidden rectangles
-    forbidden_rectangles::Array{Float64, 2}
-end
-
-@doc """
-A constructor. Input the number of particles.
-""" ->
-function Sampler(num_particles::Int64)
-	return Sampler(0, num_particles, Array(Particle, (num_particles, )),
-									Array(Float64, (num_particles, 2)),
-                                    Array(Float64, (num_particles, )),
-									Array(Int64, (num_particles, 2)),
-									Array(Int64, (num_particles, 2)),
-                                    Array(UInt16, (num_particles, num_particles)),
-                                    Array(UInt16, (num_particles, )),
-                                    transpose([-Inf -Inf]))
-end
-
-@doc """
-Generate all the particles from the prior.
-""" ->
-function initialise!(sampler::Sampler)
-	# Generate all the particles from the prior
-	# and evaluate their scalars
-	for(i in 1:sampler.num_particles)
-		sampler.particles[i] = Particle()
-		from_prior!(sampler.particles[i])
-		sampler.scalars[i, :] = calculate_scalars(sampler.particles[i])
-        sampler.tiebreakers[i] = rand()
-	end
-
-	sort_scalars!(sampler)
-    calculate_uccs!(sampler)
-	return nothing
-end
-
-@doc """
-Sort the scalars. Store various kinds of results
-""" ->
-function sort_scalars!(sampler::Sampler)
-	# Do an argsort
-	sampler.indices[:,1] = sortperm(sampler.scalars[:,1])
-	sampler.indices[:,2] = sortperm(sampler.scalars[:,2])
-
-	# Use the argsort resuls to compute
-	# the ranks of the original array elements
-	for(i in 1:sampler.num_particles)
[…]

### 2016-03-10 Code/Julia/Utils.jl

@@ -1,42 +1,0 @@
-@doc """
-Heavy tailed distribution used for proposals
-""" ->
-function randh()
-	return 10.0^(1.5 - 6.0*rand())*randn()
-end
-
-@doc """
-log(sum(exp(x)))
-""" ->
-function logsumexp(x::Array{Float64, 1})
-	top = maximum(x)
-	y = exp(x - top)
-	return log(sum(y)) + top
-end
-
-@doc """
-log(exp(a) - exp(b))
-""" ->
-function logdiffexp(a::Float64, b::Float64)
-	@assert a > b
-	if(b == -Inf)
-		return a
-	end
-	return b + log(exp(a - b) - 1.)
-end
-
-@doc """
-Compare x and y, both vectors of floats (scalars).
-Returns 1 if x is higher, -1 if x is lower, and 0 if they're not
-comparable.
-""" ->
-function compare(x::Vector{Float64}, y::Vector{Float64})
-    @assert length(x) == length(y)
-    if all(x .> y)
-        return 1
-    elseif all(x .< y)
-        return -1
-    end
-    return 0
-end
-

### 2016-03-10 Code/Julia/main.jl

@@ -1,24 +1,0 @@
-# Setup plotting and include other files
-using PyCall
-@pyimport matplotlib.pyplot as plt
-include("Models/Example.jl")
-include("Sampler.jl")
-
-# Create and initialise a Sampler
-sampler = Sampler(1000)
-initialise!(sampler)
-update!(sampler)
-
-# Plot scalars and ranks
-plt.figure(1, figsize=(13, 6))
-plt.subplot(1, 2, 1)
-plt.scatter(sampler.scalars[:, 1], sampler.scalars[:, 2], marker="o", alpha=0.1)
-plt.subplot(1, 2, 2)
-plt.scatter(sampler.ranks[:,1], sampler.ranks[:,2], marker="o", alpha=0.1)
-plt.axis([0, 1+sampler.num_particles+1, 0, 1+sampler.num_particles])
-
-# Plot upper corner counts
-plt.figure(2)
-plt.imshow(sampler.uccs, interpolation="nearest", cmap="viridis")
-plt.show()
-

### 2016-03-10 Paper/ms.tex
Working on the beginning of the paper

@@
 \DeclareUnicodeCharacter{00A0}{ }
-\newcommand{\xx}{\boldsymbol{x}}
-\newcommand{\dx}{d\boldsymbol{x}}
+\newcommand{\xx}{\boldsymbol{\theta}}
+\newcommand{\dx}{d\boldsymbol{\theta}}
 \newcommand{\data}{\boldsymbol{D}}
@@
 % Authors (Add full first names)
-\Author{Brendon J. Brewer$^{1,}$*, Ewan Cameron$^{2}$, and
+\Author{Brendon J. Brewer$^{1,}$*,
+        Ewan Cameron$^{2}$, and
         Robert J. N. Baldock$^{3}$}
@@
 $^{2}$ Spatial Ecology and Epidemiology Group, Tinbergen Building, Department
-of Zoology, University of Oxford, South Parks Road, Oxford, UK\\
+of Zoology,
+        University of Oxford, South Parks Road, Oxford, UK\\
 $^{3}$ Cavendish, University of Cambridge, Cambridge, UK}
@@
 % Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{We present a generalization of the Nested Sampling algorithm for
-estimating the partition function $Z(\beta_1, \beta_2)$ of a family of
-probability distributions indexed by two hyperparameters $\beta_1$ and
-$\beta_2$. These
-problems arise frequently in statistical mechanics and occasionally in
-Bayesian inference. We demonstrate the algorithm on three example problems.
-The first is a simple toy example where
-$Z(\beta_1, \beta_2)$ is known, where we show empirically how the accuracy
-improves as a function of the number of particles $N$. We also
+\abstract{We present a generalization of the Nested Sampling algorithm,
+which we call ``{\em TwinPeaks}'',
+that estimates
+the partition function $Z(\beta_1, \beta_2)$ for a family of
+probability distributions indexed by two hyperparameters, $\beta_1$ and
+$\beta_2$. Problems like this
+arise frequently in statistical mechanics and occasionally in
+Bayesian inference.
+We demonstrate the algorithm on three example problems.
+The first is a toy problem whose structure allows us
+to accurately calculate $Z(\beta_1, \beta_2)$ numerically using non-Monte-Carlo
+methods.
+We show empirically how the accuracy of TwinPeaks
+improves as a function of the number of particles used, $N$. We also
 calculate the partition function for a collection of atoms with a potential
@@
 Its key strength its ability to cope with phase transitions which cause
-many other methods to fail.
-In a Bayesian inference framework with unknown parameters denoted collectively
+problems for many other methods \citep{skilling}.
+In a Bayesian inference problem with unknown parameters denoted collectively
 by a vector $\xx$, the
@@
 as a function of the enclosed prior mass $X$. These cannot literally be the
-same function as the inputs are not even necessarily of the same dimensionality.
-Therefore the symbol $L$ implies that the output of the function is a
+same function as the inputs typically don't have the same dimensionality
+($\xx$ is usually a collection of quantities and $X$ is a single real number
+in $[0, 1]$).
+Therefore the symbol $L$ implies the output of the function is a
 likelihood value; so $L(\xx)$ represents a likelihood value computed from
@@
 (the prior for the $L$ value implied by $\pi(\xx)$). Hence, there is a strong
-connection between NS and ideas from rare event simulation, which has been
-explicitly investigated by \citet{walter}.
+connection between NS and ideas from rare event simulation. \citet{walter} has
+explicitly explored this connection.
 

### 2016-03-10 Paper/ms.tex
Further writing

@@
 
-$Z$ is useful because it is the probability of the data given that
+$Z$ is useful because it is the (prior) probability of the data given that
 $\xx$ is {\em one of}
@@
 Using a single NS run, we can also calculate the properties of any other distribution that
-is in some sense intermediate
+is (in some sense which we do not make precise) {\em intermediate}
 between the prior and the posterior. For example, we might be interested in
-a `power posterior' where the likelihood is raised to a power $\beta$:
+a `power posterior' where the likelihood is raised to a power $\beta$
+(this is also the family of distributions used in `annealing' or `tempering'
+methods):
 \begin{eqnarray}
@@
 instead of the usual $L(\xx)$ used to obtain the posterior.
-An example application, in the case of
-a Bayesian inference problem with a `gaussian noise' likelihood assumption,
-computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
+Bayesian inference problems with `gaussian noise' likelihood assumptions
+provide an example of where power posteriors might be useful.
+Computing $p(\xx; \beta)$ for $\beta \neq 1$ allows us to explore what the
 posterior distribution would have been if the noise variance had been greater,
@@
 to be greater, because NS allows you to test the consequences of values of the
-variance that are very implausible given the data.
+variance that are very implausible given the data, i.e. that do not have
+a high posterior probability.
 
@@
 tuning parameters (just the number of particles $N$ is inherent to NS, although
-most NS implementations also have some tuning parameters related to their
+most NS implementations have additional tuning parameters related to their
 methods for sampling Equation~\ref{eq:constrained_prior}). Annealing methods
 also tend to fail on phase change problems \citep{skilling} which can arise
-in data analysis \citep{rjobject, exoplanet}.
+in data analysis \citep{rjobject, exoplanet} as well as in statistical
+mechanics, where they are more familiar.
 
